### 2016-10-27 paper.tex
finally added all the files

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{insval}
\usepackage{natbib}

\title{Fast inhibitory loop explains functional map development with short-range inhibition.}
\author{J\'an Antol\'ik, Andrew P. Davison}

\begin{document}
 \maketitle
\loadkeyvaluestore{/home/jan/Doc/Papers/lsv1m-paper/key_value_store.py}

\section{Introduction}

Competition between populations of neurons has been proposed as one of the canonical computations of cortical 
networks, and has been hypothesized to underly range of brain functions including working memory \cite{Amit1995,Durstewitz2000},
orientation tuning \cite{Somers1995,Ben-Yishai1995}, and functional map development \cite{VonderMalsburg1973,Antolik2011}.
In developmental models of functional cortical organization the competition occurs between populations of neurons spatially offset 
along the cortical surface, and previous studies have shown that a specific arrangement of the effective inhibitory and excitatory 
interactions is required: inhibition has to have longer range than excitation (a so called Mexican-Hat profile) \cite{Muir2014,?}.
This is, however, in stark contrast to the know anatomical findings showing that it is the excitatory neurons that tend to form 
long-range arborizations spanning multiple columns while the axons of majority of inhibitory neurons are confined to area
only several hundreds of micrometers in diameter.

As pointed out by Muir et al. in their recent study \cite{Muir2014}, all previous models of ??? have either relied on the anatomically 
unsupported Mexican-hat profile of lateral interactions \cite{?,?}, or relied on other biologically unrealistic properties, such as  ... .
Furthermore, model analysis by Muir et al \cite{Muir2014} showed that the presence of competition across the cortical surface 
is predicted well by the anatomy of direct excitatory and inhibitory coupling and that multisynaptic network effects are negligible.	
In conclusion, currently no satisfactory explanation of how topological functional organization develops in cortical networks 
that is consistent with the present anatomical findings exists.

All previous models of ??? have either assumed instantaneous synaptic transmission \cite{?,?} or uniform transmission delays
across the different intra-cortical connections \cite{?,?}. This is also the case of the analysis performed by Muir et al. \cite{?}
which questions the ability of intra-cortical connections to explain cortical competition. However, a recent study by ? et al. \cite{?}
have revealed a specific pattern of transmission delays between different neural type combinations. Specifically they found that the
excitatory to excitatory connections are slow while the excitatory to inhibitory connections are fast. In this study we explore the 
possibility that this specific transmission delay pattern is the missing link that can explain how short range inhibition can lead 
to effective cortical competition. We employ computational models to show that the fast inhibitory loop allows for the disynptic
inhibition to be effective and consequently provides the effective Mexican-hat like lateral interactions, thus for the first time 
explaining the development of functional organization in cortex with no biologically implausible assumptions. To demonstrate the
proposed properties, in this study we focus on models of development of orientation maps, because this is the most well explored
example of competition driven functional cortical organization development, but these results generalize to development of other
cortical topological properties, as well as potentially other competition based computations.

\section{Materials and Methods}

Here we describe the two models and their variants used in this study and finish with description of a measure
for assessing the extent to which a model orientation maps resamble their biological counterparts. Since this study heavily leans on methodology 
developed in our previous studies, we offer here only shortened description and refer readers to the original articles for details.

\subsection{GCAL model}

In section ? through ? we use GCAL model \cite{Stevens2013}  which is a recently developed variant of LISSOM 
(Laterally Interconnected Synergetically Self-Organizing Map) algorithm introduced by Miikkulainen et al. \cite{CMVC}, 
which itself is based on earlier Self-Organizing Map models \cite{Kohonen1982}.

For clarity and consistency the following model description closely follows the methodoly sections in our 
previous work \cite{Stevens2013,Antolik2011}. The architecture of the 3 GCAL variants presented in this 
paper is depicted in figure \ref{?}. The models are implemented in the Topographica simulator, freely available at www.topographica.org. 
Each of the GCAL models consists of set of sheets of model neural units,
corresponding to: (1) the photoreceptors, (2) the combined effects of the retinal ganglion cells (RGC) and ON and OFF LGN cells.
Furthermore each model has one V1 sheet of units with direct excitatory and inhibitory interactions, or two 
V1 sheets, one corresponding to excitatory and one to inhibitory neurons.
The model sheet is a 2D array of computational elements (called units or, loosely, neurons), with activation and plasticity equations as described below
and referenced by a coordinate system we will refer to as sheet coordinates, where the center of the sheet is designated (0.0,0.0).  
The number of units simulated in each sheet is determined by the density of units per unit length in both sheet dimensions. 
All cortical sheets have nominal dimensions \insval{cortex_size}$\times$\insval{cortex_size} in sheet coordinates. The sizes of the
RGC/LGN (\insval{lgn_size}$\times$\insval{lgn_size}) and photoreceptor (\insval{visual_field_size}$\times$\insval{visual_field_size}) sheets
were chosen to ensure that each unit in the receiving sheet has a complete set of connections, thus minimizing edge effects in the
RGC/LGN and V1 sheets. The density of units per $1.0\times$1.0 area is \insval{cortex_density}$\times$\insval{cortex_density} 
for the photoreceptor and RGC/LGN ON and OFF sheets, and \insval{cortex_density}$\times$\insval{cortex_density} for both cortical sheets.

\subsection{Simulation run-time and stimuly}

The simulator operates in discrete time steps. Retinal input changes every \insval{settle_steps} time steps (and during this period is kept constant), and
therefore afferent inputs to the V1 sheet(s) are effectively updated every 20 steps. This process is a discrete simulation of an
otherwise continuous series of changes in membrane potential due to incoming spikes and consequent generation of spikes.

The activation value $\psi_{i}$ of unit i in the photoreceptor sheet (P) is given by the gray-scale value in the chosen image at that point.
In this study we use two-dimensional elongated Gaussian patterns, who's center coordinates and orientation is sampled from uniform 
distribution that covers the area of photoreceptor sheet and full range of orientations. Each 20 time steps \insval Gaussian patterns are superimposed
to form the visual input.

\subsection{The LGN/RGC ON and OFF sheets}

\subsection{Rate model}

\subsection{Orientation map quality measure}

In order to asses whether the proposed model develops orientation maps that match the structure of those found in real 
animals we need to utilize an automatic metric that tells how close the maps are to animal data. To this end we will 
utilize a map-quality metric that we have recently developed \cite{Stevens2013} based on the empirical observation that pinwheel count 
in biological orientation maps scales linearly with hypercolumn size across many different species \cite{Kaschube2010}.
Specifically the pinwheel density per hypercolumn area (\Lambda^2) converges to \pi, when averaged across sufficently 
large cortical surface. For a detailed description of the procedure for calculating this metric we refer reader to 
our previous work \cite{Stevens2013} but briefly its calculation involves three steps. First, the locations of the pinwheels
in the orientation map are detrmined as the intersections of the zero contours of the real and imaginary 
components in the polar representation of the maps, thus yielding the total pinwheel count in the map. Second the 
hypercolumn size is determined as the peak in the isotropic ring-like fourier transform of the orientation maps.
Third, using these two number we can derive the pinwheel density, but to transform it to a usefull metric between 
unity (high-quality map) and zero (low quality map) we pass it through normalized gama distribution. In order to 
avoid any edge effects contaminating this measure, only the central ?x? area of the orientation maps is analyzed for
all models presented here. We have shown that this matric reliably distinguishes low and high quality maps \cite{Stevens2013}
and is a valid measure for assessing how well the model orientation maps match animal data.

\section{Results}

In this article we will proceed through multiple gradually more complex models of orientation development to address several key questions,
eventually demonstrating that the experimentally identified fast inhibitory loop is a satisfactory explanation for how short-range 
inhibition can support development of cortical functional organization. We will use two different computational abstractions 
to explore the questions at hand. In the fist part of the study we will use computational model that is derived from the LISSOM 
family of models \cite{CMVC}. This choice has three advantages. It allows for a very straightforward explanation of why 
fast inhibitory loop enables short-range inhibition to induce competition. It makes our explanation directly comparable to the 
extensive set of published LISSOM family models, and thus demonstrates that the solution proposed here generalizes to range of other 
functional properties. And finally the LISSOM abstraction enables very fast simulations, thus allowing us to perform 
parameter search analysis that would be otherwise computationally prohibitive. However, as we will explain in ?, some simplifications
made by the LISSOM abstraction, specifically the instantaneous translation of the neuronal input into it's activity, will leave 
certain questions unanswered. These will be addressed in section ? using more detailed rate model framework.

\section{Fast excitatory to inhibitory to excitatory loop enables competition in networks with short-range inhibitory connections.}

? et al. have shown that transmition delays between different types of cortical neurons are not uniform, specifically they found that transmission delays 
between excitatory neurons are ? ms, from excitatory to inhibitory neurons are ? ms, from inhibitory to excitatory neurons are ?. The sample size for connections
between inhibitory cells in the study was not large enough to be quantitatively reliable. The key observation here is that the 
combined di-synaptic delay from excitatory to inhibitory and inhibitory to excitatory cells is approximately as long as the mono-synaptic delay
of the excitatory to excitatory connections. Drawing from this, the core insight of this study is that under such pattern of delays
the effective inhibitory interactions are well approximated by the convolution of the excitatory and inhibitory connection kernels, thus 
giving the effective inhibition longer range, consequently fulfilling the essential requirement for cortical competition to occur.

For didactic purposes, let us first explore one example set of conditions under which the above statement holds exactly:

\begin{enumerate}

\item The sum of the excitatory to inhibitory , and inhibitory to excitatory delays is exactly the same as the excitatory to excitatory delay
\item There is no inhibitory to inhibitory interaction
\item The synaptic inputs into the excitatory and inhibitory neurons are instantaneously translated into their rate response via positive rectified transfer function.
\item The connection kernels of all neurons are Gaussian kernels with spatial constant $\sigma_{e}$ for excitatory neurons and $\sigma_{i}$ for inhibitory neurons.
\item We will disregard long-range (excitatory) connectivity, and assume that both local excitatory and local inhibitory connections have the same extent , thus $\sigma_{e} = \sigma_{i}$ 
\end{enumerate}

These conditions lead to following set of equations governing the cortico-cortical interaction.
\begin{align}
\label{eqn:first}
\begin{split}
& C_{m} \frac{dV}{dt} = \frac{(V_{rest} - V)}{R_{m}} + \frac{}{{R_{m}}}           g_{ex}(E_{ex} - V) + g_{inh}(E_{inh} - V)  \\

& R_{i}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\theta)]^+
\end{split}
\end{align}

\noindent where $R_{p}(x,t)$ is the response of neuron of type $p$ located at position $x$ at time $t$, $I_{aff}(x,t)$ is the afferent input to neuron $x$ at time $t$, $N_{\sigma}$(d) 
is normal distribution of variance $\sigma$, corresponding to the lateral connection kernel. When we expand for $R_{i}$ we obtain:

\begin{equation}
R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\alpha) - \sum_{y}N_{\sigma_{i}}(\lVert y-x \rVert)[\sum_{z}N_{\sigma_{e}}(\lVert z-y \rVert)R_{e}(z,t-\theta-\beta)]^+]^+
\end{equation}

\noindent Because $\theta + \beta = \alpha$ (assumption 1) and because the inhibitory neurons receive only excitatory connections (assumption ?) we can further simplify the above as:

\begin{equation}
R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\alpha) - \sum_{z}\sum_{y}N_{\sigma_{i}}(\lVert y-x \rVert) N_{\sigma_{e}}(\lVert z-y \rVert) R_{e}(z,t-\alpha)]^+
\end{equation}

\noindent and thus due to symmetry of normal distribution:

\begin{equation}
R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\alpha) - \sum_{z}N_{\sigma_{i}} \ast  N_{\sigma_{e}}(\lVert z-x \rVert) R_{e}(z,t-\alpha)]^+ 
\end{equation}

\noindent Because convolution of two Normal distributions of variance $\sigma_{i}$  and $\sigma_{e}$ is normal distribution with variance $\sqrt{\sigma_{i}^2 + \sigma_{e}^2}$ and 
because $\sigma_{i} = \sigma_{e}$ we can simplify to:

\begin{equation}
R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\alpha) - \sum_{z}N_{\sqrt{2}\sigma_{e}}(\lVert z-x \rVert) R_{e}(z,t-\alpha)]^+
\end{equation}

This shows that under these specific assumptions the effective inhibitory interactions between excitatory neurons are $\sqrt{2}$ longer than excitatory and thus
follow the mexican-hat like profile required for for lateral cortical competition underlying functional map development. Even though this essential condition of 
cortical competition is fulfilled in this model configuration, the requirement of local inhibition puts important constraints on the extent of the effective lateral
interactions (relative to lateral excitation). It thus remains unclear wether development of high-quality orientation maps as observed experimentally is 
supported under these conditions. Conveniently, the lateral interaction in the LISSOM family of models is governed by the same equations as ? and ?. 
In the following we will use the GCAL model \cite{?}, the latest and most robust addition to the LISSOM family of models to demonstrate that the above 
specific configuration of effective lateral excitation and inhibition permits development of high-quality orientation maps. To facilitate comparison we will use exactly 
the same model configuration as in ?, except to three modifications necessitated by the analysis above:

\begin{enumerate}

\item We will change the spread of lateral inhibition to be $\sqrt{2}$ longer than lateral excitatory spread, such that the resulting lateral
interactions conform with equation ?.

\item The change in point ? will results in change in the balance of overall excitation and inhibition in the model which is critical 
to correction functional development in the model. We will thus modify the strength of the lateral inhibition to compensate for changes in ?.

\item Analogously to ?, the changes in ? change the overall balance between feedforward and lateral contributions to model cortical neuron's activity, 
which we will compensate by changing the overall strength of the lateral interactions.

\end{enumerate}

\begin{figure}[htpb!] 
\centering
\includegraphics[width=16cm]{./SVG/Figure1/figure1.png}
\caption{Orientation map development with short range inhibition and fast excitatory-to-inhibitory-to-excitatory loop. (A) Map quality (see Materials and Methods) at a range of 
lateral excitatory vs inhibitory projection strength ratios and afferent vs lateral projection strength ratios. (B-D) Functional organization in 3 example parameter configurations 
indicated by the red marks. From left to right, the orientation map, fast-fourier transform of the orientation map and afferent connection fields from the ON LGN model sheet for 25 example model V1 neurons. 
(B-C) Two examples of sub-optimal orientation maps. (D) Model configuration with the highest quality map found in this parameter search. (E) The same model configuration but in this case ran with explicit
simulation of inhibitory neurons and corresponding connections. E and D are identical confirming correctnes of our analysis.}
\label{fig:figure1}
\end{figure}

In order to find working combination of parameters in point's 2 and 3, and also to show that the model is robust to certain level of changes
in these two parameters we have performed a parameter search across these two parameters and evaluated the quality of the orientation map (see ?) for each 
parameter combination. As figure ? shows, under a range of values of both parameters the model develops high-quality orientation maps indistinguishable
from their experimental counterparts, thus concluding our first step towards showing that fast excitatory-to-inhibitory-to-excitatory loop can explain 
how short-range inhibition can induce cortical competition and consequently development of topological organization of functional properties.

Furthermore, note that the GCAL model used in figure ? only explicitly models excitatory neurons and assumes both direct excitatory and inhibitory
interactions between them, thus corresponding to equation ?. Even though above we have shown that equation ? is equivalent to equation ? to verify 
correctness of our analysis we have ran a single simulation of the GCAL model corresponding to the parameter combination with the highest map 
quality found in figure ?, but with explicitly simulated inhibitory population. In this model we thus do not model direct inhibitory interactions
between excitatory neurons, but instead add excitatory to inhibitory and inhibitory to excitatory connections of the same extend as those of 
excitatory to excitatory (see the assumption \#?). As we have shown above this model should be mathematically equivalent to the simulations 
figure ?D, whic is confirmed in figure ?E that shows that the simulation results in formation of identical orientation map as in the figure ?D.

\section{Inhibitory to inhibitory connections are consistent with development of functional organization}

In the previous section we have shown that under a set of specific assumptions short-range inhibition can induce 
effective mexican-hat like interaction and thus support development of orientation maps. However not all assumptions we 
made were in line with experimental evidence. In this section we will show that one of these assumptions is not necessary, 
specifically that addition of inhibitory connections does not invalidate results of the previous section.

To this end we will use the exact GCAL model configuration that we have found in previous section to posses the highest
quality orientation map (see figure ?). We will use GCAL configuration in which we will explicitly model the inhibitory neurons
(see figure ?, and equation ?) and consequently also explicitly the excitatory to inhibitory and inhibitory to excitatory connections.
Furthermore, we will add inhibitory to inhibitory connections to the model with the same extent as that of inhibitory to excitatory
connections ($\sigma_{i} = \sigma_{e}$ in eq ?).

\begin{figure}[htpb!] 
\centering
\includegraphics[width=16cm]{./SVG/Figure2/figure2.png}
\caption{Orientation map development with short range inhibition, fast excitatory-to-inhibitory-to-excitatory loop and inhibitory to inhibitory connections. (A) Map quality (see Materials and Methods) at a range of 
inhibitory to inhibitory vs inhibitory to excitatory projection strength ratios and excitatory to inhibitory connection strengths. (B-D) Functional organization in 3 example parameter configurations 
indicated by the red marks. From left to right, the orientation map, fast-fourier transform of the orientation map and afferent connection fields from the ON LGN model sheet for 25 example model V1 neurons. }
\label{fig:figure2}
\end{figure}

By explicitly modelling the inhibitory neurons in this model we have replaced a single parameter gowerning the strength of inhibitory
later interactions in the model from previous section with three new parameters that set the strength of excitatory to inhibitory,
inhibitory to excitatory and inhibitory to inhibitory connections (all other parameters remaind the same as the 
in the best parametrization found in previous section). Note that in principle there is redundancy in these parameters as
the overall strength of the projections from inhibitory neurons onto both excitatory and inhibitory populations is scaled by the 
excitatory to inhibitory projections strength. Therefore in figure ? we have systematically varied the strength of the added inhibitory to inhibitory
projection expressed relatively to the strength of the inhibitory to inhibitory connections (which was set to 1), while also varying the 
strength of the excitatory to inhibitory projection. We have  investigated the quality of the orientation 
maps that developed under these different levels of inhibitory to inhibitory interactions. As can be seen, high quality maps 
can develop under the full range of the inhibitory to inhibitory interaction strengths, depending on the overall 
excitatory to inhibitory drive. This shows that inclusion of direct inhibitory to inhibitory interactions does not 
invalidate the results of section ?.

\section{Long-range excitation}

In section ? we have only assumed local connectivity by assuming that both excitatory and inhibitory interaction had the same spatial extent. 
However experimental evidence shows that excitatory cells send longer connections compared to inhibitory cells \cite{?}. In this section we will
explore what happens if we add long-range excitatory connectivity into the model analyzed in section ?. Buz\'as et al. \cite{Buzas2006} have shown that
the lateral connectivity in layer 2/3 can be best described as superimposition of two gabor connectivity likelyhoods, one short-range not orientation specific
and one long-range orientation specific. Here we will assume such dual structure, leading us to add second excitatory to excitatory and excitatory to inhibitory
projection into the model described in section ?, but with 3 times larger space constant (in line with Buz\'as et al. quantitative findings). In figure \ref{fig:figure3} we have 
kept the overall strength of excitatory to excitatory and excitatory to inhibitory pathways the same, corresponding to the best parametrization found in ?,
but varied the relative strength between the short range and long range components. As can be seen ...

\begin{figure}[htpb!] 
\centering
\includegraphics[width=16cm]{./SVG/Figure2/figure3.png}
\caption{}
\label{fig:figure3}
\end{figure}

\section{Non-equal effective excitatory and inhibitory delay}

In all model variants examined sofar we have made the key assumption that the delay on the excitatory to excitatory connections is exactly equal 
to the sum of excitatory to inhibitory and inhibitory to excitatory delays. This assumption is approximately supported by the experimental evidence \cite{?}, 
but we cannot assume it holds exactly in real biological substrate.  However, we hypothesize, that the small discrepancies between the cumulative delays 
of the mono-synaptic excitatory and duo?-synaptic inhibitory interactions can be absorbed in the membrane time-constant of the neurons. In this section we will 
verify this hypothesis by extending the modeling framework used thus far with a finite membrane time-constant (see Materials and Methods ?) and proceed to determine 
the magnitude of discrepancy in the delays between the excitatory and inhibitory interactions that can be managed by the model without impairing the resulting map orientation 
map quality. We use model parametrization similar to those determined in previous section ? to lead to high quality orientation maps. For simplicity and computational 
efficiency we omit the inhibitory to inhibitory and long-range excitatory connections that have already been investigated above. The membrane time-constant of excitatory 
neurons was set to ?ms while those of inhibitory neurons to ?ms. This faster inhibitory dynamics are necessary to prevent oscillations in the system \cite{?}. 
Figure ? shows the resulting orientation maps and associated map quality measures of ? models with range of differences between the monosynaptic excitatory and 
duo?-synaptic inhibitory interactions.

\begin{figure}[htpb!] 
\centering
\includegraphics[width=16cm]{./SVG/Figure4/figure4.png}
\caption{}
\label{fig:figure3}
\end{figure}

\bibliographystyle{abbrv}
\bibliography{./mendeley}

\end{document}

### 2016-10-27 CODE/generate_figure_material.py

@@ -237,7 +237,7 @@ def figure4():
     Y = []
     qual = []
 
-    if False:
+    if True:
         for a in os.listdir(dirr):
             b = os.path.join(dirr,a);
 
@@ -278,7 +278,7 @@ def figure4():
         pylab.savefig('figure4.png')
 	pylab.show()
     
-    if True:
+    if False:
 
        fname= '/home/jan/Doc/Papers/fast_inh_paper/DATA/GCAL_EI_II/exc_inh_strength=3.5_inh_inh_strength=0.2'
        f = open(fname+'/results.pickle') 
@@ -372,7 +372,7 @@ def figure5():
             sheet_size = numpy.sqrt(len(mmap))
             mmap = numpy.resize(mmap,(sheet_size,sheet_size))
             # lets find out the pinwheel density
-            mmap = mmap[:-1,:-1]
+            mmap = mmap[8:-9,8:-9]
             X.append(abs(float(a.split('_')[5])))
             maps[X[-1]] = mmap
             rho = pinwheel_analysis(mmap)['metadata']['rho']
@@ -393,7 +393,7 @@ def figure5():
         #cax = divider.append_axes("right", size="5%", pad=0.05)
         #pylab.colorbar(im,cax=cax)
         #pylab.colorbar()
-        pylab.savefig('figure5.png')
+        pylab.savefig('figure5.png',dpi=600)
         
         vis(maps[0.001])
         rho = pinwheel_analysis(maps[0.001])['metadata']['rho']
@@ -509,9 +509,8 @@ def a():
 
 
 #figure2()
-figure3()
-#figure4()
+#figure3()
+figure4()
 #a()
-
 #figure5()
 

### 2016-10-27 CODE/pinwheel_analysis.py

@@ -1,0 +1,394 @@
+#==================================#
+#  OR map contour indentification  #
+#==================================#
+import sys
+import numpy as np
+import cmath, math, itertools
+import pylab as plt
+from numpy.fft.fftpack import fft2
+from numpy.fft.helper import fftshift
+from scipy.optimize import curve_fit
+
+
+
+
+
+
+
+
+class WarningCounter(object):
+    """
+    A simple class to count 'divide by zero' and 'invalid value'
+    exceptions to allow a suitable warning message to be generated.
+    """
+    def __init__(self):
+        self.div_by_zeros=0
+        self.invalid_values=0
+
+    def __call__(self, errtype, flag):
+        if errtype == "divide by zero":  self.div_by_zeros += 1
+        elif errtype == "invalid value": self.invalid_values += 1
+
+    def warn(self):
+        total_events = self.div_by_zeros + self.invalid_values
+        if total_events == 0: return
+        info = (total_events, self.div_by_zeros, self.invalid_values)
+        self.div_by_zeros=0; self.invalid_values=0
+        message = ("Warning: There were %d invalid intersection events:"
+                   "\n\tNumpy 'divide by zero' events: %d"
+                   "\n\tNumpy 'invalid value' events: %d\n")
+        sys.stderr.write( message %info)
+
+
+
+
+#======================#
+# Image transformation #
+#======================#
+
+def normalize_polar_channel(polar_channel):
+    """This functions normalizes an OR map (polar_channel) taking into
+    account the region of interest (ROI). The ROI is specified by
+    values set to 99. Note that this functionality is implemented to
+    reproduce the experimental approach and has not been tested (not
+    required for Topographica simulations)
+    """
+
+    def grad(r):
+        (r_x,r_y) = np.gradient(r)
+        (r_xx,r_xy)=np.gradient(r_x);
+        (r_yx,r_yy)=np.gradient(r_y);
+        return r_xx**2+ r_yy**2 + 2*r_xy**2
+
+    roi = np.ones(polar_channel.shape) # Set ROI to 0 to ignore values of -99.
+    roi[roi == -99] = 0                # In Matlab: roi(find(z==-99))=0
+
+    fst_grad = grad(roi)
+    snd_grad = grad(fst_grad)
+
[…]

### 2016-11-25 paper.tex
changes

@@
 along the cortical surface, and previous studies have shown that a specific arrangement of the effective inhibitory and excitatory 
-interactions is required: inhibition has to have longer range than excitation (a so called Mexican-Hat profile) \cite{Muir2014,?}.
-This is, however, in stark contrast to the know anatomical findings showing that it is the excitatory neurons that tend to form 
+interactions is required: inhibition has to have longer range than excitation (a so called Mexican-Hat profile) \cite{Muir2014}.
+This is, however, in stark contrast to the know anatomical arrangement: the excitatory neurons superficial layers tend to form 
 long-range arborizations spanning multiple columns while the axons of majority of inhibitory neurons are confined to area
-only several hundreds of micrometers in diameter.
+only several hundreds of micrometers in diameter \cite{Buzas2006,Budd2001}.
 
-As pointed out by Muir et al. in their recent study \cite{Muir2014}, all previous models of ??? have either relied on the anatomically 
-unsupported Mexican-hat profile of lateral interactions \cite{?,?}, or relied on other biologically unrealistic properties, such as  ... .
-Furthermore, model analysis by Muir et al \cite{Muir2014} showed that the presence of competition across the cortical surface 
-is predicted well by the anatomy of direct excitatory and inhibitory coupling and that multisynaptic network effects are negligible.	
-In conclusion, currently no satisfactory explanation of how topological functional organization develops in cortical networks 
-that is consistent with the present anatomical findings exists.
+As pointed out by Muir et al.
+in their recent study \cite{Muir2014}, to solve this apparent conflict, previous topologically organized models cortical competition have either relied on the anatomically unsupported Mexican-hat profile of lateral interactions \cite{VonderMalsburg1973,CMVC}, or relied on other biologically unrealistic properties, such as selective targeting of inhibitory neurons by long range excitatory connections \cite{?} or instanteneous synaptic transmission coupled with ommission of recurrent inhibition \cite{Kang2003,Levy2011}.
+Furthermore, model analysis by Muir et al.
+\cite{Muir2014} showed that the presence of competition across the cortical surface is predicted well by the anatomy of direct excitatory and inhibitory coupling and that multisynaptic network effects are negligible.
+In conclusion, currently no satisfactory explanation of how topological functional organization develops in cortical networks that is consistent with the present anatomical findings exists.
 
-All previous models of ??? have either assumed instantaneous synaptic transmission \cite{?,?} or uniform transmission delays
-across the different intra-cortical connections \cite{?,?}. This is also the case of the analysis performed by Muir et al. \cite{?}
-which questions the ability of intra-cortical connections to explain cortical competition. However, a recent study by ? et al. \cite{?}
-have revealed a specific pattern of transmission delays between different neural type combinations. Specifically they found that the
+All previous topologically organized models of cortical competition have either assumed instantaneous synaptic transmission \cite{?,?} or uniform transmission delays across the different intra-cortical connections \cite{?,?}.
+This is also the case of the analysis performed by Muir et al.
+\cite{Muir2014} which questions the ability of intra-cortical connections to explain cortical competition.
+However, a recent study by Ohana et al.
+\cite{Ohana2012} have revealed a specific pattern of transmission delays between different neural type combinations.
+Specifically they found that the
 excitatory to excitatory connections are slow while the excitatory to inhibitory connections are fast. In this study we explore the 
@@
 inhibition to be effective and consequently provides the effective Mexican-hat like lateral interactions, thus for the first time 
-explaining the development of functional organization in cortex with no biologically implausible assumptions. To demonstrate the
+explaining competition in topologically organized cortical networks with no biologically implausible assumptions. To demonstrate the
 proposed properties, in this study we focus on models of development of orientation maps, because this is the most well explored
@@
 
-In section ? through ? we use GCAL model \cite{Stevens2013}  which is a recently developed variant of LISSOM 
+In sections ? through ? we use GCAL model \cite{Stevens2013}  which is the most advanced variant of LISSOM 
 (Laterally Interconnected Synergetically Self-Organizing Map) algorithm introduced by Miikkulainen et al. \cite{CMVC}, 
@@
 
-For clarity and consistency the following model description closely follows the methodoly sections in our 
+For clarity and consistency the following model description closely follows the methodology sections in our 
 previous work \cite{Stevens2013,Antolik2011}. The architecture of the 3 GCAL variants presented in this 
@@
 corresponding to: (1) the photoreceptors, (2) the combined effects of the retinal ganglion cells (RGC) and ON and OFF LGN cells.
-Furthermore each model has one V1 sheet of units with direct excitatory and inhibitory interactions, or two 
-V1 sheets, one corresponding to excitatory and one to inhibitory neurons.
-The model sheet is a 2D array of computational elements (called units or, loosely, neurons), with activation and plasticity equations as described below
-and referenced by a coordinate system we will refer to as sheet coordinates, where the center of the sheet is designated (0.0,0.0).  
+Furthermore each model has one V1 sheet of units with direct excitatory and inhibitory interactions (section \ref{sec:SM1}), or two 
+V1 sheets, one corresponding to excitatory and one to inhibitory neurons (sections \ref{sec:SM2} and \ref{sec:SM3}).
+The model sheet is a 2D array of computational elements (called units or, loosely, neurons), with activation and plasticity equations as 
+described below and referenced by a coordinate system we will refer to as sheet coordinates, where the center of the sheet is designated (0.0,0.0).  
 The number of units simulated in each sheet is determined by the density of units per unit length in both sheet dimensions. 
@@
 
-\subsection{Simulation run-time and stimuly}
+\subsubsection{Simulation run-time and stimuly}
 
-The simulator operates in discrete time steps. Retinal input changes every \insval{settle_steps} time steps (and during this period is kept constant), and
-therefore afferent inputs to the V1 sheet(s) are effectively updated every 20 steps. This process is a discrete simulation of an
-otherwise continuous series of changes in membrane potential due to incoming spikes and consequent generation of spikes.
+As a simplification, GCAL ignores the detailed temporal properties of the subcortical neural responses and of signal propagation along the various types of connections.
+Instead, the model ON/OFF units have a constant, sustained output, and all connections have a constant delay, independent of the physical length of that connection.
+The simulator operates in discrete time steps.
+Retinal input changes every \insval{settle_steps} time steps (and during this period is kept constant), and therefore afferent inputs to the V1 sheet(s) are effectively updated every \insval{settle_steps} steps.
+This process is a discrete simulation of an otherwise continuous series of changes in membrane potential due to incoming spikes and consequent generation of spikes.
+One such training iteration (\insval{settle_steps} steps) in the model represents one visual fixation i.e., an iteration consists of a constant retinal activation, followed by processing at the ON/OFF and cortical levels.
 
@@
 In this study we use two-dimensional elongated Gaussian patterns, who's center coordinates and orientation is sampled from uniform 
-distribution that covers the area of photoreceptor sheet and full range of orientations. Each 20 time steps \insval Gaussian patterns are superimposed
-to form the visual input.
+distribution that covers the area of photoreceptor sheet and full range of orientations.
+Each \insval{settle_steps} time steps \insval{num_gaussian_stim} Gaussian patterns are superimposed to form the visual input.
 
-\subsection{The LGN/RGC ON and OFF sheets}
+\subsubsection{The LGN/RGC ON and OFF sheets}
+
+The ON/OFF units are called RGC/LGN units because they represent the entire pathway between the retinal photoreceptors
+and V1, including the retinal ganglion cells, LGN cells, and the connection pathways. The activation level for a unit at position $j$
+in an RGC/LGN sheet at time $t$ is defined as:
+%%
+\begin{equation}
+\eta_{j}(t)=f\left(\frac{A_{j}(t)}{c + \gamma_{L}\sum_{k}A_{k}(t)l_{kj}}\right)
+\end{equation}
+%%
+where
+%%
+\begin{equation}
+A_{j}(t) = \gamma_{F}\sum_{i\in F_j}\Psi_{i}(t)\omega_{ij}
+\end{equation}
+%%
+The activation function $f$ is a half-wave rectifying function that ensures positive activation values, constant $\gamma_{F}=\insval{retina_to_lgn_strength}$ defines the overall strength of the afferent connections from the retina, constant $\gamma_{L}=\insval{lateral_lgn_strength}$ defines the strength of the lateral connections within the RGC/LGN sheet, constant $c=\insval{lateral_lgn_gain}$ defines the slope of the gain, $\Psi_{i}$ is the activation of unit $i$ taken from the set of photoreceptors
+from which RGC/LGN unit $j$ receives input (its connection field $F_j$), $\omega_{ij}$ is the connection weight from unit $i$ in the
+retina to unit $j$ in the RGC/LGN, and $l_{kj}$ is the lateral connection weight from RGC/LGN unit $k$ to RGC/LGN unit $j$.
+Weights from the photoreceptors to units in the ON and OFF channels are set to fixed strengths with a difference-of-Gaussians
+kernel ($\sigma_{center}=\insval{lgn_center_sigma}$, $\sigma_{surround}=\insval{lgn_surround_sigma}$, in sheet dimensions), with ON connection fields having a positive center and a negative surround and vice versa for OFF.
+The lateral RGC/LGN weights are 2D Gaussians with kernel size $\sigma=\insval{lgn_lateral_sigma}$.
+The center of the afferent and lateral connection field of each ON/OFF unit is mapped to the location in the photoreceptor and lgn sheet corresponding to the location of that unit in sheet coordinates, making all these projections retinotopic.
+
+\subsubsection{Cortical model}
+
+Units in the cortical sheets each receive three types of projections represented as matrices of weights: afferent excitatory $(p=A)$, 
+lateral excitatory $(p=E)$ and lateral inhibitory $(p=I)$.  The contribution $C_{jp}$ to the activation of unit $j$ in a cortical sheet from each projection $p$ at time $t$ is given by: 
+\begin{equation} 
+C_{jp}(t+\delta t)=\sum_{i\in F_{jp}}\Psi_{i}(t)\omega_{pij} 
+\end{equation}
+
+\noindent where $\Psi_{i}(t)$ is the activation of unit $i$ taken from the set of units in the input
+sheet of projection $p$ from which unit $j$ receives input (its connection field $F_{jp}$), and $\omega_{pij}$ is the connection
+weight from unit $i$ in the input sheet of projection $p$ to unit $j$ in the in the output sheet of projection $p$. All connection field
+weights are initialized with uniform random noise multiplied by a 2D Gaussian profile, cut off at the distance specified below.
+Contributions from each projection are weighted and summed and passed via a non-linearity $f$ to calculate the activation of a cortical neuron i at time $t$:
+%%
+\begin{equation}
+\Psi_{i}(t)= f(\sum_{p}\gamma_{p}C_{ip}(t))
+\end{equation}
+%%
+where $\gamma_{p}$ is a constant determining the sign (negative for inhibitory) and strength of projection $p$. 
+The transfer function $f$ is a half-wave rectifying function that ensures positive activation values. It has a 
+variable threshold point ($\theta$) dependent on the average activity of the unit as described in the next subsection, 
+but in all cases the gain is fixed at unity.
+Table \ref{table:param} shows the strength, initial Gaussian kernel spatial extent, and the cut-off distance values for all projections for the first form of the GCAL model examined in section \ref{sec:SM1}.
+Any modifications to these base parameters
+in the GCAL model variants examined in sections \ref{sec:SM2,sec:SM3} are then reported in the respective sections.
+
+Once all \insval{settle_steps} settling steps are complete, the settled cortica activation pattern is deemed to be the response of cortical sheets to 
+the presented pattern.
+At this point we use the response of cortical neurons to update their threshold point ($\theta$) (using the adaptation process described below) and to update the afferent weights via Hebbian learning.
+Cortical activity is then reset to zero, and a new pattern is presented.
+Note that both adaptation and learning could instead be performed at every settling step, but this would greatly decrease computational efficiency.
+
+\subsubsection{Homeostatic adaptation}
+
+The threshold $\theta$ of all cortical excitatory units is updated at the end of each settling phase based on following 
+equations:
+
+\begin{equation}
+	\theta_{t+1} = \theta_{t} + \xi (\tilde{\Psi}(t) - \mu) 
+\end{equation}
+
+where $\xi=\insva{homeostatic_learning_rate}$ is the time constant of the threshold adaptation,
+$\mu=\insval{target_activity}$ is a constant defining the target average activity, and $\tilde{\Psi}$ is 
+the recent average activity of the unit:
+
+\begin{equation}
+  \tilde{\Psi}(t) = (1 - \chi)\Psi(t) + \chi\tilde{\Psi}(t-1)  
+\end{equation}
+
+\noindent where $\Psi(t)$ is the output of the unit at time t and $\chi=\insval{activity_average_smoothing}$ is time constant 
+controling the decay of the influence of the past activities. The effect of this scaling mechanism is to bring the average activity of each 
+cortical unit closer to the specified target.
+If the activity in a V1 unit moves away from the target during training, the threshold for activation is thus automatically raised or lowered to bring it closer to the target.
+Note that an alternative rule with only a single smoothing parameter (rather than
+$\xi$ and  $\chi$) could be formulated, but the rule as presented here makes it simple for the modeler to set a desired target activity.
+
+\subsubsection{Hebbian adaptation}
+
+In the model, as images are presented to the photoreceptors, cortical afferent connection weights $\omega_{i,j,A} from the ON/OFF sheets are adjusted once per iterations (after cortical settling is completed) using a simple Hebbian learning rule.
+This rule results in connections that reflect correlations between the presynaptic ON/OFF unit activities and the postsynaptic cortical response.
+Hebbian connection weight adjustment at each iteration is dependent on the presynaptic activity, the postsynaptic response, and the Hebbian learning rate:
+
+\begin{equation}
+\omega_{ijA}(t)=\frac{\omega_{ij}(t-1)+\beta_{p}\Psi_{j}(t)\Psi_{i}(t)}{\sum_{p \in \{ON,OFF\}}\sum_{k}\left(\omega_{kj,p}(t-1)+\beta_{p}\Psi_{j}(t)\Psi_{k}(t)\right)}
+\end{equation}
+
+where $\beta_{p}$ is the Hebbian learning rate for the connection fields in the two afferent projections from RGC/LGN $p \in \{\mathrm{ON},\mathrm{OFF}\}$.
+I.e., the afferent weights from RGC/LGN are normalized jointly.
+Learning rate parameters are specified as
+a fixed value $\iota_{p}=\insval{learning_rate}$ for each projection, and then the unit-specific values used in the equation above are calculated as
+$\beta_{p}=\frac{\iota_{p}}{\upsilon_{p}}$, where $\upsilon_{p}$ is the number of connections per connection field in projection $p$.
 
 \subsection{Rate model}
+
+\subsection{Orientation map analysis}
+
+Model orientation maps are calculated based on the vector average method \cite{CMVC}. We first determine the prefered frequency of neurons 
+across the map. Due to the simplified stereotypical stimulus used in this study (elongated Gaussian inputs) the spatial frequency 
+preference of all neurons lies in a very narrow band, and we thus use the mean prefered spatial frequency across all cortical neurons as the
+value for the spatial frequency parameteter across all subsequent analysis.
+Next, sine grating inputs that cover the full range of remaining parameter values (combinations of all orientations and phases) are presented, and for each orientation, the peak response of the neuron is recorded.
+The orientation preference is calculated by constructing a vector for each orientation $\theta$ (between 0 and 180\circ), with the peak response as
+the length and $\theta$ as its orientation.
+These vectors are summed and the preferred orientation is calculated as half of the orientation of the summed vector.
+The selectivity is given by the magnitude of the summed vector.
 
@@
 large cortical surface. For a detailed description of the procedure for calculating this metric we refer reader to 
-our previous work \cite{Stevens2013} but briefly its calculation involves three steps. First, the locations of the pinwheels
+our previous work \cite{Stevens2013}, but briefly, its calculation involves three steps. First, the locations of the pinwheels
 in the orientation map are detrmined as the intersections of the zero contours of the real and imaginary 
@@
 functional properties. And finally the LISSOM abstraction enables very fast simulations, thus allowing us to perform 
-parameter search analysis that would be otherwise computationally prohibitive. However, as we will explain in ?, some simplifications
+parameter search analysis that would be otherwise computationally prohibitive. However, as we will explain in section \ref{sec:SM4}, some simplifications
 made by the LISSOM abstraction, specifically the instantaneous translation of the neuronal input into it's activity, will leave 
-certain questions unanswered. These will be addressed in section ? using more detailed rate model framework.
+certain questions unanswered. These will be addressed in section \ref{sec:SM4} using more detailed rate model framework.
 
-\section{Fast excitatory to inhibitory to excitatory loop enables competition in networks with short-range inhibitory connections.}
+\subsection{Fast excitatory to inhibitory to excitatory loop enables competition in networks with short-range inhibitory connections.} \label{sec:SM1}
 
-? et al. have shown that transmition delays between different types of cortical neurons are not uniform, specifically they found that transmission delays 
-between excitatory neurons are ? ms, from excitatory to inhibitory neurons are ? ms, from inhibitory to excitatory neurons are ?. The sample size for connections
+Ohana et al.
+\cite{Ohana2012} have shown that transmition delays between different types of cortical neurons are not uniform, specifically they found that on average the transmission delays  between excitatory neurons are $\approx$1.4\,ms, from excitatory to inhibitory neurons are $\approx$0.54\,ms, from inhibitory to excitatory neurons are $\approx$0.98\,ms.
+The sample size for connections
 between inhibitory cells in the study was not large enough to be quantitatively reliable. The key observation here is that the 
@@
 
-\noindent Because $\theta + \beta = \alpha$ (assumption 1) and because the inhibitory neurons receive only excitatory connections (assumption ?) we can further simplify the above as:
+\noindent Because $\theta + \beta = \alpha$ (assumption 1) and because the inhibitory neurons receive only excitatory connections (assumption 2) we can further simplify the above as:
 
@@
 supported under these conditions. Conveniently, the lateral interaction in the LISSOM family of models is governed by the same equations as ? and ?. 
-In the following we will use the GCAL model \cite{?}, the latest and most robust addition to the LISSOM family of models to demonstrate that the above 
-specific configuration of effective lateral excitation and inhibition permits development of high-quality orientation maps. To facilitate comparison we will use exactly 
-the same model configuration as in ?, except to three modifications necessitated by the analysis above:
+In the following we will use the GCAL model \cite{Stevens2013}, the latest and most robust addition to the LISSOM family of models to demonstrate that the above specific configuration of effective lateral excitation and inhibition permits development of high-quality orientation maps.
+To facilitate comparison we will use exactly the same model configuration as in Stevens et al.
+\cite{Stevens2013}, except to three modifications necessitated by the analysis above:
 
@@
 In order to find working combination of parameters in point's 2 and 3, and also to show that the model is robust to certain level of changes
-in these two parameters we have performed a parameter search across these two parameters and evaluated the quality of the orientation map (see ?) for each 
-parameter combination. As figure ? shows, under a range of values of both parameters the model develops high-quality orientation maps indistinguishable
+in these two parameters we have performed a parameter search across these two parameters and evaluated the quality of the orientation map (see Materials and Methods) for each 
+parameter combination. As figure \ref{fig:figure1} shows, under a range of values of both parameters the model develops high-quality orientation maps indistinguishable
 from their experimental counterparts, thus concluding our first step towards showing that fast excitatory-to-inhibitory-to-excitatory loop can explain 
@@
 
-Furthermore, note that the GCAL model used in figure ? only explicitly models excitatory neurons and assumes both direct excitatory and inhibitory
+Furthermore, note that the GCAL model used in figure \ref{fig:figure1} only explicitly models excitatory neurons and assumes both direct excitatory and inhibitory
 interactions between them, thus corresponding to equation ?. Even though above we have shown that equation ? is equivalent to equation ? to verify 
 correctness of our analysis we have ran a single simulation of the GCAL model corresponding to the parameter combination with the highest map 
-quality found in figure ?, but with explicitly simulated inhibitory population. In this model we thus do not model direct inhibitory interactions
+quality found in figure \ref{fig:figure1}, but with explicitly simulated inhibitory population. In this model we thus do not model direct inhibitory interactions
 between excitatory neurons, but instead add excitatory to inhibitory and inhibitory to excitatory connections of the same extend as those of 
-excitatory to excitatory (see the assumption \#?). As we have shown above this model should be mathematically equivalent to the simulations 
-figure ?D, whic is confirmed in figure ?E that shows that the simulation results in formation of identical orientation map as in the figure ?D.
+excitatory to excitatory (see the assumption \#5). As we have shown above this model should be mathematically equivalent to the simulations 
+figure \ref{fig:figure1}D, which is confirmed in figure \ref{fig:figure1}E that shows that the simulation results in formation of identical orientation map as in the figure \ref{fig:figure1}D.
 
-\section{Inhibitory to inhibitory connections are consistent with development of functional organization}
+CORRECT THIS END ?????
+
+\subsection{Inhibitory to inhibitory connections are consistent with development of functional organization} \label{sec:SM2}
 
@@
 To this end we will use the exact GCAL model configuration that we have found in previous section to posses the highest
-quality orientation map (see figure ?). We will use GCAL configuration in which we will explicitly model the inhibitory neurons
+quality orientation map (see figure \ref{fig:figure1}). We will use GCAL configuration in which we will explicitly model the inhibitory neurons
 (see figure ?, and equation ?) and consequently also explicitly the excitatory to inhibitory and inhibitory to excitatory connections.
@@
 the overall strength of the projections from inhibitory neurons onto both excitatory and inhibitory populations is scaled by the 
-excitatory to inhibitory projections strength. Therefore in figure ? we have systematically varied the strength of the added inhibitory to inhibitory
+excitatory to inhibitory projections strength. Therefore in figure \ref{fig:figure2} we have systematically varied the strength of the added inhibitory to inhibitory
 projection expressed relatively to the strength of the inhibitory to inhibitory connections (which was set to 1), while also varying the 
@@
 excitatory to inhibitory drive. This shows that inclusion of direct inhibitory to inhibitory interactions does not 
-invalidate the results of section ?.
+invalidate the results of section \ref{sec:SM3}.
 
-\section{Long-range excitation}
+\subsection{Long-range excitation} \label{sec:SM3}
 
-In section ? we have only assumed local connectivity by assuming that both excitatory and inhibitory interaction had the same spatial extent. 
+In section \ref{sec:SM2} we have only assumed local connectivity by assuming that both excitatory and inhibitory interaction had the same spatial extent. 
 However experimental evidence shows that excitatory cells send longer connections compared to inhibitory cells \cite{?}. In this section we will
-explore what happens if we add long-range excitatory connectivity into the model analyzed in section ?. Buz\'as et al. \cite{Buzas2006} have shown that
-the lateral connectivity in layer 2/3 can be best described as superimposition of two gabor connectivity likelyhoods, one short-range not orientation specific
-and one long-range orientation specific. Here we will assume such dual structure, leading us to add second excitatory to excitatory and excitatory to inhibitory
-projection into the model described in section ?, but with 3 times larger space constant (in line with Buz\'as et al. quantitative findings). In figure \ref{fig:figure3} we have 
-kept the overall strength of excitatory to excitatory and excitatory to inhibitory pathways the same, corresponding to the best parametrization found in ?,
-but varied the relative strength between the short range and long range components. As can be seen ...
+explore what happens if we add long-range excitatory connectivity into the model analyzed in section \ref{sec:SM2}. Buz\'as et al. \cite{Buzas2006} have
+shown that the lateral connectivity in layer 2/3 can be best described as superimposition of two gabor connectivity likelyhoods, one short-range not orientation specific and one long-range orientation specific.
+Here we will assume such dual structure, leading us to add second excitatory to excitator and excitatory to inhibitory projection into the model described in section \ref{sec:SM2}, but with 3 times larger space constant (in line with Buz\'as et al.
+quantitative findings).
+In figure \ref{fig:figure3} we have  kept the overall strength of excitatory to excitatory and excitatory to inhibitory pathways the same, corresponding to the best parametrization found in \ref{sec:SM2}, but varied the relative strength between the short range and long range components.
+As can be seen ...
 
@@
 
-\section{Non-equal effective excitatory and inhibitory delay}
+\subsection{Non-equal effective excitatory and inhibitory delay} \label{sec:SM4}
 
 In all model variants examined sofar we have made the key assumption that the delay on the excitatory to excitatory connections is exactly equal 
-to the sum of excitatory to inhibitory and inhibitory to excitatory delays. This assumption is approximately supported by the experimental evidence \cite{?}, 
-but we cannot assume it holds exactly in real biological substrate.  However, we hypothesize, that the small discrepancies between the cumulative delays 
-of the mono-synaptic excitatory and duo?-synaptic inhibitory interactions can be absorbed in the membrane time-constant of the neurons. In this section we will 
-verify this hypothesis by extending the modeling framework used thus far with a finite membrane time-constant (see Materials and Methods ?) and proceed to determine 
-the magnitude of discrepancy in the delays between the excitatory and inhibitory interactions that can be managed by the model without impairing the resulting map orientation 
-map quality. We use model parametrization similar to those determined in previous section ? to lead to high quality orientation maps. For simplicity and computational 
-efficiency we omit the inhibitory to inhibitory and long-range excitatory connections that have already been investigated above. The membrane time-constant of excitatory 
-neurons was set to ?ms while those of inhibitory neurons to ?ms. This faster inhibitory dynamics are necessary to prevent oscillations in the system \cite{?}. 
-Figure ? shows the resulting orientation maps and associated map quality measures of ? models with range of differences between the monosynaptic excitatory and 
-duo?-synaptic inhibitory interactions.
+to the sum of excitatory to inhibitory and inhibitory to excitatory delays.
+This assumption is approximately supported by the experimental evidence \cite{Ohana2012}, but we cannot assume it holds exactly in real biological substrate.
+However, we hypothesize, that the small discrepancies between the delay of the mono-synaptic excitatory and cumulative delay of the bi-synaptic inhibitory interactions can be absorbed in the membrane time-constant of the neurons.
+In this section we will verify this hypothesis by extending the modeling framework used thus far with a finite membrane time-constant (see Materials and Methods) and proceed to determine the magnitude of discrepancy in the delays between the excitatory and inhibitory interactions that can be managed by the model without impairing the resulting map orientation map quality.
+We use model parametrization similar to those determined in previous section \ref{sec:SM1}.
+For simplicity and computational efficiency we omit the inhibitory to inhibitory and long-range excitatory connections that have already been investigated in sections \ref{sec:SM2} and \ref{sec:SM3}.
+The membrane time-constant of excitatory
+neurons was set to \insval{membrane_time_constant_E}ms while those of inhibitory neurons to \insval{membrane_time_constant_I}ms.
+This faster inhibitory dynamics are necessary to prevent oscillations in the system \cite{Kang2003}.
+Figure \ref{fig:figure3} shows the resulting orientation maps and associated map quality measures of ?
+models with range of differences between the delays of monosynaptic excitatory and bi-synaptic inhibitory interactions.
 
@@
 
+As can be seen when the differences between the excitatory delay and cumulative inhibitory delay is small ($\less ?$) the high quality orientation maps develop in the model.
+However, as expected, as the difference between the delays increases the ability of the model to learn topological organized representation of orientation preference diminishes.
+
+\section{Discussion}
+
+TODO
+
 \bibliographystyle{abbrv}

### 2016-11-25 key_value_store.py

@@ -1,0 +1,29 @@
+{
+	# ARCHITECTURE
+	'cortex_size' : ?,
+	'lgn_size' : ?,
+	'visual_field_size' : ?,
+	'cortex_density' : ?,
+	'settle_steps' : ?,
+	'num_gaussian_stim' : ?,
+
+
+	# LGN
+	'retina_to_lgn_strength' : ?,
+	'lateral_lgn_strength' : ?,
+	'lateral_lgn_gain' : ?,
+	'lgn_center_sigma' : ?,
+	'lgn_surround_sigma' : ?,
+	'lgn_lateral_sigma' : ?,
+
+	# Homeostasis
+	'homeostatic_learning_rate'  : ?,
+	'target_activity' : ?,
+	'activity_average_smoothing' : ?,
+	
+	# Hebbian adaptation
+	'learning_rate' : ?,
+
+
+
+}

### 2016-11-25 paper.tex
changes

@@
 
-All previous topologically organized models of cortical competition have either assumed instantaneous synaptic transmission \cite{?,?} or uniform transmission delays across the different intra-cortical connections \cite{?,?}.
+All previous topologically organized models of cortical competition have either assumed instantaneous synaptic transmission \cite{Kang2003,Levy2011,VonderMalsburg1973,CMVC} or uniform transmission delays across the different intra-cortical connections \cite{?}.
 This is also the case of the analysis performed by Muir et al.
-\cite{Muir2014} which questions the ability of intra-cortical connections to explain cortical competition.
+\cite{Muir2014} which questions the ability of the accepted cortical architecture to explain cortical competition.
 However, a recent study by Ohana et al.
 \cite{Ohana2012} have revealed a specific pattern of transmission delays between different neural type combinations.
-Specifically they found that the
-excitatory to excitatory connections are slow while the excitatory to inhibitory connections are fast. In this study we explore the 
+Specifically they found that the excitatory to excitatory connections are slow while the excitatory to inhibitory connections are fast.
+In this study we explore the
 possibility that this specific transmission delay pattern is the missing link that can explain how short range inhibition can lead 
-to effective cortical competition. We employ computational models to show that the fast inhibitory loop allows for the disynptic
-inhibition to be effective and consequently provides the effective Mexican-hat like lateral interactions, thus for the first time 
+to effective cortical competition.
+We employ computational models to show that the fast excitatory-to-inhibitory-to-excitatory pathway allows for the disynptic inhibition to generate the effective Mexican-hat like lateral interactions, thus for the first time
 explaining competition in topologically organized cortical networks with no biologically implausible assumptions. To demonstrate the
@@
 
+TODO?????
+
 \subsection{Orientation map analysis}
@@
 
-In this article we will proceed through multiple gradually more complex models of orientation development to address several key questions,
-eventually demonstrating that the experimentally identified fast inhibitory loop is a satisfactory explanation for how short-range 
-inhibition can support development of cortical functional organization. We will use two different computational abstractions 
-to explore the questions at hand. In the fist part of the study we will use computational model that is derived from the LISSOM 
-family of models \cite{CMVC}. This choice has three advantages. It allows for a very straightforward explanation of why 
+In this article we will proceed through multiple gradually more complex models of orientation development, addressing several 
+of the major issues with modelling of this phenomenam, eventually demonstrating that the experimentally identified fast inhibitory loop is a satisfactory explanation for how short-range inhibition can support development of cortical functional organization.
+We will use two different computational abstractions to explore the questions at hand.
+In the fist part of the study we will use computational model that is derived from the LISSOM family of models \cite{CMVC}.
+This choice has three advantages.
+It allows for a very straightforward explanation of why
 fast inhibitory loop enables short-range inhibition to induce competition. It makes our explanation directly comparable to the 
@@
 functional properties. And finally the LISSOM abstraction enables very fast simulations, thus allowing us to perform 
-parameter search analysis that would be otherwise computationally prohibitive. However, as we will explain in section \ref{sec:SM4}, some simplifications
-made by the LISSOM abstraction, specifically the instantaneous translation of the neuronal input into it's activity, will leave 
+parameter search analysis that would be otherwise computationally prohibitive.
+However, as we will explain further in section \ref{sec:SM4}, some simplifications made by the LISSOM abstraction, specifically the instantaneous translation of the neuronal input into it's activity, will leave
 certain questions unanswered. These will be addressed in section \ref{sec:SM4} using more detailed rate model framework.
@@
 Ohana et al.
-\cite{Ohana2012} have shown that transmition delays between different types of cortical neurons are not uniform, specifically they found that on average the transmission delays  between excitatory neurons are $\approx$1.4\,ms, from excitatory to inhibitory neurons are $\approx$0.54\,ms, from inhibitory to excitatory neurons are $\approx$0.98\,ms.
+\cite{Ohana2012} have shown that transmition delays between different types of cortical neurons are not uniform, specifically they found that on average the transmission delays  between excitatory neurons are $\approx$1.4\,ms, from excitatory to inhibitory neurons are $\approx$0.5\,ms, and from inhibitory to excitatory neurons are $\approx$0.98\,ms.
 The sample size for connections
@@
 of the excitatory to excitatory connections. Drawing from this, the core insight of this study is that under such pattern of delays
-the effective inhibitory interactions are well approximated by the convolution of the excitatory and inhibitory connection kernels, thus 
-giving the effective inhibition longer range, consequently fulfilling the essential requirement for cortical competition to occur.
-
-For didactic purposes, let us first explore one example set of conditions under which the above statement holds exactly:
+the effective inhibitory interactions are well approximated by the convolution of the excitatory and inhibitory connection kernels, as we will
+show below. This gives the effective inhibition longer range, consequently fulfilling the essential requirement for cortical competition to occur.
+For didactic purposes, let us first explore one specific set of conditions under which the above statement holds exactly, we will explore the 
+situation when these conditions are relaxed in subsequent sections:
 
@@
 
-\item The sum of the excitatory to inhibitory , and inhibitory to excitatory delays is exactly the same as the excitatory to excitatory delay
+\item The sum of the excitatory to inhibitory and inhibitory to excitatory delays is exactly the same as the excitatory to excitatory delay
 \item There is no inhibitory to inhibitory interaction
@@
 \item The connection kernels of all neurons are Gaussian kernels with spatial constant $\sigma_{e}$ for excitatory neurons and $\sigma_{i}$ for inhibitory neurons.
-\item We will disregard long-range (excitatory) connectivity, and assume that both local excitatory and local inhibitory connections have the same extent , thus $\sigma_{e} = \sigma_{i}$ 
+\item We will assume only local connectivity (disregarding long-range excitatory connections), and assume that both excitatory and inhibitory have the same extent, thus $\sigma_{e} = \sigma_{i}$ 
 \end{enumerate}
@@
 
-This shows that under these specific assumptions the effective inhibitory interactions between excitatory neurons are $\sqrt{2}$ longer than excitatory and thus
-follow the mexican-hat like profile required for for lateral cortical competition underlying functional map development. Even though this essential condition of 
-cortical competition is fulfilled in this model configuration, the requirement of local inhibition puts important constraints on the extent of the effective lateral
-interactions (relative to lateral excitation). It thus remains unclear wether development of high-quality orientation maps as observed experimentally is 
-supported under these conditions. Conveniently, the lateral interaction in the LISSOM family of models is governed by the same equations as ? and ?. 
+This shows that under these specific assumptions the effective inhibitory interactions between excitatory neurons are $\sqrt{2}$ longer than excitatory and thus follow the mexican-hat like profile required for for lateral cortical competition underlying functional map development.
+Even though this essential condition of cortical competition is fulfilled in this model configuration, important constraints on the extent of the effective lateral
+interactions (relative to lateral excitation) remain.
+It is thus still unclear wether development of high-quality orientation maps as observed experimentally is supported under these conditions.
+Conveniently, the lateral interaction in the LISSOM family of models is governed by the same equations as ?
+and ?.
 In the following we will use the GCAL model \cite{Stevens2013}, the latest and most robust addition to the LISSOM family of models to demonstrate that the above specific configuration of effective lateral excitation and inhibition permits development of high-quality orientation maps.
 To facilitate comparison we will use exactly the same model configuration as in Stevens et al.
-\cite{Stevens2013}, except to three modifications necessitated by the analysis above:
+\cite{Stevens2013} (see Materials and Methods), except to three modifications necessitated by the analysis above:
 
@@
 \item The change in point ? will results in change in the balance of overall excitation and inhibition in the model which is critical 
-to correction functional development in the model. We will thus modify the strength of the lateral inhibition to compensate for changes in ?.
+for successfull functional development in the model. We will thus modify the strength of the lateral inhibition to compensate for changes in ?.
 
@@
 correctness of our analysis we have ran a single simulation of the GCAL model corresponding to the parameter combination with the highest map 
-quality found in figure \ref{fig:figure1}, but with explicitly simulated inhibitory population. In this model we thus do not model direct inhibitory interactions
-between excitatory neurons, but instead add excitatory to inhibitory and inhibitory to excitatory connections of the same extend as those of 
-excitatory to excitatory (see the assumption \#5). As we have shown above this model should be mathematically equivalent to the simulations 
-figure \ref{fig:figure1}D, which is confirmed in figure \ref{fig:figure1}E that shows that the simulation results in formation of identical orientation map as in the figure \ref{fig:figure1}D.
-
-CORRECT THIS END ?????
+quality found in figure \ref{fig:figure1}, but with explicitly simulated inhibitory population (figure \ref{fig:figure1}E).
+In this model we thus do not model direct inhibitory interactions between excitatory neurons, but instead add excitatory to inhibitory and inhibitory to excitatory connections of the same extend as those of excitatory to excitatory pathway (see the assumption \#5).
+As we have shown above this model should be mathematically equivalent to the simulations figure \ref{fig:figure1}D.
+Note, however, that the GCAL simulations represent a discrete approximation in both time and space of equations ?
+and we thus expect small numerical discrepancies.
+Indeed, orientation maps shown in figure \ref{fig:figure1}E  are nearly identical to those in the figure \ref{fig:figure1}D, with only minor numerical differences, confirming the validity of our approach.
 
@@
 effective mexican-hat like interaction and thus support development of orientation maps. However not all assumptions we 
-made were in line with experimental evidence. In this section we will show that one of these assumptions is not necessary, 
+made were in line with experimental evidence. In this section we will numerically show that one of these assumptions is not necessary, 
 specifically that addition of inhibitory connections does not invalidate results of the previous section.
@@
 By explicitly modelling the inhibitory neurons in this model we have replaced a single parameter gowerning the strength of inhibitory
-later interactions in the model from previous section with three new parameters that set the strength of excitatory to inhibitory,
-inhibitory to excitatory and inhibitory to inhibitory connections (all other parameters remaind the same as the 
+later interactions in the model from previous section with three new parameters that set: (1) strength of excitatory to inhibitory,
+(2) inhibitory to excitatory and (3) inhibitory to inhibitory connections (all other parameters remaind the same as the 
 in the best parametrization found in previous section). Note that in principle there is redundancy in these parameters as

### 2016-11-25 key_value_store.py

@@ -25,5 +25,10 @@
 	'learning_rate' : ?,
 
 
+	# time constant Model
+	'membrane_time_constant_E' : ?,
+	'membrane_time_constant_I' : ?,
+
+
 
 }

### 2016-11-25 CODE/generate_figure_material.py
changes

@@ -6,23 +6,26 @@ import sys
 from pinwheel_analysis import *
 import pylab
 from mpl_toolkits.axes_grid1 import make_axes_locatable
-
 sys.path.append('/home/jan/projects/topographica/')
 import topo
 from topo.command.analysis import save_plotgroup
 from topo.command import load_snapshot
 
-f = open('/home/jan/Doc/Papers/fast_inh_paper/DATA/GCAL_EI_II_LONGE/a-p_exc_strength=9_-p_exc_inh_strength=10.3_-p_exc_short_long_ratio=0.5_-p_cortex_exc_target_activity=0.003/results.pickle')
-d = pickle.load(f);f.close()
-#rho = pinwheel_analysis(d['orprefmap'][5:-6,5:-6])['metadata']['rho']
-#print "RHO:", rho
-#metric = gamma_metric(rho,k=7.0)
-#print "METRIC:",metric
-vis(d['orprefmap'][5:-6,5:-6])
-    
-#pylab.figure();pylab.imshow(d['orprefmap'][5:-6,5:-6],interpolation='none',cmap='hsv')
-#pylab.show()
-#0/0
+if False:
+    f = open('/home/jan/Doc/Papers/fast_inh_paper/DATA/GCAL_EI_II/exc_inh_strength=3.4_inh_inh_strength=0.3/results.pickle')
+    d = pickle.load(f);f.close()
+    #mmap = d['orprefmap'][25:-24,25:-24]
+    mmap = d['orprefmap'][2:-3,2:-3]
+    rho = pinwheel_analysis(mmap)['metadata']['rho']
+    print "RHO:", rho
+    metric = gamma_metric(rho,k=5.0)
+    print "METRIC:",metric
+    vis(mmap)
+    print "MEAN:", numpy.mean(mmap)
+    pylab.figure();pylab.imshow(mmap,interpolation='none',cmap='hsv')
+    pylab.figure();pylab.imshow(abs(fftshift(fft2(mmap-0.5))),interpolation='none',cmap='gray')
+    pylab.show()
+    0/0
 
 
 def figure2():
@@ -33,7 +36,7 @@ def figure2():
     Y = []
     qual = []
     
-    if False:
+    if True:
         for a in os.listdir(dirr):
             b = os.path.join(dirr,a);
 
@@ -48,9 +51,13 @@ def figure2():
             
             X.append(d['lat_strength_ratio'])
             Y.append(d['exc_inh_ratio'])
[…]

### 2016-11-25 CODE/pinwheel_analysis.py

@@ -202,6 +202,46 @@ def identify_pinwheels(re_contours,  im_contours, intersections):
 ## Hypercolumn distance estimation #
 #==================================#
 
+def azimuthalAverage(image, center=None):
+    """
+    Calculate the azimuthally averaged radial profile.
+
+    image - The 2D image
+    center - The [x,y] pixel coordinates used as the center. The default is 
+             None, which then uses the center of the image (including 
+             fracitonal pixels).
+    
+    """
+    import numpy as np
+    # Calculate the indices from the image
+    y, x = np.indices(image.shape)
+
+    if not center:
+        center = np.array([(x.max()-x.min())/2.0, (x.max()-x.min())/2.0])
+
+    r = np.hypot(x - center[0], y - center[1])
+
+    # Get sorted radii
+    ind = np.argsort(r.flat)
+    r_sorted = r.flat[ind]
+    i_sorted = image.flat[ind]
+
+    # Get the integer part of the radii (bin size = 1)
+    r_int = r_sorted.astype(int)
+
+    # Find all pixels that fall within each radial bin.
+    deltar = r_int[1:] - r_int[:-1]  # Assumes all radii represented
+    rind = np.where(deltar)[0]       # location of changed radius
+    nr = rind[1:] - rind[:-1]        # number of radius bin
+    
+    # Cumulative sum to figure out sums for each radius bin
+    csim = np.cumsum(i_sorted, dtype=float)
+    tbin = csim[rind[1:]] - csim[rind[:-1]]
+
+    radial_prof = tbin / nr
+
+    return radial_prof
+
 # For featuremapper.analysis
 
 # - Crop to smallest dimension, assume FFT ring in center
@@ -267,38 +307,56 @@ def hypercolumn_distance(preference, init_fit=[0.35,3,1.3,0.15,-0.003,0]):
     the analysed values in a dictionary.
     """
     fft_spectrum = power_spectrum(preference)
-    amplitudes = wavenumber_spectrum(fft_spectrum)
-    ks = np.array(range(len(amplitudes)))
-    
+    amplitudes = np.array(azimuthalAverage(1-fft_spectrum))
[…]

### 2016-11-25 build_figures.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
 
-for x in ./SVG/Figure1/figure1.svg  ./SVG/Figure2/figure2.svg 
+for x in ./SVG/Figure1/figure1.svg  ./SVG/Figure2/figure2.svg ./SVG/Figure3/figure3.svg  ./SVG/Figure4/figure4.svg 
   do
   dir=`dirname $x`
   base=`basename $x .svg`

### 2016-12-13 paper.tex
text changes and new figures

@@
 
+\newcommand{\etal}{\textit{et al.}}
+
 \title{Fast inhibitory loop explains functional map development with short-range inhibition.}
-\author{J\'an Antol\'ik, Andrew P. Davison}
+\author{J\'an Antol\'ik}
 
@@
  \maketitle
-\loadkeyvaluestore{/home/jan/Doc/Papers/lsv1m-paper/key_value_store.py}
+\loadkeyvaluestore{/home/jan/Doc/Papers/fast_inh_paper/key_value_store.py}
 
@@
 As pointed out by Muir et al.
-in their recent study \cite{Muir2014}, to solve this apparent conflict, previous topologically organized models cortical competition have either relied on the anatomically unsupported Mexican-hat profile of lateral interactions \cite{VonderMalsburg1973,CMVC}, or relied on other biologically unrealistic properties, such as selective targeting of inhibitory neurons by long range excitatory connections \cite{?} or instanteneous synaptic transmission coupled with ommission of recurrent inhibition \cite{Kang2003,Levy2011}.
-Furthermore, model analysis by Muir et al.
-\cite{Muir2014} showed that the presence of competition across the cortical surface is predicted well by the anatomy of direct excitatory and inhibitory coupling and that multisynaptic network effects are negligible.
-In conclusion, currently no satisfactory explanation of how topological functional organization develops in cortical networks that is consistent with the present anatomical findings exists.
+in their recent study \cite{Muir2014}, to solve this apparent conflict, previous topologically organized models cortical competition have either relied on the anatomically unsupported Mexican-hat profile of lateral interactions \cite{VonderMalsburg1973,CMVC}, or relied on other biologically unrealistic properties, such as selective targeting of inhibitory neurons by long range excitatory connections \cite{law:phd09,Rutishauser2012} or instantaneous synaptic transmission coupled with omission of recurrent inhibition \cite{Kang2003,Levy2011}.
+Model analysis by Muir et al.
+\cite{Muir2014} showed that the presence of competition across the cortical surface is predicted well by the anatomy of direct excitatory and inhibitory coupling and that multi-synaptic network effects are negligible.
+In conclusion, currently no satisfactory explanation of how topological functional organization dev?lops in cortical networks that is consistent with the present anatomical findings exists.
 
-All previous topologically organized models of cortical competition have either assumed instantaneous synaptic transmission \cite{Kang2003,Levy2011,VonderMalsburg1973,CMVC} or uniform transmission delays across the different intra-cortical connections \cite{?}.
+All previous topologically organized models of cortical competition have either assumed instantaneous synaptic transmission \cite{VonderMalsburg1973,Kang2003,Levy2011,Grabska-Barwinska2008}} or uniform transmission delays across the different intra-cortical connections \cite{CMVC,Antolik2011}.
 This is also the case of the analysis performed by Muir et al.
@@
 Here we describe the two models and their variants used in this study and finish with description of a measure
-for assessing the extent to which a model orientation maps resamble their biological counterparts. Since this study heavily leans on methodology 
+for assessing the extent to which a model orientation maps resemble their biological counterparts. Since this study heavily leans on methodology 
 developed in our previous studies, we offer here only shortened description and refer readers to the original articles for details.
+
+\begin{figure}[htpb!] 
+\centering
+\includegraphics[width=16cm]{./SVG/FigureModelArchitecture/drawing.png}
+\caption{General model architecture. }
+\label{fig:model_architecture}
+\end{figure}
 
@@
 
-In sections ? through ? we use GCAL model \cite{Stevens2013}  which is the most advanced variant of LISSOM 
+In sections \ref{sec:SM1} through \ref{sec:SM3} we use GCAL model \cite{Stevens2013}  which is the most advanced variant of LISSOM 
 (Laterally Interconnected Synergetically Self-Organizing Map) algorithm introduced by Miikkulainen et al. \cite{CMVC}, 
@@
 
-As a simplification, GCAL ignores the detailed temporal properties of the subcortical neural responses and of signal propagation along the various types of connections.
+As a simplification, GCAL ignores the detailed temporal properties of the sub-cortical neural responses and of signal propagation along the various types of connections.
 Instead, the model ON/OFF units have a constant, sustained output, and all connections have a constant delay, independent of the physical length of that connection.
@@
 Any modifications to these base parameters
-in the GCAL model variants examined in sections \ref{sec:SM2,sec:SM3} are then reported in the respective sections.
+in the GCAL model variants examined in sections \ref{sec:SM2} and \ref{sec:SM3} are then reported in the respective sections.
 
-Once all \insval{settle_steps} settling steps are complete, the settled cortica activation pattern is deemed to be the response of cortical sheets to 
+Once all \insval{settle_steps} settling steps are complete, the settled cortical activation pattern is deemed to be the response of cortical sheets to 
 the presented pattern.
@@
 
-where $\xi=\insva{homeostatic_learning_rate}$ is the time constant of the threshold adaptation,
+where $\xi=\insval{homeostatic_learning_rate}$ is the time constant of the threshold adaptation,
 $\mu=\insval{target_activity}$ is a constant defining the target average activity, and $\tilde{\Psi}$ is 
@@
 \noindent where $\Psi(t)$ is the output of the unit at time t and $\chi=\insval{activity_average_smoothing}$ is time constant 
-controling the decay of the influence of the past activities. The effect of this scaling mechanism is to bring the average activity of each 
+controlling the decay of the influence of the past activities. The effect of this scaling mechanism is to bring the average activity of each 
 cortical unit closer to the specified target.
@@
 
-In the model, as images are presented to the photoreceptors, cortical afferent connection weights $\omega_{i,j,A} from the ON/OFF sheets are adjusted once per iterations (after cortical settling is completed) using a simple Hebbian learning rule.
+In the model, as images are presented to the photoreceptors, cortical afferent connection weights $\omega_{i,j,A}$ from the ON/OFF sheets are adjusted once per iterations (after cortical settling is completed) using a simple Hebbian learning rule.
 This rule results in connections that reflect correlations between the presynaptic ON/OFF unit activities and the postsynaptic cortical response.
@@
 
+\begin{figure}[htpb!] 
+\centering
+\includegraphics[width=16cm]{./SVG/FigureModelVariants/drawing.png}
+\caption{The four considered model variants. }
+\label{fig:model_architecture}
+\end{figure}
+
 \subsection{Rate model}
-
-TODO?????
 
@@
 
-Model orientation maps are calculated based on the vector average method \cite{CMVC}. We first determine the prefered frequency of neurons 
+Model orientation maps are calculated based on the vector average method \cite{CMVC}. We first determine the preferred frequency of neurons 
 across the map. Due to the simplified stereotypical stimulus used in this study (elongated Gaussian inputs) the spatial frequency 
-preference of all neurons lies in a very narrow band, and we thus use the mean prefered spatial frequency across all cortical neurons as the
-value for the spatial frequency parameteter across all subsequent analysis.
+preference of all neurons lies in a very narrow band, and we thus use the mean preferred spatial frequency across all cortical neurons as the
+value for the spatial frequency parameter across all subsequent analysis.
 Next, sine grating inputs that cover the full range of remaining parameter values (combinations of all orientations and phases) are presented, and for each orientation, the peak response of the neuron is recorded.
-The orientation preference is calculated by constructing a vector for each orientation $\theta$ (between 0 and 180\circ), with the peak response as
+The orientation preference is calculated by constructing a vector for each orientation $\theta$ (between 0 and 180$\circ$), with the peak response as
 the length and $\theta$ as its orientation.
@@
 in biological orientation maps scales linearly with hypercolumn size across many different species \cite{Kaschube2010}.
-Specifically the pinwheel density per hypercolumn area (\Lambda^2) converges to \pi, when averaged across sufficently 
+Specifically the pinwheel density per hypercolumn area ($\Lambda^2$) converges to $\pi$, when averaged across sufficiently 
 large cortical surface. For a detailed description of the procedure for calculating this metric we refer reader to 
 our previous work \cite{Stevens2013}, but briefly, its calculation involves three steps. First, the locations of the pinwheels
-in the orientation map are detrmined as the intersections of the zero contours of the real and imaginary 
+in the orientation map are determined as the intersections of the zero contours of the real and imaginary 
 components in the polar representation of the maps, thus yielding the total pinwheel count in the map. Second the 
-hypercolumn size is determined as the peak in the isotropic ring-like fourier transform of the orientation maps.
-Third, using these two number we can derive the pinwheel density, but to transform it to a usefull metric between 
-unity (high-quality map) and zero (low quality map) we pass it through normalized gama distribution. In order to 
-avoid any edge effects contaminating this measure, only the central ?x? area of the orientation maps is analyzed for
-all models presented here. We have shown that this matric reliably distinguishes low and high quality maps \cite{Stevens2013}
+hypercolumn size is determined as the peak in the isotropic ring-like Fourier transform of the orientation maps.
+Third, using these two number we can derive the pinwheel density, but to transform it to a useful metric between 
+unity (high-quality map) and zero (low quality map) we pass it through normalized Gama distribution. 
+We have shown that this metric reliably distinguishes low and high quality maps \cite{Stevens2013}
 and is a valid measure for assessing how well the model orientation maps match animal data.
@@
 In this article we will proceed through multiple gradually more complex models of orientation development, addressing several 
-of the major issues with modelling of this phenomenam, eventually demonstrating that the experimentally identified fast inhibitory loop is a satisfactory explanation for how short-range inhibition can support development of cortical functional organization.
+of the major issues with modeling of this phenomena, eventually demonstrating that the experimentally identified fast inhibitory loop is a satisfactory explanation for how short-range inhibition can support development of cortical functional organization.
 We will use two different computational abstractions to explore the questions at hand.
@@
 Ohana et al.
-\cite{Ohana2012} have shown that transmition delays between different types of cortical neurons are not uniform, specifically they found that on average the transmission delays  between excitatory neurons are $\approx$1.4\,ms, from excitatory to inhibitory neurons are $\approx$0.5\,ms, and from inhibitory to excitatory neurons are $\approx$0.98\,ms.
+\cite{Ohana2012} have shown that transmission delays between different types of cortical neurons are not uniform, specifically they found that on average the transmission delays  between excitatory neurons are $\approx$1.4\,ms, from excitatory to inhibitory neurons are $\approx$0.5\,ms, and from inhibitory to excitatory neurons are $\approx$0.98\,ms.
 The sample size for connections
@@
 \begin{split}
-& C_{m} \frac{dV}{dt} = \frac{(V_{rest} - V)}{R_{m}} + \frac{}{{R_{m}}}  g_{ex}(E_{ex} - V) + g_{inh}(E_{inh} - V)  \\
-
+& R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\theta) - N_{\sigma_{i}}(\lVert y-x \rVert)R_{i}(y,t-\theta) + I_{aff}(x,t)]^+ \\
 & R_{i}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\theta)]^+
@@
 
-\noindent where $R_{p}(x,t)$ is the response of neuron of type $p$ located at position $x$ at time $t$, $I_{aff}(x,t)$ is the afferent input to neuron $x$ at time $t$, $N_{\sigma}$(d) 
+\noindent where $R_{p}(x,t)$ is the response of neuron of type $p$ located at position $x$ at time $t$, $I_{aff}(x,t)$ is the afferent input to neuron $x$ at time $t$, $N_{\sigma}$
 is normal distribution of variance $\sigma$, corresponding to the lateral connection kernel. When we expand for $R_{i}$ we obtain:
@@
 \begin{equation}
+\label{eqn:last}
 R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\alpha) - \sum_{z}N_{\sqrt{2}\sigma_{e}}(\lVert z-x \rVert) R_{e}(z,t-\alpha)]^+
@@
 
-This shows that under these specific assumptions the effective inhibitory interactions between excitatory neurons are $\sqrt{2}$ longer than excitatory and thus follow the mexican-hat like profile required for for lateral cortical competition underlying functional map development.
+This shows that under these specific assumptions the effective inhibitory interactions between excitatory neurons are $\sqrt{2}$ longer than excitatory and thus follow the Mexican-hat like profile required for for lateral cortical competition underlying functional map development.
 Even though this essential condition of cortical competition is fulfilled in this model configuration, important constraints on the extent of the effective lateral
 interactions (relative to lateral excitation) remain.
-It is thus still unclear wether development of high-quality orientation maps as observed experimentally is supported under these conditions.
-Conveniently, the lateral interaction in the LISSOM family of models is governed by the same equations as ?
-and ?.
-In the following we will use the GCAL model \cite{Stevens2013}, the latest and most robust addition to the LISSOM family of models to demonstrate that the above specific configuration of effective lateral excitation and inhibition permits development of high-quality orientation maps.
-To facilitate comparison we will use exactly the same model configuration as in Stevens et al.
-\cite{Stevens2013} (see Materials and Methods), except to three modifications necessitated by the analysis above:
+It is thus still unclear whether development of high-quality orientation maps as observed experimentally is supported under these conditions.
+Conveniently, the lateral interaction in the LISSOM family of models is governed by the same equations as \ref{eqn:first}.
+In the following we will use the GCAL model \cite{Stevens2013}, the latest and most robust addition to the LISSOM family of models
+to demonstrate that the above specific configuration of effective lateral excitation and inhibition permits development of high-quality orientation maps.
+To facilitate comparison we will use exactly the same model configuration
+as in Stevens et al. \cite{Stevens2013} (see Materials and Methods), except to three modifications necessitated by the analysis above:
 
@@
 \item We will change the spread of lateral inhibition to be $\sqrt{2}$ longer than lateral excitatory spread, such that the resulting lateral
-interactions conform with equation ?.
+interactions conform with equation \ref{eqn:last}.
 
-\item The change in point ? will results in change in the balance of overall excitation and inhibition in the model which is critical 
-for successfull functional development in the model. We will thus modify the strength of the lateral inhibition to compensate for changes in ?.
+\item The change in point 1 will results in change in the balance of overall excitation and inhibition in the model which is critical 
+for successful functional development in the model. We will thus modify the strength of the lateral inhibition to compensate for changes due to the modification 1.
 
-\item Analogously to ?, the changes in ? change the overall balance between feedforward and lateral contributions to model cortical neuron's activity, 
+\item Analogously to 2, the changes in 1 change the overall balance between feedforward and lateral contributions to model cortical neuron's activity, 
 which we will compensate by changing the overall strength of the lateral interactions.
@@
 indicated by the red marks.
-From left to right, the orientation map, fast-fourier transform of the orientation map and afferent connection fields from the ON LGN model sheet for 25 example model V1 neurons.
-(B-C) Two examples of sub-optimal orientation maps.
-(D) Model configuration with the highest quality map found in this parameter search.
-(E) The same model configuration but in this case ran with explicit
-simulation of inhibitory neurons and corresponding connections. E and D are identical confirming correctnes of our analysis.}
+From left to right, the orientation map, fast-Fourier transform of the orientation map and afferent connection fields from the ON LGN model sheet for 25 example model V1 neurons.
+(B-C) Two examples of sub-optimal orientation maps. (D) Model configuration with the highest quality map found in this parameter search. 
+(E) The same model configuration but in this case ran with explicit simulation of inhibitory neurons and corresponding connections. E and D are nearly identical confirming correctness of our analysis.
+}
 \label{fig:figure1}
@@
 Furthermore, note that the GCAL model used in figure \ref{fig:figure1} only explicitly models excitatory neurons and assumes both direct excitatory and inhibitory
-interactions between them, thus corresponding to equation ?. Even though above we have shown that equation ? is equivalent to equation ? to verify 
+interactions between them, thus corresponding to equation \ref{eqn:last}. Even though above we have shown that equation \ref{eqn:last} is equivalent to equation \ref{eqn:first} to verify 
 correctness of our analysis we have ran a single simulation of the GCAL model corresponding to the parameter combination with the highest map 
-quality found in figure \ref{fig:figure1}, but with explicitly simulated inhibitory population (figure \ref{fig:figure1}E).
-In this model we thus do not model direct inhibitory interactions between excitatory neurons, but instead add excitatory to inhibitory and inhibitory to excitatory connections of the same extend as those of excitatory to excitatory pathway (see the assumption \#5).
-As we have shown above this model should be mathematically equivalent to the simulations figure \ref{fig:figure1}D.
-Note, however, that the GCAL simulations represent a discrete approximation in both time and space of equations ?
-and we thus expect small numerical discrepancies.
-Indeed, orientation maps shown in figure \ref{fig:figure1}E  are nearly identical to those in the figure \ref{fig:figure1}D, with only minor numerical differences, confirming the validity of our approach.
+quality found in figure \ref{fig:figure1}, but with explicitly simulated inhibitory population (figure \ref{fig:figure1}E). In this model we thus do not model direct inhibitory 
+interactions between excitatory neurons, but instead add excitatory to inhibitory and inhibitory to excitatory connections of the same extend as those of excitatory to excitatory pathway (see the assumption \#5).
+As we have shown above this model should be mathematically equivalent to the simulations figure \ref{fig:figure1}D. Note, however, that the GCAL simulations represent a discrete approximation 
+in both time and space of equations \ref{eqn:first},\ref{eqn:last} and we thus expect small numerical discrepancies.
+Indeed, orientation maps shown in figure \ref{fig:figure1}E  are nearly identical to those
+in the figure \ref{fig:figure1}D, with only barely perceptible numerical differences, confirming the validity of our approach.
 
@@
 In the previous section we have shown that under a set of specific assumptions short-range inhibition can induce 
-effective mexican-hat like interaction and thus support development of orientation maps. However not all assumptions we 
+effective Mexican-hat like interaction and thus support development of orientation maps. However not all assumptions we 
 made were in line with experimental evidence. In this section we will numerically show that one of these assumptions is not necessary, 
@@
 quality orientation map (see figure \ref{fig:figure1}). We will use GCAL configuration in which we will explicitly model the inhibitory neurons
-(see figure ?, and equation ?) and consequently also explicitly the excitatory to inhibitory and inhibitory to excitatory connections.
+(see figure ?, and equation \ref{eqn:first}) and consequently also explicitly the excitatory to inhibitory and inhibitory to excitatory connections.
 Furthermore, we will add inhibitory to inhibitory connections to the model with the same extent as that of inhibitory to excitatory
-connections ($\sigma_{i} = \sigma_{e}$ in eq ?).
+connections ($\sigma_{i} = \sigma_{e}$ in eq \ref{eqn:first}).
 
@@
 indicated by the red marks.
-From left to right, the orientation map, fast-fourier transform of the orientation map and afferent connection fields from the ON LGN model sheet for 25 example model V1 neurons.
+From left to right, the orientation map, fast-Fourier transform of the orientation map and afferent connection fields from the ON LGN model sheet for 25 example model V1 neurons.
 }
@@
 
-By explicitly modelling the inhibitory neurons in this model we have replaced a single parameter gowerning the strength of inhibitory
+By explicitly modeling the inhibitory neurons in this model we have replaced a single parameter governing the strength of inhibitory
 later interactions in the model from previous section with three new parameters that set: (1) strength of excitatory to inhibitory,
-(2) inhibitory to excitatory and (3) inhibitory to inhibitory connections (all other parameters remaind the same as the 
+(2) inhibitory to excitatory and (3) inhibitory to inhibitory connections (all other parameters remained the same as the 
 in the best parametrization found in previous section). Note that in principle there is redundancy in these parameters as
@@
 In section \ref{sec:SM2} we have only assumed local connectivity by assuming that both excitatory and inhibitory interaction had the same spatial extent. 
-However experimental evidence shows that excitatory cells send longer connections compared to inhibitory cells \cite{?}. In this section we will
+However experimental evidence shows that excitatory cells send longer connections compared to inhibitory cells \cite{Buzas2006,Budd2001}. In this section we will
 explore what happens if we add long-range excitatory connectivity into the model analyzed in section \ref{sec:SM2}. Buz\'as et al. \cite{Buzas2006} have
-shown that the lateral connectivity in layer 2/3 can be best described as superimposition of two gabor connectivity likelyhoods, one short-range not orientation specific and one long-range orientation specific.
-Here we will assume such dual structure, leading us to add second excitatory to excitator and excitatory to inhibitory projection into the model described in section \ref{sec:SM2}, but with 3 times larger space constant (in line with Buz\'as et al.
-quantitative findings).
-In figure \ref{fig:figure3} we have  kept the overall strength of excitatory to excitatory and excitatory to inhibitory pathways the same, corresponding to the best parametrization found in \ref{sec:SM2}, but varied the relative strength between the short range and long range components.
-As can be seen ...
+shown that the lateral connectivity in layer 2/3 can be best described as superimposition of two gabor connectivity likelyhoods, one short-range not orientation 
+specific and one long-range orientation specific. Here we will assume such dual structure, leading us to add second excitatory to excitatory and excitatory to 
+inhibitory projection into the model described in section \ref{sec:SM2}, but with 3 times larger space constant (in line with Buz\'as \etal \, \cite{Buzas2006} quantitative findings). 
+In figure \ref{fig:figure3} we have  kept the overall strength of excitatory to excitatory and excitatory to inhibitory pathways the same, corresponding to the best 
+parametrization found in \ref{sec:SM2}, but varied the relative strength between the short range and long range components (y axis of figure \ref{fig:figure3}). 
+Adding the long-range excitation changes the spatial configurations over which excitatory interactions win over inhibitory ones, as well as the 
+overall magnitude of the resulting net local excitation. Consequently, the proportion between the magnitude of the net local excitation due to the lateral interactions and the excitation
+due to afferent inputs is changed, which is a crucial parameter for map development.
+To compensate for these changes, we also systematically explore the ratio of the overall magnitudes of the lateral and afferent
+interactions (x axis of figure \ref{fig:figure3}).
+To make the parameter search computationally feasible, we perform the parameter search only in the region of parameters that allow for sufficiently strong
+long-range excitatory projections.
+As can be seen in figure \ref{fig:figure3}D and E, under appropriately strong lateral interactions, substantial long-range excitatory connection support development of high quality orientation maps, demonstrating that the proposed model is consistent with the experimentally identified long-range excitatory connectivity.
 
@@
 \centering
-\includegraphics[width=16cm]{./SVG/Figure2/figure3.png}
-\caption{}
+\includegraphics[width=16cm]{./SVG/Figure3/figure3.png}
+\caption{Orientation map development with short range inhibition, fast excitatory-to-inhibitory-to-excitatory loop and long-range excitation. (A) Map quality (see Materials and Methods) at a range of 
+short vs.
+long range excitatory connection strength ratios (y-axis) and range of afferent vs.
+lateral connection strength ratios (x-axis).
+(B-E) Functional organization in 4 example parameter configurations
+indicated by the red marks.
+From left to right, the orientation map, fast-Fourier transform of the orientation map and afferent connection fields from the ON LGN model sheet for 25 example model V1 neurons.}
 \label{fig:figure3}
@@
 
-In all model variants examined sofar we have made the key assumption that the delay on the excitatory to excitatory connections is exactly equal 
+In all model variants examined so far we have made the key assumption that the delay on the excitatory to excitatory connections is exactly equal 
 to the sum of excitatory to inhibitory and inhibitory to excitatory delays.
@@
 For simplicity and computational efficiency we omit the inhibitory to inhibitory and long-range excitatory connections that have already been investigated in sections \ref{sec:SM2} and \ref{sec:SM3}.
-The membrane time-constant of excitatory
-neurons was set to \insval{membrane_time_constant_E}ms while those of inhibitory neurons to \insval{membrane_time_constant_I}ms.
+The membrane time-constant of excitatory neurons was set to \insval{membrane_time_constant_E}ms while those of inhibitory neurons to \insval{membrane_time_constant_I}ms.
 This faster inhibitory dynamics are necessary to prevent oscillations in the system \cite{Kang2003}.
-Figure \ref{fig:figure3} shows the resulting orientation maps and associated map quality measures of ?
-models with range of differences between the delays of monosynaptic excitatory and bi-synaptic inhibitory interactions.
+We set the excitatory to excitatory delay to 1.4 ms and excitatory to inhibitory delay to 0.5 ms based on Ohana \etal\,\cite{Ohana2012}.
+In order to understand how closely does the
+cumulative bi-synaptic inhibition delay has to match that of the direct excitatory to excitatory delay, we will vary the delay on the inhibitory to excitatory projection (note that we could achieve the same by varying excitatory
+to inhibitory delay, and this choice was arbitrary).
+Figure \ref{fig:figure3} shows the resulting orientation maps and associated map quality measures of models with range of differences between the delays of monosynaptic excitatory and bi-synaptic inhibitory interactions, that are in the figure expressed as the sum of the excitatory to inhibitory and inhibitory to excitatory delays.
 
@@
 \includegraphics[width=16cm]{./SVG/Figure4/figure4.png}
-\caption{}
-\label{fig:figure3}
+\caption{Orientation map development in a rate model with synaptic time constant. (A) Map quality (see Materials and Methods) at a range of 
+cumulative inhibitory delays expressed as the difference from the direct excitatory to excitatory delay. (B-F) Functional organization in 5 parameter configurations 
+indicated by the red marks. The orientation map (top), and afferent connection fields from the ON LGN model sheet for 81 example model V1 neurons (bottom). 
+(B-D) Three examples of configuration where good quality orientation maps develop. (E) If the cumulative inhibitory delay is longer by more than approximately 0.8ms in comparison
+to the direct excitatory delay the map quality starts to drop. (F) In the configuration corresponding to the case where delays on all connections are equal (i.e. the cumulative delay
+of the inhibitory interactions is twice as long as on the direct excitatory ones) orientation maps fail to develop.}
+\label{fig:figure4}
 \end{figure}
 
-As can be seen when the differences between the excitatory delay and cumulative inhibitory delay is small ($\less ?$) the high quality orientation maps develop in the model.
+As can be seen when the differences between the excitatory delay (1.4ms) and cumulative inhibitory delay is small (\textless 0.8 ms) high quality orientation maps develop in the model, confirming that sufficiently
+large discrepancy between the direct excitatory and bi-synaptic inhibitory delays can be accommodated in the model.
 However, as expected, as the difference between the delays increases the ability of the model to learn topological organized representation of orientation preference diminishes.
+Crucially, if the delays across all the projections would be equal, as is typically assumed, the model fails to develop orientation maps in line with the analysis by Muir \etal \, \cite{Muir2014}, thus confirming that the specific delay pattern between neural types identified by Ohana \etal \, \cite{Ohana2012} is key to achieving competitive dynamics in topologically organized neural models.
 
@@
 
-TODO
+In this study we have shown how recent findings on dependence of neural transmission delays on the
+type of pre- and post-synaptic neuron \cite{Ohana2012} can resolve long standing question on how short-range inhibition
+can support cortical competition and consequently the development cortical functional topological organization.
+Under simplifying assumptions, we have analytically shown how dy-synaptic inhibition that is as fast as mono-synaptic 
+excitation can extend the effective range of inhibitory interactions, in contrast to the recent analytical 
+results showing that in the case of equal synaptic delays on all connections the dy-synaptic inhibition has negligable effects \cite{Muir2014}.\
+We have also shown that these findings are applicable to the problem of functional development in primary visual cortex. We have then proceeded 
+to show using computational methods that the proposed models are robust to the addition of other well established features of cortical anatomy, commonly 
+ignored by similar studies, including the long-range excitatory connections and mutual inhibition among inhibitory neurons. Finally, we have shown 
+that the proposed mechanisms are robust to the variations of the exact delay ratio between the mono-synaptic excitation and di-synaptic inhibition. 
+Overall, this study represents an important advance in our understanding how orientation map development can be supported by cortical neural substrate. These
+results generelize to the development of other functional features in the cortex and other cortical competition based mechanisms in general.
+
+The most related past explanation of how cortical competition can arise under short range inhibition is that of Kang \etal\,\cite{Kang2003}, who have shown 
+that under the assumption of faster inibitory time constant (as opposed to excitatory), the effective excitatory and inhibitory interactions will follow the 
+Mexican hat profile and thus support competition along the cortical surface. Indeed, in our final model that explicitly considers membrane time-constant presented in section \ref{sec:SM4}, 
+we assume that inhibitory neurons have faster membrane time constants than excitatory ones, as otherwise we observe oscilatory behavior in line with the analytical findings of Kang \etal\,\cite{Kang2003}.
+Crucially, Kang \etal\,\cite{Kang2003} however assumed instantenous neural transmission, and when this biologically implausible assumption is rectified by addition of tansmission delays that are uniform
+accross the connections between the different pre- and post-synaptic neural types, we find that the competitive dynamics in the neural model break (section \ref{sec:SM4}E) in line with the analytical and
+computational results of Muir \etal\,\cite{Muir2014}.
+However, when we replace the tansmission delays with the neural-type specific  pattern uncovered by Ohana \etal\,\cite{Ohana2012} the competitive dymanics
+in the model are rescued and we observe development of high quality orientation maps (section \ref{sec:SM4}AB), in line with the analytical results under simplifieng conditions (section \ref{sec:SM1}).
+
+The analytical results in this study were obtained only under simplifying assumptions, specificaly instantaneous translation of inputs to membrane potential, equal extent of 
+excitatory and inhibitory connections and lack of inhibitory-to-inhibitory interactions. Even though we have shown computationally that these assumptions are not neccessary 
+for achieving the cortical competition and the consequential orientation map development sought in these study, further analytical work than can circuimvent these simplifications would 
+undoubtly provide deeper understanding of the dynamics of the studied neural system and its dependence on the various parameters. This sentiment is underlied by the parameter 
+explorations presented here, which show that even though the model is robust to changes of the considered parameters, the existence or not of dynamics supporting development of 
+orientation maps forms a complex pattern within the explored parameter spaces. Furthermore, the relative computational complexity of the studied models and the extensive set of parameters
+involved preclude systematic search across the full parameter space, and we have only explored parameters that we empirically found to have biggest impact. Finally, one simplifying assumption 
+that we have not treated in this study is the lack of direct thalamic input onto inhibitory cells. Since inhibitory cells in cortical layer 4 do receive thalamic input \cite{Binzegger2004} the 
+inclusion of external input in the inhibitory population needs to be considered in future.
 

### 2016-12-13 CODE/generate_figure_material.py

@@ -36,7 +36,7 @@ def figure2():
     Y = []
     qual = []
     
-    if True:
+    if False:
         for a in os.listdir(dirr):
             b = os.path.join(dirr,a);
 
@@ -106,7 +106,7 @@ def figure2():
        fig.add_axes(ax)
        ax.imshow(power_spectrum(d['orprefmap'][61:-60,61:-60]),interpolation='none',cmap='gray',aspect='normal')
        pylab.savefig('/home/jan/Doc/Papers/fast_inh_paper/SVG/Figure1/generated_data/fft_map2.png',  pad_inches=0)
-       
+
 
        load_snapshot('/home/jan/Doc/Papers/fast_inh_paper/DATA/GCAL_SHORTRANGE/lat_strength_ratio=2.6_exc_inh_ratio=0.8/snapshot.typ')
        save_plotgroup("Projection",projection=topo.sim["V1"].projections().values()[0],density=1.5,saver_params={'filename_prefix' : '2'})
@@ -129,7 +129,27 @@ def figure2():
      
        load_snapshot('/home/jan/Doc/Papers/fast_inh_paper/DATA/GCAL_SHORTRANGE/lat_strength_ratio=2.8_exc_inh_ratio=0.65/results.pickle')
        save_plotgroup("Projection",projection=topo.sim["V1"].projections().values()[0],density=1.5,saver_params={'filename_prefix' : '1'})
+
+    if True:
+       load_snapshot('/home/jan/projects/topographica/GCAL_EI/a-p_exc_strength=3.3_-p_inh_strength=2.805/snapshot.typ')
+       mmap = topo.sim["V1"].sheet_views["OrientationPreference"].view()[0]
+
+       fig = pylab.figure(dpi=600,facecolor='w',figsize=(5,5))
+       ax = plt.Axes(fig, [0., 0., 1., 1.])
+       ax.set_axis_off()
+       fig.add_axes(ax)
+       ax.imshow(mmap[61:-60,61:-60],interpolation='none',cmap='hsv',aspect='normal')
+       pylab.savefig('/home/jan/Doc/Papers/fast_inh_paper/SVG/Figure1/generated_data/or_map4.png',  pad_inches=0)
        
+       fig = pylab.figure(dpi=600,facecolor='w',figsize=(5,5))
+       ax = plt.Axes(fig, [0., 0., 1., 1.])
+       ax.set_axis_off()
[…]

### 2016-12-13 build_figures.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
 
-for x in ./SVG/Figure1/figure1.svg  ./SVG/Figure2/figure2.svg ./SVG/Figure3/figure3.svg  ./SVG/Figure4/figure4.svg 
+for x in ./SVG/Figure1/figure1.svg  ./SVG/Figure2/figure2.svg ./SVG/Figure3/figure3.svg  ./SVG/Figure4/figure4.svg ./SVG/FigureModelArchitecture/drawing.svg ./SVG/FigureModelVariants/drawing.svg
   do
   dir=`dirname $x`
   base=`basename $x .svg`

### 2016-12-13 key_value_store.py

@@ -1,34 +1,31 @@
 {
 	# ARCHITECTURE
-	'cortex_size' : ?,
-	'lgn_size' : ?,
-	'visual_field_size' : ?,
-	'cortex_density' : ?,
-	'settle_steps' : ?,
-	'num_gaussian_stim' : ?,
+	'cortex_size' : 1.0,
+	'lgn_size' : 1.5,
+	'visual_field_size' : 3.5,
+	'cortex_density' : 96,
+	'settle_steps' : 16,
+	'num_gaussian_stim' : 2,
 
 
 	# LGN
-	'retina_to_lgn_strength' : ?,
-	'lateral_lgn_strength' : ?,
-	'lateral_lgn_gain' : ?,
-	'lgn_center_sigma' : ?,
-	'lgn_surround_sigma' : ?,
-	'lgn_lateral_sigma' : ?,
+	'retina_to_lgn_strength' : 1.5,
+	'lateral_lgn_strength' : 0.6,
+	'lateral_lgn_gain' : 0.11,
+	'lgn_center_sigma' : 0.07385,
+	'lgn_surround_sigma' : 0.2954,
+	'lgn_lateral_sigma' : 0.25,
 
 	# Homeostasis
-	'homeostatic_learning_rate'  : ?,
-	'target_activity' : ?,
-	'activity_average_smoothing' : ?,
+	'homeostatic_learning_rate'  : 0.01,
+	'target_activity' : 0.24,
+	'activity_average_smoothing' : 0.991,
 	
 	# Hebbian adaptation
-	'learning_rate' : ?,
+	'learning_rate' : 0.2,
 
 
 	# time constant Model
-	'membrane_time_constant_E' : ?,
-	'membrane_time_constant_I' : ?,
-
-
-
+	'membrane_time_constant_E' : 2, #ms
+	'membrane_time_constant_I' : 0.5, #ms
 }

### 2016-12-14 paper.tex
figure captions

@@
 \includegraphics[width=16cm]{./SVG/FigureModelArchitecture/drawing.png}
-\caption{General model architecture. }
+\caption{General model architecture. (A) Inhibitory cortical population corresponds to a regular lattice of units in cortical space. (B) As in A but 
+for excitatory population. (C) Thalamic neurons are modeled as simple difference-of-Gaussian filter followed by a threshold-linear trasfer function. 
+(D) The intra-cortical connectivity between the excitatory and inhibitory populations and their transmission delays.
+The intra-cortical connectivity is the only aspect of the general architecture presented here that changes between the diffrent model variants explored in this study and each variant is detail in figure \ref{fig:model_variants}.
+(F) The centers of RFs of the LGN neurons form a regular latice accross the visual space covered by the model.
+This retinotopic mapping of connection fields between layers of the model is also maintained in the thalamo-cortical projection (lines between D and F)}
 \label{fig:model_architecture}
@@
 
-\subsection{GCAL model}
+\subsection{GCAL model}\label{sec:gcal}
 
@@
 previous work \cite{Stevens2013,Antolik2011}. The architecture of the 3 GCAL variants presented in this 
-paper is depicted in figure \ref{?}. The models are implemented in the Topographica simulator, freely available at www.topographica.org. 
-Each of the GCAL models consists of set of sheets of model neural units,
-corresponding to: (1) the photoreceptors, (2) the combined effects of the retinal ganglion cells (RGC) and ON and OFF LGN cells.
+paper is depicted in figure \ref{fig:fig:model_architecture} and figure \ref{model_variants}A-C. The models are implemented in the Topographica simulator, freely available at www.topographica.org. 
+Each of the GCAL models consists of set of sheets of model neural units, corresponding to: (1) the photoreceptors, (2) the combined effects of the retinal ganglion cells (RGC) and ON and OFF LGN cells.
 Furthermore each model has one V1 sheet of units with direct excitatory and inhibitory interactions (section \ref{sec:SM1}), or two
@@
 \includegraphics[width=16cm]{./SVG/FigureModelVariants/drawing.png}
-\caption{The four considered model variants. }
-\label{fig:model_architecture}
+\caption{The four considered intra-cortical connectivty variants. (A-C) Are modelled using firing rate models of neural units with instantenous 
+translation of inputs into membrane-potential (see section \ref{sec:gcal}). 
+(A) Variant 1, assuming only local equaly wide excitatory and inhibitory connectivity, and ignoring direct inhibitory to inhibitory interactions.
+(B) Variant 2, as in A but with added direct inhibitory to inhibitory connections of equal extent as the other excitatory and inhibitory projections.
+(C) Variant 3, as in A but with added long-range excitatory connections represented based on Buz\'as \etal\,\cite{Buzas2006} as a second wider gaussian.
+(D) The same as A but modelled with neural units that take into account membrane time-cosntant (see section \ref{sec:rate}).
+The transmission delay of the inhibitory to excitatory projections was varied in the experiments presneted in section \ref{sec:SM4}.}
+\label{fig:model_variants}
 \end{figure}
 
-\subsection{Rate model}
+\subsection{Rate model} \label{sec:rate}
+
+The architecture of this model is identical to the GCAL models presented above (see figure \ref{fig:fig:model_architecture}) 
+with the exception of the equation ? describing the neural model which is replaced with equation taking into account the 
+membrane time constant:
+
+\begin{equation}
+\Psi_{i}(t)= f(\sum_{p}\gamma_{p}C_{ip}(t))
+\end{equation}
+
+As a consequence the evolution of the network dynamics have to be simulated at higher resolution. We chose the update step to be ?ms, and we 
+let the activity in the model to settle for ?ms, thus resulting in ? settling steps as opposed to the \ref{} of the GCAL model.
+
+figure \ref{fig:fig:model_architecture} and figure \ref{model_variants}A-C
 
@@
 
+In this paper we have decided to investigate cortical competitive mechamnisms through the prism of orientation map development.
+The advantage of this approach is that it allows us not only to show that some form of competition is possible, but also that it is of the form that actually supports implementation of specific cortical computations.
+Given that we show that our model implements effective Mexican hat lateral interactions (section \ref{sec:SM1}) and these have been in the past shown to be sufficient to explain cortical organization of other functional features (i.e.
+retinotopy, ocular dominance, spatial frequency and color) it is very likely that our results will generalzie to these other dimensions of sensory input as well.
+Cortical competition of other forms have been proposed to underly broad variety of other cortical operations, including associative memory, noise suppression, decission making, saliency detection and other forms of attentional computations.
+Even though additional work will be required to determine if the mechanisms proposed here can generelize to these other neural computations, this study offers
+promising framework for anatomically plausible mechinistic explanations of these important aspect of brain function.
+
 \bibliographystyle{abbrv}

### 2017-01-10 paper.tex
changes

@@
-\documentclass[a4paper,10pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{fullpage}
+\documentclass[utf8]{frontiersSCNS}
+\usepackage{url,hyperref,lineno,microtype,subcaption}
+\usepackage[onehalfspacing]{setspace}
 \usepackage{mathtools}
 \usepackage{graphicx}
+\usepackage{gensymb}
 \usepackage{insval}
-\usepackage{natbib}
+
+\linenumbers
 
@@
 
-\title{Fast inhibitory loop explains functional map development with short-range inhibition.}
-\author{J\'an Antol\'ik}
+%\title[Running Title]{Fast inhibitory loop explains functional map development with short-range inhibition.}
+%\author{J\'an Antol\'ik}
+
+%\begin{document}
+%\maketitle
+
+\def\keyFont{\fontsize{8}{11}\helveticabold }
+\def\firstAuthorLast{Antol\'ik} 
+\def\Authors{J\'an Antol\'ik\,$^{1,*}$}
+\def\Address{$^{1}$ Unit\'e de Neurosciences, Information et Complexit\'e,CNRS UPR 3293, Gif-sur-Yvette, France }
+% The Corresponding Author should be marked with an asterisk
+% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
+\def\corrAuthor{J\'an Antol\'ik}
+\def\corrEmail{antolikjan@gmail.com}
 
 \begin{document}
+
+\onecolumn
+\firstpage{1}
+
+\title[Running Title]{Fast inhibitory loop explains functional map development with short-range inhibition.}
+
+\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
+\address{} %This field will be automatically populated
+\correspondance{} %This field will be automatically populated
+\extraAuth{}
+
 \maketitle
-\loadkeyvaluestore{/home/jan/Doc/Papers/fast_inh_paper/key_value_store.py}
+
+\begin{abstract}
+
+\section{}
+
+Competitive interactions are believed to underlie variety of cortical processing, ranging from memory formation, attention and 
+development of cortical functional organization (e.g. development of orientaion maps in primary visual cortex). 
+In the latter case, the competitive interactions happen along the cortical 
+surface, with local populations of neurons reinforcing each other while competing with those displaced more distally. This 
+specific configuration of lateral interactions is however in stark contrast with the known properties of the antatomical substrate, i.e.
+excitatory connections (mediating reinforcment) having longer reach than inhibitory ones (mediating competition). No 
+satisfactory biologically plausible resolution of this conflict between anatomical measures, and assumed cortical function has been proposed.
+Recently a specific pattern of delays between different types of neurons in cat cortex has been discovered, where direct mono-synaptic
+excitation has approximately the same delay, as the combined delays of the disynaptic inhibitory interactions between excitatory neurons 
+(i.e. the sum of delays from excitatory to inhibitory and from inhibitory to excitatory neuron). Here we show that this spefic pattern
+of delays represents a biologically plausible explanation for how short-range inhibition can support competitive interactions in cortex
+that underlie development of orientation maps in primary visual cortex. We demonstrate this statement analytically under simplyfing conditions, 
+and subsequently show using network simulations that development of orientation maps is preserved when long-range excitation, 
+direct inhibitory to inhibitory interactions, and moderate inequality in the delays between excitatory and inhibitory pathways is 
+added.
+
+\tiny
+\keyFont{ \section{Keywords:} primary visual cortex, rate model, cortical functional development, hebbian learning, cortical horizontal connectivity, orientation map} 
+\end{abstract}
+
+\loadkeyvaluestore{/Users/jan/Documents/doc/Papers/fast_inh_paper/key_value_store.py}
 
@@
 Competition between populations of neurons has been proposed as one of the canonical computations of cortical 
-networks, and has been hypothesized to underly range of brain functions including working memory \cite{Amit1995,Durstewitz2000},
-orientation tuning \cite{Somers1995,Ben-Yishai1995}, and functional map development \cite{VonderMalsburg1973,Antolik2011}.
+networks, and has been hypothesized to underly range of brain functions including working memory \citep{Amit1995,Durstewitz2000},
+orientation tuning \citep{Somers1995,Ben-Yishai1995}, and functional map development \citep{VonderMalsburg1973,CMVC,Antolik2011}.
 In developmental models of functional cortical organization the competition occurs between populations of neurons spatially offset 
-along the cortical surface, and previous studies have shown that a specific arrangement of the effective inhibitory and excitatory 
-interactions is required: inhibition has to have longer range than excitation (a so called Mexican-Hat profile) \cite{Muir2014}.
-This is, however, in stark contrast to the know anatomical arrangement: the excitatory neurons superficial layers tend to form 
-long-range arborizations spanning multiple columns while the axons of majority of inhibitory neurons are confined to area
-only several hundreds of micrometers in diameter \cite{Buzas2006,Budd2001}.
+along the cortical surface, whereby local populations mutually rainforce each other via excitatory connections (short-range 
+excitation) while long-range inhibition facilitates competition between the local populations and stabilizes the activity in the 
+netowrk. Such so called Mexican-hat arrangement of recurrent interactions, however, is in stark contrast to the know anatomical arrangement
+of cortical cicuitry: the excitatory neurons (especially in superficial layers) tend to form long-range arborizations spanning 
+multiple columns while the axons of majority of inhibitory neurons are confined to area only several hundreds of micrometers 
+in diameter \citep{Buzas2006,Budd2001}.
 
-As pointed out by Muir et al.
-in their recent study \cite{Muir2014}, to solve this apparent conflict, previous topologically organized models cortical competition have either relied on the anatomically unsupported Mexican-hat profile of lateral interactions \cite{VonderMalsburg1973,CMVC}, or relied on other biologically unrealistic properties, such as selective targeting of inhibitory neurons by long range excitatory connections \cite{law:phd09,Rutishauser2012} or instantaneous synaptic transmission coupled with omission of recurrent inhibition \cite{Kang2003,Levy2011}.
-Model analysis by Muir et al.
-\cite{Muir2014} showed that the presence of competition across the cortical surface is predicted well by the anatomy of direct excitatory and inhibitory coupling and that multi-synaptic network effects are negligible.
-In conclusion, currently no satisfactory explanation of how topological functional organization dev?lops in cortical networks that is consistent with the present anatomical findings exists.
+As pointed out by Muir \etal\, in their recent study \citep{Muir2014}, to solve this apparent conflict, previous topologically organized models of cortical competition have either relied on the anatomically unsupported Mexican-hat profile of lateral interactions \citep{VonderMalsburg1973,CMVC}, or relied on other biologically unrealistic properties, such as selective targeting of inhibitory neurons by long range excitatory connections \citep{law:phd09,Rutishauser2012} or instantaneous synaptic transmission coupled with omission of recurrent inhibition \citep{Kang2003,Levy2011,Grabska-Barwinska2008}.
 
-All previous topologically organized models of cortical competition have either assumed instantaneous synaptic transmission \cite{VonderMalsburg1973,Kang2003,Levy2011,Grabska-Barwinska2008}} or uniform transmission delays across the different intra-cortical connections \cite{CMVC,Antolik2011}.
-This is also the case of the analysis performed by Muir et al.
-\cite{Muir2014} which questions the ability of the accepted cortical architecture to explain cortical competition.
-However, a recent study by Ohana et al.
-\cite{Ohana2012} have revealed a specific pattern of transmission delays between different neural type combinations.
+A common hypothesis about how to reconcile Mexican-hat lateral interacions with anatomical reality is that the range of the effective inhibitory influence of an excitatory neuron onto other excitatory neurons (via the disynaptic pathway from exctiatory to inhibitory to excitatory neurons - the disynaptic inhibition) will correspond to the cumulative reach of the direct excitatory and the direct inhibitory connections.
+The effective
+inhibitory interactions between excitatory neurons will thus have longer range than the direct excitatory connections, supporting the use
+of Mexican-hat lateral interaction in single population models (i.e. models without explicit sepparation of excitatory and
+inhibitory populations). Even though very intuitive, this explanation has never been explicitly demonstrated, and  
+it ommits the fact that under the reasonable null-hypothesis of equal transmission delays on all connections, 
+the recurrent disynaptic inhibition will lag the direct recurrent excitation. As it turns out this is crucial, 
+as recent model analysis by Muir \etal\,\citep{Muir2014} shows that, under the assumption of uniform tranmission delays, the presence of competition across the cortical surface is predicted well by the anatomy of direct excitatory and inhibitory coupling and that multi-synaptic network effects are negligible, effectively rejecting the disynaptic explanation behind Maxican-hat interaction.
+In conclusion, currently no satisfactory explanation of how topological functional organization develops in cortical networks that is consistent with the present anatomical findings exists.
+
+As we will show in this study, the nature of transmission delays between different neural type (excitatory,inhibitory) combinations
+holds the key to resolving this long standing open question.
+Recent study by Ohana \etal\,\citep{Ohana2012} have revealed a specific pattern of transmission delays between different neural type combinations.
 Specifically they found that the excitatory to excitatory connections are slow while the excitatory to inhibitory connections are fast.
@@
 (F) The centers of RFs of the LGN neurons form a regular latice accross the visual space covered by the model.
-This retinotopic mapping of connection fields between layers of the model is also maintained in the thalamo-cortical projection (lines between D and F)}
+This retinotopic mapping of connection fields between layers of the model is also maintained in the thalamo-cortical projection (lines between D and F).
+These thalamo-cortical projections are the only connections in the model that undergo hebbian adaptation during the simulated visually driven development.}
 \label{fig:model_architecture}
@@
 
-In sections \ref{sec:SM1} through \ref{sec:SM3} we use GCAL model \cite{Stevens2013}  which is the most advanced variant of LISSOM 
-(Laterally Interconnected Synergetically Self-Organizing Map) algorithm introduced by Miikkulainen et al. \cite{CMVC}, 
-which itself is based on earlier Self-Organizing Map models \cite{Kohonen1982}.
+In sections \ref{sec:SM1} through \ref{sec:SM3} (model 1, 2 and 3; see figure \ref{fig:model_variants}) we use GCAL model \citep{Stevens2013}  which is the most advanced variant of LISSOM (Laterally Interconnected Synergetically Self-Organizing Map) algorithm introduced by Miikkulainen \etal\,\citep{CMVC}, which itself is based on earlier Self-Organizing Map models \citep{Kohonen1982}.
 
 For clarity and consistency the following model description closely follows the methodology sections in our 
-previous work \cite{Stevens2013,Antolik2011}. The architecture of the 3 GCAL variants presented in this 
-paper is depicted in figure \ref{fig:fig:model_architecture} and figure \ref{model_variants}A-C. The models are implemented in the Topographica simulator, freely available at www.topographica.org. 
+previous work \citep{Stevens2013,Antolik2011}. The architecture of the 3 GCAL variants presented in this 
+paper is depicted in figure \ref{fig:model_architecture} and figure \ref{fig:model_variants}A-C. The models are implemented in the Topographica simulator, freely available at www.topographica.org. 
 Each of the GCAL models consists of set of sheets of model neural units, corresponding to: (1) the photoreceptors, (2) the combined effects of the retinal ganglion cells (RGC) and ON and OFF LGN cells.
-Furthermore each model has one V1 sheet of units with direct excitatory and inhibitory interactions (section \ref{sec:SM1}), or two
-V1 sheets, one corresponding to excitatory and one to inhibitory neurons (sections \ref{sec:SM2} and \ref{sec:SM3}).
+Furthermore each model has one V1 sheet of units with direct excitatory and inhibitory interactions (model 1; see figure \ref{fig:model_variants}), or two
+V1 sheets, one corresponding to excitatory and one to inhibitory neurons (model 2 and 3; see figure \ref{fig:model_variants}).
 The model sheet is a 2D array of computational elements (called units or, loosely, neurons), with activation and plasticity equations as 
@@
 %%
-where
+\noindent where
 %%
@@
 lateral excitatory $(p=E)$ and lateral inhibitory $(p=I)$.  The contribution $C_{jp}$ to the activation of unit $j$ in a cortical sheet from each projection $p$ at time $t$ is given by: 
-\begin{equation} 
+\begin{equation} \label{eq:vm_model}
 C_{jp}(t+\delta t)=\sum_{i\in F_{jp}}\Psi_{i}(t)\omega_{pij} 
@@
 %%
-\begin{equation}
+\begin{equation} \label{eq:neural_model}
 \Psi_{i}(t)= f(\sum_{p}\gamma_{p}C_{ip}(t))
@@
 %%
-where $\gamma_{p}$ is a constant determining the sign (negative for inhibitory) and strength of projection $p$. 
+\noindent where $\gamma_{p}$ is a constant determining the sign (negative for inhibitory) and strength of projection $p$. 
 The transfer function $f$ is a half-wave rectifying function that ensures positive activation values. It has a 
 variable threshold point ($\theta$) dependent on the average activity of the unit as described in the next subsection, 
-but in all cases the gain is fixed at unity.
-Table \ref{table:param} shows the strength, initial Gaussian kernel spatial extent, and the cut-off distance values for all projections for the first form of the GCAL model examined in section \ref{sec:SM1}.
-Any modifications to these base parameters
-in the GCAL model variants examined in sections \ref{sec:SM2} and \ref{sec:SM3} are then reported in the respective sections.
+but in all cases the gain is fixed at unity. The projection strength scaling factor of the afferent project $\gamma_{A}$ was set to 1.5 
+based on Stevens \etal\,\citep{Stevens2013} while the values of the lateral excitatory  and inhibitory scaling factors, $\gamma_{E}$
+and $\gamma_{I}$ respectively, were varied (see figure \ref{fig:figure1}) to find a balance between excitation and inhibition, and between afferent and lateral influences, to provide robust formation of activity bubbles that allows smooth maps to form.
 
@@
 
-where $\xi=\insval{homeostatic_learning_rate}$ is the time constant of the threshold adaptation,
+\noindent where $\xi=\insval{homeostatic_learning_rate}$ is the time constant of the threshold adaptation,
 $\mu=\insval{target_activity}$ is a constant defining the target average activity, and $\tilde{\Psi}$ is 
@@
 
+Initial connection field weights are isotropic 2D Gaussians for the lateral excitatory projection and uniformly random within a Gaussian envelope for afferent and lateral inhibitory projections.
+Specifically, a neuron located at (i, j) will have the following weights in projection $p$:
+
+\begin{equation}
+\omega_{ijp} = \frac{1}{Z_{p}} u\, exp(- \frac{x^2+y^2}{2\sigma^2_p} )
+\end{equation}
+
+\noindent where (x, y) is the sheet-coordinate location of the presynaptic neuron, $u=1$ for the lateral excitatory projection ($p$=E), and u is a scalar value drawn from a uniform random distribution for the afferent and lateral inhibitory projections ($p$ = A, I),
+$\sigma_{p}$ determines the width of the Gaussian in sheet coordinates ($\sigma_{A}=0.27$,$\sigma_{E}=0.035$,$\sigma_{I}=0.035 \sqrt{2}$), 
+and $Z_{p}$ is a constant normalizing term that ensures that the total of all weights $\omega_{ijp}$ to neuron $j$ in projection $p$ is 1.0.
+Weights for each projection are only defined within a specific maximum circular radius $r_p$ ($r_{A}=0.27$,$r_{E}=0.15$,$r_{I}=0.15 \sqrt{2}$).
+
 In the model, as images are presented to the photoreceptors, cortical afferent connection weights $\omega_{i,j,A}$ from the ON/OFF sheets are adjusted once per iterations (after cortical settling is completed) using a simple Hebbian learning rule.
@@
 
-where $\beta_{p}$ is the Hebbian learning rate for the connection fields in the two afferent projections from RGC/LGN $p \in \{\mathrm{ON},\mathrm{OFF}\}$.
+\noindent where $\beta_{p}$ is the Hebbian learning rate for the connection fields in the two afferent projections from RGC/LGN $p \in \{\mathrm{ON},\mathrm{OFF}\}$.
 I.e., the afferent weights from RGC/LGN are normalized jointly.
@@
 $\beta_{p}=\frac{\iota_{p}}{\upsilon_{p}}$, where $\upsilon_{p}$ is the number of connections per connection field in projection $p$.  
+The base parameters described here correspond to the first model variant (figure \ref{fig:figure1}).
+Any modifications of these base paremters in the other two GCAL model variants 2 and 3 (figure \ref{fig:figure2},\ref{fig:figure3}) examined in
+sections \ref{sec:SM2} and \ref{sec:SM3} are then reported in the respective sections.
 
@@
 \includegraphics[width=16cm]{./SVG/FigureModelVariants/drawing.png}
-\caption{The four considered intra-cortical connectivty variants. (A-C) Are modelled using firing rate models of neural units with instantenous 
+\caption{The four considered intra-cortical connectivty variants.
+As in figure \ref{fig:model_architecture}, in all pannels, the top line represent excitatory and bottom line inhibitory population of neurons.
+Models (A-C) are using firing rate models of neural units with instantenous
 translation of inputs into membrane-potential (see section \ref{sec:gcal}). 
@@
 (B) Variant 2, as in A but with added direct inhibitory to inhibitory connections of equal extent as the other excitatory and inhibitory projections.
-(C) Variant 3, as in A but with added long-range excitatory connections represented based on Buz\'as \etal\,\cite{Buzas2006} as a second wider gaussian.
+(C) Variant 3, as in A but with added long-range excitatory connections represented based on Buz\'as \etal\,\citep{Buzas2006} as a second wider gaussian.
 (D) The same as A but modelled with neural units that take into account membrane time-cosntant (see section \ref{sec:rate}).
-The transmission delay of the inhibitory to excitatory projections was varied in the experiments presneted in section \ref{sec:SM4}.}
+The transmission delay of the inhibitory to excitatory projections in model 4 was varied in the experiments presneted in section \ref{sec:SM4}
+(figure \ref{fig:figure4}).}
 \label{fig:model_variants}
@@
 
-\subsection{Rate model} \label{sec:rate}
+\subsection{Rate model with membrane time constant} \label{sec:rate}
 
-The architecture of this model is identical to the GCAL models presented above (see figure \ref{fig:fig:model_architecture}) 
-with the exception of the equation ? describing the neural model which is replaced with equation taking into account the 
-membrane time constant:
+The architecture of this model (model variant 4; figure \ref{fig:model_architecture} and \ref{fig:model_variants}D) is identical to the GCAL models 
+(model variants 1 through 3; figure \ref{fig:model_variants}A-C) with the exception of the equations \ref{eq:vm_model} and \ref{eq:neural_model}, which were replaced with equation taking into account the membrane time constant:
 
 \begin{equation}
-\Psi_{i}(t)= f(\sum_{p}\gamma_{p}C_{ip}(t))
+\tau_{z} \frac{\delta\Psi_{i}}{\delta t} = - \Psi_{i} + \sum_{p}\gamma_{p}\sum_{j\in F_{ip}}f(\Psi_{j}(t-\nu_{p}))\omega_{pji} 
 \end{equation}
-
-As a consequence the evolution of the network dynamics have to be simulated at higher resolution. We chose the update step to be ?ms, and we 
-let the activity in the model to settle for ?ms, thus resulting in ? settling steps as opposed to the \ref{} of the GCAL model.
-
-figure \ref{fig:fig:model_architecture} and figure \ref{model_variants}A-C
+\noindent where $\tau_{z}$ ($z \in \{E,I\}$) is the membrane time constant of excitatory and inhibitory neurons 
+($\tau_{E}$ = \insval{membrane_time_constant_E} ms, $\tau_{I}$ = \insval{membrane_time_constant_I} ms), and $\nu_{p}$ is the transmission delay 
+on the projection $p$ ($\nu_{EE}$ = \insval{td_EE}ms, $\nu_{EI}$ = \insval{td_EI} ms,$\nu_{IE}$ was varied (see section \ref{sec:SM4})).
+As a consequence the evolution of the network dynamics have to be simulated at higher resolution.
+We chose the update step to be \insval{rm_update_step}ms, and we let the activity in the model to settle for \insval{rm_settle_time}ms, thus resulting in \insval{rm_settle_steps} settling steps as opposed to the \insval{settle_steps} of the GCAL model.
+The sub-cortical parametrization of the model is identical to the model variants 1-3 (figure \ref{fig:model_variants}A-C), but because the addition of membrane time constants substantially changes the cortical dynamics
+the cortical parameters had to be re-adjusted. Specifically the strength of the thalamo-cortical connections was set to $\gamma_{A}=\insval{w_A}$,
+the strength of the lateral excitatory-to-excitatory connections was set to $\gamma_{EE}=\insval{w_EE}$, the strength of the lateral excitatory-to-inhibitory connections was set to $\gamma_{EI}=\insval{w_EI}$, and the strength of the lateral inhibitory-to-excitatory connections was set to $\gamma_{IE}=\insval{w_IE}$.
 
@@
 
-Model orientation maps are calculated based on the vector average method \cite{CMVC}. We first determine the preferred frequency of neurons 
+Model orientation maps are calculated based on the vector average method \citep{CMVC}. We first determine the preferred frequency of neurons 
 across the map. Due to the simplified stereotypical stimulus used in this study (elongated Gaussian inputs) the spatial frequency 
@@
 Next, sine grating inputs that cover the full range of remaining parameter values (combinations of all orientations and phases) are presented, and for each orientation, the peak response of the neuron is recorded.
-The orientation preference is calculated by constructing a vector for each orientation $\theta$ (between 0 and 180$\circ$), with the peak response as
+The orientation preference is calculated by constructing a vector for each orientation $\theta$ (between 0 and 180$\degree$), with the peak response as
 the length and $\theta$ as its orientation.
@@
 animals we need to utilize an automatic metric that tells how close the maps are to animal data. To this end we will 
-utilize a map-quality metric that we have recently developed \cite{Stevens2013} based on the empirical observation that pinwheel count 
-in biological orientation maps scales linearly with hypercolumn size across many different species \cite{Kaschube2010}.
+utilize a map-quality metric that we have recently developed \citep{Stevens2013} based on the empirical observation that pinwheel count 
+in biological orientation maps scales linearly with hypercolumn size across many different species \citep{Kaschube2010}.
 Specifically the pinwheel density per hypercolumn area ($\Lambda^2$) converges to $\pi$, when averaged across sufficiently 
 large cortical surface. For a detailed description of the procedure for calculating this metric we refer reader to 
-our previous work \cite{Stevens2013}, but briefly, its calculation involves three steps. First, the locations of the pinwheels
+our previous work \citep{Stevens2013}, but briefly, its calculation involves three steps. First, the locations of the pinwheels
 in the orientation map are determined as the intersections of the zero contours of the real and imaginary 
@@
 unity (high-quality map) and zero (low quality map) we pass it through normalized Gama distribution. 
-We have shown that this metric reliably distinguishes low and high quality maps \cite{Stevens2013}
+We have shown that this metric reliably distinguishes low and high quality maps \citep{Stevens2013}
 and is a valid measure for assessing how well the model orientation maps match animal data.
@@
 We will use two different computational abstractions to explore the questions at hand.
-In the fist part of the study we will use computational model that is derived from the LISSOM family of models \cite{CMVC}.
+In the fist part of the study we will use computational model that is derived from the LISSOM family of models \citep{CMVC}.
 This choice has three advantages.
@@
 
-Ohana et al.
-\cite{Ohana2012} have shown that transmission delays between different types of cortical neurons are not uniform, specifically they found that on average the transmission delays  between excitatory neurons are $\approx$1.4\,ms, from excitatory to inhibitory neurons are $\approx$0.5\,ms, and from inhibitory to excitatory neurons are $\approx$0.98\,ms.
+Ohana \etal\,\citep{Ohana2012} have shown that transmission delays between different types of cortical neurons are not uniform, specifically they found that on average the transmission delays  between excitatory neurons are $\sim$1.4\,ms, from excitatory to inhibitory neurons are $\sim$0.5\,ms, and from inhibitory to excitatory neurons are $\sim$0.98\,ms.
 The sample size for connections
@@
 show below. This gives the effective inhibition longer range, consequently fulfilling the essential requirement for cortical competition to occur.
-For didactic purposes, let us first explore one specific set of conditions under which the above statement holds exactly, we will explore the 
-situation when these conditions are relaxed in subsequent sections:
+For didactic purposes, and to simplify analytical treatment, let us first explore one specific set of conditions under which the above statement holds exactly, we will explore the situation when these conditions are relaxed in subsequent sections:
 
@@
 \begin{split}
-& R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\theta) - N_{\sigma_{i}}(\lVert y-x \rVert)R_{i}(y,t-\theta) + I_{aff}(x,t)]^+ \\
-& R_{i}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\theta)]^+
+& R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\theta_{EE}) - N_{\sigma_{i}}(\lVert y-x \rVert)R_{i}(y,t-\theta_{IE}) + I_{aff}(x,t)]^+ \\
+& R_{i}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\theta_{EI})]^+
 \end{split}
@@
 
-\noindent where $R_{p}(x,t)$ is the response of neuron of type $p$ located at position $x$ at time $t$, $I_{aff}(x,t)$ is the afferent input to neuron $x$ at time $t$, $N_{\sigma}$
-is normal distribution of variance $\sigma$, corresponding to the lateral connection kernel. When we expand for $R_{i}$ we obtain:
+\noindent where $R_{p}(x,t)$ is the response of neuron of type $p$ located at position $x$ at time $t$, $I_{aff}(x,t)$ is the afferent input to neuron at position $x$ at time $t$, $N_{\sigma}$
+is normal distribution of variance $\sigma$, corresponding to the lateral connection kernel, and $\theta_{ab}$ is delay on connections from
+neural type $a$ to neural type $b$. When we expand for $R_{i}$ we obtain:
 
 \begin{equation}
-R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\alpha) - \sum_{y}N_{\sigma_{i}}(\lVert y-x \rVert)[\sum_{z}N_{\sigma_{e}}(\lVert z-y \rVert)R_{e}(z,t-\theta-\beta)]^+]^+
+\begin{split}
+R_{e}(x,t) = & [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\theta_{EE}) - \sum_{y}N_{\sigma_{i}}(\lVert y-x \rVert)[\sum_{z}N_{\sigma_{e}}(\lVert z-y \rVert)R_{e}(z,t-\theta_{EI}-\theta_{IE})]^+\\
+& + I_{aff}(x,t)]^+
+\end{split}
 \end{equation}
 
-\noindent Because $\theta + \beta = \alpha$ (assumption 1) and because the inhibitory neurons receive only excitatory connections (assumption 2) we can further simplify the above as:
+\noindent Because $\theta_{EI} + \theta_{IE} = \theta_{EE}$ (assumption 1) and because the inhibitory neurons receive only excitatory connections (assumption 2) we can further simplify the above as:
 
 \begin{equation}
-R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\alpha) - \sum_{z}\sum_{y}N_{\sigma_{i}}(\lVert y-x \rVert) N_{\sigma_{e}}(\lVert z-y \rVert) R_{e}(z,t-\alpha)]^+
+R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\theta_{EE}) - \sum_{z}\sum_{y}N_{\sigma_{i}}(\lVert y-x \rVert) N_{\sigma_{e}}(\lVert z-y \rVert) R_{e}(z,t-\theta_{EE}) + I_{aff}(x,t)]^+
 \end{equation}
@@
 \begin{equation}
-R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\alpha) - \sum_{z}N_{\sigma_{i}} \ast  N_{\sigma_{e}}(\lVert z-x \rVert) R_{e}(z,t-\alpha)]^+ 
+R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\theta_{EE}) - \sum_{z}N_{\sigma_{i}} \ast  N_{\sigma_{e}}(\lVert z-x \rVert) R_{e}(z,t-\theta_{EE}) + I_{aff}(x,t)]^+ 
 \end{equation}
@@
 \label{eqn:last}
-R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\alpha) - \sum_{z}N_{\sqrt{2}\sigma_{e}}(\lVert z-x \rVert) R_{e}(z,t-\alpha)]^+
+R_{e}(x,t) = [\sum_{y}N_{\sigma_{e}}(\lVert y-x \rVert)R_{e}(y,t-\theta_{EE}) - \sum_{z}N_{\sqrt{2}\sigma_{e}}(\lVert z-x \rVert) R_{e}(z,t-\theta_{EE}) + I_{aff}(x,t)]^+
 \end{equation}
 
-This shows that under these specific assumptions the effective inhibitory interactions between excitatory neurons are $\sqrt{2}$ longer than excitatory and thus follow the Mexican-hat like profile required for for lateral cortical competition underlying functional map development.
+This shows that under these specific assumptions the effective inhibitory interactions between excitatory neurons are $\sqrt{2}$ longer than excitatory and thus follow the Mexican-hat like profile required for lateral cortical competition underlying functional map development.
 Even though this essential condition of cortical competition is fulfilled in this model configuration, important constraints on the extent of the effective lateral
@@
 Conveniently, the lateral interaction in the LISSOM family of models is governed by the same equations as \ref{eqn:first}.
-In the following we will use the GCAL model \cite{Stevens2013}, the latest and most robust addition to the LISSOM family of models
-to demonstrate that the above specific configuration of effective lateral excitation and inhibition permits development of high-quality orientation maps.
+In the following we will use the GCAL model \citep{Stevens2013}, the latest and most robust addition to the LISSOM family of models to demonstrate that the above specific configuration of effective lateral excitation and inhibition permits development of high-quality orientation maps.
+GCAL is the only model of stimulus dependent functional development which achieves emergence of biologically realistic
+orientation maps in terms of pinwheel density, a signature that is a usefull objective measure of orientation map quality. We will utlize this
+map quality measure to reliably detect model configurations which permit succesfull emergence of orientation maps.
 To facilitate comparison we will use exactly the same model configuration 
-as in Stevens et al. \cite{Stevens2013} (see Materials and Methods), except to three modifications necessitated by the analysis above:
+as in Stevens \etal\,\citep{Stevens2013} (see Materials and Methods), except to three modifications necessitated by the analysis above (see figure \ref{fig:model_variants}A):
 
@@
 
+In order to find working combination of parameters in point's 2 and 3, and also to show that the model is robust to certain level of changes
+in these two parameters we have performed a parameter search across these two parameters and evaluated the quality of the orientation map (see Materials and Methods) for each 
+parameter combination. As figure \ref{fig:figure1} shows, under a range of values of both parameters the model develops high-quality orientation maps indistinguishable
+from their experimental counterparts, thus concluding our first step towards showing that fast excitatory-to-inhibitory-to-excitatory loop can explain 
+how short-range inhibition can induce cortical competition and consequently development of topological organization of functional properties.
+
 \begin{figure}[htpb!] 
@@
 
-In order to find working combination of parameters in point's 2 and 3, and also to show that the model is robust to certain level of changes
-in these two parameters we have performed a parameter search across these two parameters and evaluated the quality of the orientation map (see Materials and Methods) for each 
-parameter combination. As figure \ref{fig:figure1} shows, under a range of values of both parameters the model develops high-quality orientation maps indistinguishable
-from their experimental counterparts, thus concluding our first step towards showing that fast excitatory-to-inhibitory-to-excitatory loop can explain 
-how short-range inhibition can induce cortical competition and consequently development of topological organization of functional properties.
-
 Furthermore, note that the GCAL model used in figure \ref{fig:figure1} only explicitly models excitatory neurons and assumes both direct excitatory and inhibitory
@@
 interactions between excitatory neurons, but instead add excitatory to inhibitory and inhibitory to excitatory connections of the same extend as those of excitatory to excitatory pathway (see the assumption \#5).
-As we have shown above this model should be mathematically equivalent to the simulations figure \ref{fig:figure1}D. Note, however, that the GCAL simulations represent a discrete approximation 
+As we have shown above this model should be mathematically equivalent to the simulation shown in figure \ref{fig:figure1}D. Note, however, that the GCAL simulations represent a discrete approximation 
 in both time and space of equations \ref{eqn:first},\ref{eqn:last} and we thus expect small numerical discrepancies.
@@
 
-In the previous section we have shown that under a set of specific assumptions short-range inhibition can induce 
+In the previous section we have shown that under a set of specific assumptions that facilitate analytical treatment, short-range inhibition can induce 
 effective Mexican-hat like interaction and thus support development of orientation maps. However not all assumptions we 
-made were in line with experimental evidence. In this section we will numerically show that one of these assumptions is not necessary, 
-specifically that addition of inhibitory connections does not invalidate results of the previous section.
+made were in line with experimental evidence. In this section we will show, now only numerically, that one of these assumptions is not necessary, 
+specifically that addition of inhibitory to inhibitory connections does not prevent emergence of orientation maps.
 
@@
 quality orientation map (see figure \ref{fig:figure1}). We will use GCAL configuration in which we will explicitly model the inhibitory neurons
-(see figure ?, and equation \ref{eqn:first}) and consequently also explicitly the excitatory to inhibitory and inhibitory to excitatory connections.
+(see figure \ref{fig:model_variants}B, and equation \ref{eqn:first}) and consequently also explicitly the excitatory to inhibitory and inhibitory to excitatory connections.
 Furthermore, we will add inhibitory to inhibitory connections to the model with the same extent as that of inhibitory to excitatory
@@
 (2) inhibitory to excitatory and (3) inhibitory to inhibitory connections (all other parameters remained the same as the 
-in the best parametrization found in previous section). Note that in principle there is redundancy in these parameters as
+in the best parametrization found in previous section). Note that in principle, there is redundancy in these parameters as
 the overall strength of the projections from inhibitory neurons onto both excitatory and inhibitory populations is scaled by the 
@@
 excitatory to inhibitory drive. This shows that inclusion of direct inhibitory to inhibitory interactions does not 
-invalidate the results of section \ref{sec:SM3}.
+invalidate the results of section \ref{sec:SM1} (figure \ref{fig:figure1}).
 
 \subsection{Long-range excitation} \label{sec:SM3}
-
-In section \ref{sec:SM2} we have only assumed local connectivity by assuming that both excitatory and inhibitory interaction had the same spatial extent. 
-However experimental evidence shows that excitatory cells send longer connections compared to inhibitory cells \cite{Buzas2006,Budd2001}. In this section we will
-explore what happens if we add long-range excitatory connectivity into the model analyzed in section \ref{sec:SM2}. Buz\'as et al. \cite{Buzas2006} have
-shown that the lateral connectivity in layer 2/3 can be best described as superimposition of two gabor connectivity likelyhoods, one short-range not orientation 
-specific and one long-range orientation specific. Here we will assume such dual structure, leading us to add second excitatory to excitatory and excitatory to 
-inhibitory projection into the model described in section \ref{sec:SM2}, but with 3 times larger space constant (in line with Buz\'as \etal \, \cite{Buzas2006} quantitative findings). 
-In figure \ref{fig:figure3} we have  kept the overall strength of excitatory to excitatory and excitatory to inhibitory pathways the same, corresponding to the best 
-parametrization found in \ref{sec:SM2}, but varied the relative strength between the short range and long range components (y axis of figure \ref{fig:figure3}). 
-Adding the long-range excitation changes the spatial configurations over which excitatory interactions win over inhibitory ones, as well as the 
-overall magnitude of the resulting net local excitation. Consequently, the proportion between the magnitude of the net local excitation due to the lateral interactions and the excitation
-due to afferent inputs is changed, which is a crucial parameter for map development.
-To compensate for these changes, we also systematically explore the ratio of the overall magnitudes of the lateral and afferent
-interactions (x axis of figure \ref{fig:figure3}).
-To make the parameter search computationally feasible, we perform the parameter search only in the region of parameters that allow for sufficiently strong
-long-range excitatory projections.
-As can be seen in figure \ref{fig:figure3}D and E, under appropriately strong lateral interactions, substantial long-range excitatory connection support development of high quality orientation maps, demonstrating that the proposed model is consistent with the experimentally identified long-range excitatory connectivity.
 
@@
 
+In model variants 1 nad 2 (figure \ref{fig:figure1} and \ref{fig:figure2}) we have only assumed local connectivity by setting both excitatory and inhibitory interactions to have the same spatial extent.
+However experimental evidence shows that excitatory cells send longer connections compared to inhibitory cells \citep{Buzas2006,Budd2001}.
+In this section we will explore what happens if we add long-range excitatory connectivity into the model variant 1.
+Buz\'as \etal\,\citep{Buzas2006} have
+shown that the lateral connectivity in layer 2/3 can be best described as superimposition of two gabor connectivity likelyhoods, one short-range not orientation 
+specific and one long-range orientation specific. Here we will assume such dual structure, leading us to add second excitatory to excitatory and excitatory to 
+inhibitory projection into the model variant 1, but with 3 times larger space constant (see figure \ref{fig:model_variants}C), in line with Buz\'as \etal \, \citep{Buzas2006} quantitative findings.
+
+In figure \ref{fig:figure3} we will examine what strength of the long-range excitatory connections, relative to the short-range excitatory ones (y axis in figure \ref{fig:figure3}A), leads to development of high-quality orientation map.
+Adding the long-range excitation changes the spatial configurations over which excitatory interactions win over inhibitory ones, as well as the 
+overall magnitude of the resulting net local excitation. Consequently, the proportion between the magnitude of the net local excitation due to the lateral interactions and the excitation
+due to afferent inputs is changed, which is a crucial parameter for map development.
+To compensate for these changes, we also systematically explore the ratio of the overall magnitudes of the lateral and afferent
+interactions (x axis of figure \ref{fig:figure3}).
+To make the parameter search computationally feasible, we perform the parameter search only in the region of parameters that allow for sufficiently strong
+long-range excitatory projections.
+As can be seen in figure \ref{fig:figure3}D and E, under appropriately strong lateral interactions, substantial long-range excitatory connection still permit development of high quality orientation maps, demonstrating that the proposed model is consistent with the experimentally identified long-range excitatory connectivity.
+
 \subsection{Non-equal effective excitatory and inhibitory delay} \label{sec:SM4}
@@
 In all model variants examined so far we have made the key assumption that the delay on the excitatory to excitatory connections is exactly equal 
-to the sum of excitatory to inhibitory and inhibitory to excitatory delays.
-This assumption is approximately supported by the experimental evidence \cite{Ohana2012}, but we cannot assume it holds exactly in real biological substrate.
+to the sum of excitatory to inhibitory and inhibitory to excitatory delays. This assumption is approximately supported by the experimental evidence 
+\citep{Ohana2012}, but we cannot assume it holds exactly in real biological substrate.
 However, we hypothesize, that the small discrepancies between the delay of the mono-synaptic excitatory and cumulative delay of the bi-synaptic inhibitory interactions can be absorbed in the membrane time-constant of the neurons.
 In this section we will verify this hypothesis by extending the modeling framework used thus far with a finite membrane time-constant (see Materials and Methods) and proceed to determine the magnitude of discrepancy in the delays between the excitatory and inhibitory interactions that can be managed by the model without impairing the resulting map orientation map quality.
-We use model parametrization similar to those determined in previous section \ref{sec:SM1}.
-For simplicity and computational efficiency we omit the inhibitory to inhibitory and long-range excitatory connections that have already been investigated in sections \ref{sec:SM2} and \ref{sec:SM3}.
+
+We use model parametrization similar to those determined for the
+model variant 1 (section \ref{sec:SM1}; figure \ref{fig:model_variants}A).
+For simplicity and computational efficiency we omit the inhibitory to inhibitory and long-range excitatory connections that have already been investigated with model variants 2 and 3.
 The membrane time-constant of excitatory neurons was set to \insval{membrane_time_constant_E}ms while those of inhibitory neurons to \insval{membrane_time_constant_I}ms.
-This faster inhibitory dynamics are necessary to prevent oscillations in the system \cite{Kang2003}.
-We set the excitatory to excitatory delay to 1.4 ms and excitatory to inhibitory delay to 0.5 ms based on Ohana \etal\,\cite{Ohana2012}.
+This faster inhibitory dynamics are necessary to prevent oscillations in the system \citep{Kang2003}.
+We set the excitatory to excitatory delay to 1.4 ms and excitatory to inhibitory delay to 0.5 ms based on Ohana \etal\,\citep{Ohana2012}.
 In order to understand how closely does the
@@
 to inhibitory delay, and this choice was arbitrary).
-Figure \ref{fig:figure3} shows the resulting orientation maps and associated map quality measures of models with range of differences between the delays of monosynaptic excitatory and bi-synaptic inhibitory interactions, that are in the figure expressed as the sum of the excitatory to inhibitory and inhibitory to excitatory delays.
 
@@
 
+Figure \ref{fig:figure4} shows the resulting orientation maps and associated map quality measures of models with range of differences between the delays of monosynaptic excitatory and bi-synaptic inhibitory interactions, that are in the figure expressed as the sum of the excitatory to inhibitory and inhibitory to excitatory delays minus the excitatory to excitatory delay (i.e.
+the figure shows how much slower the inhibitory disynaptic interactions were in comparison to the monosynaptic excitatory ones).
 As can be seen when the differences between the excitatory delay (1.4ms) and cumulative inhibitory delay is small (\textless 0.8 ms) high quality orientation maps develop in the model, confirming that sufficiently
@@
 However, as expected, as the difference between the delays increases the ability of the model to learn topological organized representation of orientation preference diminishes.
-Crucially, if the delays across all the projections would be equal, as is typically assumed, the model fails to develop orientation maps in line with the analysis by Muir \etal \, \cite{Muir2014}, thus confirming that the specific delay pattern between neural types identified by Ohana \etal \, \cite{Ohana2012} is key to achieving competitive dynamics in topologically organized neural models.
+Crucially, if the delays across all the projections would be equal (figure \ref{fig:figure4}), as is typically assumed, the model fails to develop orientation maps in line with the analysis by Muir \etal \, \citep{Muir2014}, thus confirming that the specific delay pattern between neural types identified by Ohana \etal\,\citep{Ohana2012} is key to achieving competitive dynamics in topologically organized neural models.
 
@@
 In this study we have shown how recent findings on dependence of neural transmission delays on the
-type of pre- and post-synaptic neuron \cite{Ohana2012} can resolve long standing question on how short-range inhibition
+type of pre- and post-synaptic neuron \citep{Ohana2012} can resolve long standing question on how short-range inhibition
 can support cortical competition and consequently the development cortical functional topological organization.
-Under simplifying assumptions, we have analytically shown how dy-synaptic inhibition that is as fast as mono-synaptic 
+Under simplifying assumptions, we have analytically shown how disynaptic inhibition that is as fast as mono-synaptic 
 excitation can extend the effective range of inhibitory interactions, in contrast to the recent analytical 
-results showing that in the case of equal synaptic delays on all connections the dy-synaptic inhibition has negligable effects \cite{Muir2014}.\
+results showing that in the case of equal synaptic delays on all connections the disynaptic inhibition has negligable effects \citep{Muir2014}.\
 We have also shown that these findings are applicable to the problem of functional development in primary visual cortex. We have then proceeded 
@@
 
-The most related past explanation of how cortical competition can arise under short range inhibition is that of Kang \etal\,\cite{Kang2003}, who have shown 
+The most related past explanation of how cortical competition can arise under short range inhibition is that of Kang \etal\,\citep{Kang2003}, who have shown 
 that under the assumption of faster inibitory time constant (as opposed to excitatory), the effective excitatory and inhibitory interactions will follow the 
-Mexican hat profile and thus support competition along the cortical surface. Indeed, in our final model that explicitly considers membrane time-constant presented in section \ref{sec:SM4}, 
-we assume that inhibitory neurons have faster membrane time constants than excitatory ones, as otherwise we observe oscilatory behavior in line with the analytical findings of Kang \etal\,\cite{Kang2003}.
-Crucially, Kang \etal\,\cite{Kang2003} however assumed instantenous neural transmission, and when this biologically implausible assumption is rectified by addition of tansmission delays that are uniform
-accross the connections between the different pre- and post-synaptic neural types, we find that the competitive dynamics in the neural model break (section \ref{sec:SM4}E) in line with the analytical and
-computational results of Muir \etal\,\cite{Muir2014}.
-However, when we replace the tansmission delays with the neural-type specific  pattern uncovered by Ohana \etal\,\cite{Ohana2012} the competitive dymanics
-in the model are rescued and we observe development of high quality orientation maps (section \ref{sec:SM4}AB), in line with the analytical results under simplifieng conditions (section \ref{sec:SM1}).
+Mexican hat profile and thus support competition along the cortical surface. Indeed, in our final model variant 4 that explicitly considers membrane time-constant presented in section \ref{sec:SM4}, 
+we assume that inhibitory neurons have faster membrane time constants than excitatory ones, as otherwise we observe oscilatory behavior in line with the analytical findings of Kang \etal\,\citep{Kang2003}.
+Crucially, Kang \etal\,\citep{Kang2003} however assumed instantenous neural transmission, and when this biologically implausible assumption is rectified by addition of tansmission delays that are uniform
+accross the connections between the different pre- and post-synaptic neural types, we find that the competitive dynamics in the neural model break (figure \ref{fig:figure4}F) in line with the analytical and
+computational results of Muir \etal\,\citep{Muir2014}.
+However, when we replace the tansmission delays with the neural-type specific  pattern uncovered by Ohana \etal\,\citep{Ohana2012} the competitive dymanics
+in the model are rescued and we observe development of high quality orientation maps (section \ref{sec:SM4}AB), in line with the analytical results under simplified conditions (section \ref{sec:SM1}).
 
@@
 involved preclude systematic search across the full parameter space, and we have only explored parameters that we empirically found to have biggest impact. Finally, one simplifying assumption 
-that we have not treated in this study is the lack of direct thalamic input onto inhibitory cells. Since inhibitory cells in cortical layer 4 do receive thalamic input \cite{Binzegger2004} the 
+that we have not treated in this study is the lack of direct thalamic input onto inhibitory cells. Since inhibitory cells in cortical layer 4 do receive thalamic input \citep{Binzegger2004} the 
 inclusion of external input in the inhibitory population needs to be considered in future.
@@
 
-\bibliographystyle{abbrv}
+\section*{Conflict of Interest Statement}
+
+The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
+
+\section*{Author Contributions}
+
+JA performed the calculus, simulations and wrote the manusript.
+
+\section*{Funding}
+The research leading to these results has received funding from the European Union Seventh Framework Programme (FP7/2007-2013) under grant
+agreement no. 604102 (Human Brain Project).
+
+\section*{Acknowledgments}
+I would like to thank James A. Bednar and Andrew Davison for reviewing the original draft and 
+providing valuable comments on the presented research.
+
+\bibliographystyle{frontiersinSCNS_ENG_HUMS}
 \bibliography{./mendeley}

### 2017-01-10 key_value_store.py

@@ -28,4 +28,14 @@
 	# time constant Model
 	'membrane_time_constant_E' : 2, #ms
 	'membrane_time_constant_I' : 0.5, #ms
+	'rm_update_step' : 0.1, #ms
+	'rm_settle_time' : '150', #ms #!!!! double check when return to GIF!!!!!!
+	'rm_settle_steps' : '1500',#!!!! double check when return to GIF!!!!!!
+	'td_EE' : '1.4',
+	'td_EI' : '0.5',
+	'w_A' : 0.5,
+	'w_EE' : '?',
+	'w_EI' : 1.0,
+	'w_IE' : '?',
+	
 }

### 2017-01-11 paper.tex
minor changes

@@
 Competitive interactions are believed to underlie variety of cortical processing, ranging from memory formation, attention and 
-development of cortical functional organization (e.g. development of orientaion maps in primary visual cortex). 
+development of cortical functional organization (e.g. development of orientation maps in primary visual cortex). 
 In the latter case, the competitive interactions happen along the cortical 
 surface, with local populations of neurons reinforcing each other while competing with those displaced more distally. This 
-specific configuration of lateral interactions is however in stark contrast with the known properties of the antatomical substrate, i.e.
-excitatory connections (mediating reinforcment) having longer reach than inhibitory ones (mediating competition). No 
+specific configuration of lateral interactions is however in stark contrast with the known properties of the anatomical substrate, i.e.
+excitatory connections (mediating reinforcement) having longer reach than inhibitory ones (mediating competition). No 
 satisfactory biologically plausible resolution of this conflict between anatomical measures, and assumed cortical function has been proposed.
@@
 excitation has approximately the same delay, as the combined delays of the disynaptic inhibitory interactions between excitatory neurons 
-(i.e. the sum of delays from excitatory to inhibitory and from inhibitory to excitatory neuron). Here we show that this spefic pattern
+(i.e. the sum of delays from excitatory to inhibitory and from inhibitory to excitatory neuron). Here we show that this specific pattern
 of delays represents a biologically plausible explanation for how short-range inhibition can support competitive interactions in cortex
-that underlie development of orientation maps in primary visual cortex. We demonstrate this statement analytically under simplyfing conditions, 
+that underlie development of orientation maps in primary visual cortex. We demonstrate this statement analytically under simplifying conditions, 
 and subsequently show using network simulations that development of orientation maps is preserved when long-range excitation, 
@@
 \tiny
-\keyFont{ \section{Keywords:} primary visual cortex, rate model, cortical functional development, hebbian learning, cortical horizontal connectivity, orientation map} 
+\keyFont{ \section{Keywords:} primary visual cortex, rate model, cortical functional development, Hebbian learning, cortical horizontal connectivity, orientation map} 
 \end{abstract}
@@
 In developmental models of functional cortical organization the competition occurs between populations of neurons spatially offset 
-along the cortical surface, whereby local populations mutually rainforce each other via excitatory connections (short-range 
+along the cortical surface, whereby local populations mutually reinforce each other via excitatory connections (short-range 
 excitation) while long-range inhibition facilitates competition between the local populations and stabilizes the activity in the 
-netowrk. Such so called Mexican-hat arrangement of recurrent interactions, however, is in stark contrast to the know anatomical arrangement
-of cortical cicuitry: the excitatory neurons (especially in superficial layers) tend to form long-range arborizations spanning 
+network. Such so called Mexican-hat arrangement of recurrent interactions, however, is in stark contrast to the know anatomical arrangement
+of cortical circuitry: the excitatory neurons (especially in superficial layers) tend to form long-range arborizations spanning 
 multiple columns while the axons of majority of inhibitory neurons are confined to area only several hundreds of micrometers 
@@
 
-A common hypothesis about how to reconcile Mexican-hat lateral interacions with anatomical reality is that the range of the effective inhibitory influence of an excitatory neuron onto other excitatory neurons (via the disynaptic pathway from exctiatory to inhibitory to excitatory neurons - the disynaptic inhibition) will correspond to the cumulative reach of the direct excitatory and the direct inhibitory connections.
+A common hypothesis about how to reconcile Mexican-hat lateral interactions with anatomical reality is that the range of the effective inhibitory influence of an excitatory neuron onto other excitatory neurons (via the disynaptic pathway from excitatory to inhibitory to excitatory neurons - the disynaptic inhibition) will correspond to the cumulative reach of the direct excitatory and the direct inhibitory connections.
 The effective
 inhibitory interactions between excitatory neurons will thus have longer range than the direct excitatory connections, supporting the use
-of Mexican-hat lateral interaction in single population models (i.e. models without explicit sepparation of excitatory and
+of Mexican-hat lateral interaction in single population models (i.e. models without explicit separation of excitatory and
 inhibitory populations). Even though very intuitive, this explanation has never been explicitly demonstrated, and  
-it ommits the fact that under the reasonable null-hypothesis of equal transmission delays on all connections, 
+it omits the fact that under the reasonable null-hypothesis of equal transmission delays on all connections, 
 the recurrent disynaptic inhibition will lag the direct recurrent excitation. As it turns out this is crucial, 
-as recent model analysis by Muir \etal\,\citep{Muir2014} shows that, under the assumption of uniform tranmission delays, the presence of competition across the cortical surface is predicted well by the anatomy of direct excitatory and inhibitory coupling and that multi-synaptic network effects are negligible, effectively rejecting the disynaptic explanation behind Maxican-hat interaction.
+as recent model analysis by Muir \etal\,\citep{Muir2014} shows that, under the assumption of uniform transmission delays, the presence of competition across the cortical surface is predicted well by the anatomy of direct excitatory and inhibitory coupling and that multi-synaptic network effects are negligible, effectively rejecting the disynaptic explanation behind Maxican-hat interaction.
 In conclusion, currently no satisfactory explanation of how topological functional organization develops in cortical networks that is consistent with the present anatomical findings exists.
@@
 \caption{General model architecture. (A) Inhibitory cortical population corresponds to a regular lattice of units in cortical space. (B) As in A but 
-for excitatory population. (C) Thalamic neurons are modeled as simple difference-of-Gaussian filter followed by a threshold-linear trasfer function. 
+for excitatory population. (C) Thalamic neurons are modeled as simple difference-of-Gaussian filter followed by a threshold-linear transfer function. 
 (D) The intra-cortical connectivity between the excitatory and inhibitory populations and their transmission delays.
-The intra-cortical connectivity is the only aspect of the general architecture presented here that changes between the diffrent model variants explored in this study and each variant is detail in figure \ref{fig:model_variants}.
-(F) The centers of RFs of the LGN neurons form a regular latice accross the visual space covered by the model.
+The intra-cortical connectivity is the only aspect of the general architecture presented here that changes between the different model variants explored in this study and each variant is detail in figure \ref{fig:model_variants}.
+(F) The centers of RFs of the LGN neurons form a regular latice across the visual space covered by the model.
 This retinotopic mapping of connection fields between layers of the model is also maintained in the thalamo-cortical projection (lines between D and F).
-These thalamo-cortical projections are the only connections in the model that undergo hebbian adaptation during the simulated visually driven development.}
+These thalamo-cortical projections are the only connections in the model that undergo Hebbian adaptation during the simulated visually driven development.}
 \label{fig:model_architecture}
@@
 The base parameters described here correspond to the first model variant (figure \ref{fig:figure1}).
-Any modifications of these base paremters in the other two GCAL model variants 2 and 3 (figure \ref{fig:figure2},\ref{fig:figure3}) examined in
+Any modifications of these base parameters in the other two GCAL model variants 2 and 3 (figure \ref{fig:figure2},\ref{fig:figure3}) examined in
 sections \ref{sec:SM2} and \ref{sec:SM3} are then reported in the respective sections.
@@
 \includegraphics[width=16cm]{./SVG/FigureModelVariants/drawing.png}
-\caption{The four considered intra-cortical connectivty variants.
-As in figure \ref{fig:model_architecture}, in all pannels, the top line represent excitatory and bottom line inhibitory population of neurons.
-Models (A-C) are using firing rate models of neural units with instantenous
+\caption{The four considered intra-cortical connectivity variants.
+As in figure \ref{fig:model_architecture}, in all panels, the top line represent excitatory and bottom line inhibitory population of neurons.
+Models (A-C) are using firing rate models of neural units with instantaneous
 translation of inputs into membrane-potential (see section \ref{sec:gcal}). 
@@
 (B) Variant 2, as in A but with added direct inhibitory to inhibitory connections of equal extent as the other excitatory and inhibitory projections.
-(C) Variant 3, as in A but with added long-range excitatory connections represented based on Buz\'as \etal\,\citep{Buzas2006} as a second wider gaussian.
-(D) The same as A but modelled with neural units that take into account membrane time-cosntant (see section \ref{sec:rate}).
-The transmission delay of the inhibitory to excitatory projections in model 4 was varied in the experiments presneted in section \ref{sec:SM4}
+(C) Variant 3, as in A but with added long-range excitatory connections represented based on Buz\'as \etal\,\citep{Buzas2006} as a second wider Gaussian.
+(D) The same as A but modeled with neural units that take into account membrane time-constant (see section \ref{sec:rate}).
+The transmission delay of the inhibitory to excitatory projections in model 4 was varied in the experiments presented in section \ref{sec:SM4}
 (figure \ref{fig:figure4}).}
@@
 GCAL is the only model of stimulus dependent functional development which achieves emergence of biologically realistic
-orientation maps in terms of pinwheel density, a signature that is a usefull objective measure of orientation map quality. We will utlize this
-map quality measure to reliably detect model configurations which permit succesfull emergence of orientation maps.
+orientation maps in terms of pinwheel density, a signature that is a useful objective measure of orientation map quality. We will utilize this
+map quality measure to reliably detect model configurations which permit successful emergence of orientation maps.
 To facilitate comparison we will use exactly the same model configuration 
@@
 
-In model variants 1 nad 2 (figure \ref{fig:figure1} and \ref{fig:figure2}) we have only assumed local connectivity by setting both excitatory and inhibitory interactions to have the same spatial extent.
+In model variants 1 and 2 (figure \ref{fig:figure1} and \ref{fig:figure2}) we have only assumed local connectivity by setting both excitatory and inhibitory interactions to have the same spatial extent.
 However experimental evidence shows that excitatory cells send longer connections compared to inhibitory cells \citep{Buzas2006,Budd2001}.
@@
 Buz\'as \etal\,\citep{Buzas2006} have
-shown that the lateral connectivity in layer 2/3 can be best described as superimposition of two gabor connectivity likelyhoods, one short-range not orientation 
+shown that the lateral connectivity in layer 2/3 can be best described as superimposition of two gabor connectivity likelihoods, one short-range not orientation 
 specific and one long-range orientation specific. Here we will assume such dual structure, leading us to add second excitatory to excitatory and excitatory to 
@@
 excitation can extend the effective range of inhibitory interactions, in contrast to the recent analytical 
-results showing that in the case of equal synaptic delays on all connections the disynaptic inhibition has negligable effects \citep{Muir2014}.\
+results showing that in the case of equal synaptic delays on all connections the disynaptic inhibition has negligible effects \citep{Muir2014}.\
 We have also shown that these findings are applicable to the problem of functional development in primary visual cortex. We have then proceeded 
@@
 Overall, this study represents an important advance in our understanding how orientation map development can be supported by cortical neural substrate. These
-results generelize to the development of other functional features in the cortex and other cortical competition based mechanisms in general.
+results generalize to the development of other functional features in the cortex and other cortical competition based mechanisms in general.
 
 The most related past explanation of how cortical competition can arise under short range inhibition is that of Kang \etal\,\citep{Kang2003}, who have shown 
-that under the assumption of faster inibitory time constant (as opposed to excitatory), the effective excitatory and inhibitory interactions will follow the 
+that under the assumption of faster inhibitory time constant (as opposed to excitatory), the effective excitatory and inhibitory interactions will follow the 
 Mexican hat profile and thus support competition along the cortical surface. Indeed, in our final model variant 4 that explicitly considers membrane time-constant presented in section \ref{sec:SM4}, 
-we assume that inhibitory neurons have faster membrane time constants than excitatory ones, as otherwise we observe oscilatory behavior in line with the analytical findings of Kang \etal\,\citep{Kang2003}.
-Crucially, Kang \etal\,\citep{Kang2003} however assumed instantenous neural transmission, and when this biologically implausible assumption is rectified by addition of tansmission delays that are uniform
-accross the connections between the different pre- and post-synaptic neural types, we find that the competitive dynamics in the neural model break (figure \ref{fig:figure4}F) in line with the analytical and
+we assume that inhibitory neurons have faster membrane time constants than excitatory ones, as otherwise we observe oscillatory behavior in line with the analytical findings of Kang \etal\,\citep{Kang2003}.
+Crucially, Kang \etal\,\citep{Kang2003} however assumed instantaneous neural transmission, and when this biologically implausible assumption is rectified by addition of transmission delays that are uniform
+across the connections between the different pre- and post-synaptic neural types, we find that the competitive dynamics in the neural model break (figure \ref{fig:figure4}F) in line with the analytical and
 computational results of Muir \etal\,\citep{Muir2014}.
-However, when we replace the tansmission delays with the neural-type specific  pattern uncovered by Ohana \etal\,\citep{Ohana2012} the competitive dymanics
+However, when we replace the transmission delays with the neural-type specific  pattern uncovered by Ohana \etal\,\citep{Ohana2012} the competitive dynamics
 in the model are rescued and we observe development of high quality orientation maps (section \ref{sec:SM4}AB), in line with the analytical results under simplified conditions (section \ref{sec:SM1}).
 
-The analytical results in this study were obtained only under simplifying assumptions, specificaly instantaneous translation of inputs to membrane potential, equal extent of 
-excitatory and inhibitory connections and lack of inhibitory-to-inhibitory interactions. Even though we have shown computationally that these assumptions are not neccessary 
-for achieving the cortical competition and the consequential orientation map development sought in these study, further analytical work than can circuimvent these simplifications would 
-undoubtly provide deeper understanding of the dynamics of the studied neural system and its dependence on the various parameters. This sentiment is underlied by the parameter 
+The analytical results in this study were obtained only under simplifying assumptions, specifically instantaneous translation of inputs to membrane potential, equal extent of 
+excitatory and inhibitory connections and lack of inhibitory-to-inhibitory interactions. Even though we have shown computationally that these assumptions are not necessary 
+for achieving the cortical competition and the consequential orientation map development sought in these study, further analytical work than can circumvent these simplifications would 
+undoubtedly provide deeper understanding of the dynamics of the studied neural system and its dependence on the various parameters. This sentiment is underlaid by the parameter 
 explorations presented here, which show that even though the model is robust to changes of the considered parameters, the existence or not of dynamics supporting development of 
@@
 
-In this paper we have decided to investigate cortical competitive mechamnisms through the prism of orientation map development.
+In this paper we have decided to investigate cortical competitive mechanisms through the prism of orientation map development.
 The advantage of this approach is that it allows us not only to show that some form of competition is possible, but also that it is of the form that actually supports implementation of specific cortical computations.
 Given that we show that our model implements effective Mexican hat lateral interactions (section \ref{sec:SM1}) and these have been in the past shown to be sufficient to explain cortical organization of other functional features (i.e.
-retinotopy, ocular dominance, spatial frequency and color) it is very likely that our results will generalzie to these other dimensions of sensory input as well.
-Cortical competition of other forms have been proposed to underly broad variety of other cortical operations, including associative memory, noise suppression, decission making, saliency detection and other forms of attentional computations.
-Even though additional work will be required to determine if the mechanisms proposed here can generelize to these other neural computations, this study offers
-promising framework for anatomically plausible mechinistic explanations of these important aspect of brain function.
+retinotopy, ocular dominance, spatial frequency and color) it is very likely that our results will generalize to these other dimensions of sensory input as well.
+Cortical competition of other forms have been proposed to underly broad variety of other cortical operations, including associative memory, noise suppression, decision making, saliency detection and other forms of attentional computations.
+Even though additional work will be required to determine if the mechanisms proposed here can generalize to these other neural computations, this study offers
+promising framework for anatomically plausible mechanistic explanations of these important aspect of brain function.
 

### 2017-01-11 key_value_store.py

@@ -34,8 +34,8 @@
 	'td_EE' : '1.4',
 	'td_EI' : '0.5',
 	'w_A' : 0.5,
-	'w_EE' : '?',
+	'w_EE' : 3.5,
 	'w_EI' : 1.0,
-	'w_IE' : '?',
+	'w_IE' : 2.94,
 	
 }

### 2017-01-11 build_figures.sh
added tifs

@@ -6,4 +6,5 @@ for x in ./SVG/Figure1/figure1.svg  ./SVG/Figure2/figure2.svg ./SVG/Figure3/figu
   base=`basename $x .svg`
   pwd
   cd ${dir}; inkscape ${base}.svg --export-area-drawing --export-dpi=600 -e ${base}.png; cd ../../
+  cd ${dir}; inkscape ${base}.svg --export-area-drawing --export-dpi=600 -e ${base}.tif; cd ../../
 done

### 2017-01-11 paper.tex
dsa

@@
 Competition between populations of neurons has been proposed as one of the canonical computations of cortical 
-networks, and has been hypothesized to underly range of brain functions including working memory \citep{Amit1995,Durstewitz2000},
-orientation tuning \citep{Somers1995,Ben-Yishai1995}, and functional map development \citep{VonderMalsburg1973,CMVC,Antolik2011}.
+networks, and has been hypothesized to underly range of brain functions including working memory \citep{Amit1995,Durstewitz2000,Deco2003},
+attention \citep{Deco2003}, orientation tuning \citep{Somers1995,Ben-Yishai1995,Antolik2011}, and functional map development \citep{VonderMalsburg1973,CMVC,Antolik2011}.
 In developmental models of functional cortical organization the competition occurs between populations of neurons spatially offset

### 2017-02-23 paper.tex
changed title
corrections from Evan

--- 2017-01-11 paper.tex
@@
 
-\title[Running Title]{Fast inhibitory loop explains functional map development with short-range inhibition.}
+\title[Running Title]{Rapid long-range disynaptic inhibition explains the formation of cortical orientation maps.}
 
@@
 
-Competitive interactions are believed to underlie variety of cortical processing, ranging from memory formation, attention and 
+Competitive interactions are believed to underlie many types of cortical processing, ranging from memory formation, attention and 
 development of cortical functional organization (e.g. development of orientation maps in primary visual cortex). 
@@
 excitation has approximately the same delay, as the combined delays of the disynaptic inhibitory interactions between excitatory neurons 
-(i.e. the sum of delays from excitatory to inhibitory and from inhibitory to excitatory neuron). Here we show that this specific pattern
-of delays represents a biologically plausible explanation for how short-range inhibition can support competitive interactions in cortex
-that underlie development of orientation maps in primary visual cortex. We demonstrate this statement analytically under simplifying conditions, 
+(i.e. the sum of delays from excitatory to inhibitory and from inhibitory to excitatory neurons). Here we show that this specific pattern
+of delays represents a biologically plausible explanation for how short-range inhibition can support competitive interactions
+that underlie the development of orientation maps in primary visual cortex. We demonstrate this statement analytically under simplifying conditions, 
 and subsequently show using network simulations that development of orientation maps is preserved when long-range excitation, 
@@
 Competition between populations of neurons has been proposed as one of the canonical computations of cortical 
-networks, and has been hypothesized to underly range of brain functions including working memory \citep{Amit1995,Durstewitz2000},
+networks, and has been hypothesized to underly a range of brain functions including working memory \citep{Amit1995,Durstewitz2000},
 orientation tuning \citep{Somers1995,Ben-Yishai1995}, and functional map development \citep{VonderMalsburg1973,CMVC,Antolik2011}.
@@
 of cortical circuitry: the excitatory neurons (especially in superficial layers) tend to form long-range arborizations spanning 
-multiple columns while the axons of majority of inhibitory neurons are confined to area only several hundreds of micrometers 
+multiple columns while the axons of a majority of inhibitory neurons are confined to an area only several hundreds of micrometers 
 in diameter \citep{Buzas2006,Budd2001}.
@@
 inhibitory interactions between excitatory neurons will thus have longer range than the direct excitatory connections, supporting the use
-of Mexican-hat lateral interaction in single population models (i.e. models without explicit separation of excitatory and
+of Mexican-hat lateral interactions in population models (i.e. models without explicit separation of excitatory and
 inhibitory populations). Even though very intuitive, this explanation has never been explicitly demonstrated, and  
-it omits the fact that under the reasonable null-hypothesis of equal transmission delays on all connections, 
+it omits the fact that under the reasonable null-hypothesis of equal transmission delays for all connections, 
 the recurrent disynaptic inhibition will lag the direct recurrent excitation. As it turns out this is crucial, 
@@
 
-As we will show in this study, the nature of transmission delays between different neural type (excitatory,inhibitory) combinations
+As we will show in this study, the nature of the transmission delays between different neural type (excitatory,inhibitory) combinations
 holds the key to resolving this long standing open question.
-Recent study by Ohana \etal\,\citep{Ohana2012} have revealed a specific pattern of transmission delays between different neural type combinations.
+A recent study by Ohana \etal\,\citep{Ohana2012} revealed a specific pattern of transmission delays between different neural type combinations.
 Specifically they found that the excitatory to excitatory connections are slow while the excitatory to inhibitory connections are fast.
@@
 to effective cortical competition.
-We employ computational models to show that the fast excitatory-to-inhibitory-to-excitatory pathway allows for the disynptic inhibition to generate the effective Mexican-hat like lateral interactions, thus for the first time
+We employ computational models to show that the fast excitatory-to-inhibitory-to-excitatory pathway allows for the disynaptic inhibition to generate the effective Mexican-hat like lateral interactions, thus for the first time
 explaining competition in topologically organized cortical networks with no biologically implausible assumptions. To demonstrate the
@@
 
-Here we describe the two models and their variants used in this study and finish with description of a measure
-for assessing the extent to which a model orientation maps resemble their biological counterparts. Since this study heavily leans on methodology 
-developed in our previous studies, we offer here only shortened description and refer readers to the original articles for details.
+Here we describe the two models and their variants used in this study and finish with a description of a measure
+for assessing the extent to which model orientation maps resemble their biological counterparts. Since this study heavily leans on methodology 
+developed in our previous studies, we offer here a short description and refer readers to the original articles for the details.
 
@@
 \includegraphics[width=16cm]{./SVG/FigureModelArchitecture/drawing.png}
-\caption{General model architecture. (A) Inhibitory cortical population corresponds to a regular lattice of units in cortical space. (B) As in A but 
-for excitatory population. (C) Thalamic neurons are modeled as simple difference-of-Gaussian filter followed by a threshold-linear transfer function. 
+\caption{General model architecture. (A) The inhibitory cortical population corresponds to a regular lattice of units in cortical space. (B) As in A but 
+for the excitatory population. (C) Thalamic neurons are modeled as simple difference-of-Gaussian filters followed by a threshold-linear transfer function. 
 (D) The intra-cortical connectivity between the excitatory and inhibitory populations and their transmission delays.
-The intra-cortical connectivity is the only aspect of the general architecture presented here that changes between the different model variants explored in this study and each variant is detail in figure \ref{fig:model_variants}.
-(F) The centers of RFs of the LGN neurons form a regular latice across the visual space covered by the model.
+The intra-cortical connectivity is the only aspect of the general architecture presented here that changes between the different model variants explored in this study and each variant is detailed in figure \ref{fig:model_variants}.
+(F) The RF centers of the LGN neurons form a regular latice across the visual space covered by the model.
 This retinotopic mapping of connection fields between layers of the model is also maintained in the thalamo-cortical projection (lines between D and F).
-These thalamo-cortical projections are the only connections in the model that undergo Hebbian adaptation during the simulated visually driven development.}
+These thalamo-cortical projections are the only connections in the model that undergo Hebbian adaptation during simulated visually driven development.}
 \label{fig:model_architecture}
@@
 
-In sections \ref{sec:SM1} through \ref{sec:SM3} (model 1, 2 and 3; see figure \ref{fig:model_variants}) we use GCAL model \citep{Stevens2013}  which is the most advanced variant of LISSOM (Laterally Interconnected Synergetically Self-Organizing Map) algorithm introduced by Miikkulainen \etal\,\citep{CMVC}, which itself is based on earlier Self-Organizing Map models \citep{Kohonen1982}.
+In sections \ref{sec:SM1} through \ref{sec:SM3} (model 1, 2 and 3; see figure \ref{fig:model_variants}) we use the GCAL model \citep{Stevens2013}  which is the most advanced variant of the LISSOM (Laterally Interconnected Synergetically Self-Organizing Map) algorithm introduced by Miikkulainen \etal\,\citep{CMVC}, which itself is based on earlier Self-Organizing Map models \citep{Kohonen1982}.
 
@@
 paper is depicted in figure \ref{fig:model_architecture} and figure \ref{fig:model_variants}A-C. The models are implemented in the Topographica simulator, freely available at www.topographica.org. 
-Each of the GCAL models consists of set of sheets of model neural units, corresponding to: (1) the photoreceptors, (2) the combined effects of the retinal ganglion cells (RGC) and ON and OFF LGN cells.
+Each of the GCAL models consists of a set of sheets of model neural units, corresponding to: (1) the photoreceptors, (2) the combined effects of the retinal ganglion cells (RGC) and ON and OFF LGN cells.
 Furthermore each model has one V1 sheet of units with direct excitatory and inhibitory interactions (model 1; see figure \ref{fig:model_variants}), or two
@@
 
-\subsubsection{Simulation run-time and stimuly}
+\subsubsection{Simulation run-time and stimuli}
 
@@
 The activation value $\psi_{i}$ of unit i in the photoreceptor sheet (P) is given by the gray-scale value in the chosen image at that point.
-In this study we use two-dimensional elongated Gaussian patterns, who's center coordinates and orientation is sampled from uniform 
-distribution that covers the area of photoreceptor sheet and full range of orientations.
-Each \insval{settle_steps} time steps \insval{num_gaussian_stim} Gaussian patterns are superimposed to form the visual input.
+In this study we use two-dimensional elongated Gaussian patterns, whose center coordinates and orientation are sampled from a uniform 
+distribution that covers the area of the photoreceptor sheet and the full range of orientations.
+Every \insval{settle_steps} time steps \insval{num_gaussian_stim} Gaussian patterns are superimposed to form the visual input.
 
@@
 The lateral RGC/LGN weights are 2D Gaussians with kernel size $\sigma=\insval{lgn_lateral_sigma}$.
-The center of the afferent and lateral connection field of each ON/OFF unit is mapped to the location in the photoreceptor and lgn sheet corresponding to the location of that unit in sheet coordinates, making all these projections retinotopic.
+The center of the afferent and lateral connection field of each ON/OFF unit is mapped to the location in the photoreceptor and LGN sheet corresponding to the location of that unit in sheet coordinates, making all these projections retinotopic.
 
@@
 sheet of projection $p$ from which unit $j$ receives input (its connection field $F_{jp}$), and $\omega_{pij}$ is the connection
-weight from unit $i$ in the input sheet of projection $p$ to unit $j$ in the in the output sheet of projection $p$. All connection field
+weight from unit $i$ in the input sheet of projection $p$ to unit $j$ in the output sheet of projection $p$. All connection field
 weights are initialized with uniform random noise multiplied by a 2D Gaussian profile, cut off at the distance specified below.
@@
 variable threshold point ($\theta$) dependent on the average activity of the unit as described in the next subsection, 
-but in all cases the gain is fixed at unity. The projection strength scaling factor of the afferent project $\gamma_{A}$ was set to 1.5 
+but in all cases the gain is fixed at unity. The projection strength scaling factor of the afferent projection $\gamma_{A}$ was set to 1.5 
 based on Stevens \etal\,\citep{Stevens2013} while the values of the lateral excitatory and inhibitory scaling factors, $\gamma_{E}$
-and $\gamma_{I}$ respectively, were varied (see figure \ref{fig:figure1}) to find a balance between excitation and inhibition, and between afferent and lateral influences, to provide robust formation of activity bubbles that allows smooth maps to form.
+and $\gamma_{I}$ respectively, were varied (see figure \ref{fig:figure1}) to find a balance between excitation and inhibition, and between afferent and lateral influences, to provide robust formation of activity bubbles that facilitates the formation of smooth maps.
 
@@
 
-The threshold $\theta$ of all cortical excitatory units is updated at the end of each settling phase based on following 
+The threshold $\theta$ of all cortical excitatory units is updated at the end of each settling phase based on the following 
 equations:
@@
 
-\noindent where $\Psi(t)$ is the output of the unit at time t and $\chi=\insval{activity_average_smoothing}$ is time constant 
+\noindent where $\Psi(t)$ is the output of the unit at time t and $\chi=\insval{activity_average_smoothing}$ is a time constant 
 controlling the decay of the influence of the past activities. The effect of this scaling mechanism is to bring the average activity of each 
@@
 
-Initial connection field weights are isotropic 2D Gaussians for the lateral excitatory projection and uniformly random within a Gaussian envelope for afferent and lateral inhibitory projections.
+The initial connection field weights are isotropic 2D Gaussians for the lateral excitatory projection and uniformly random within a Gaussian envelope for afferent and lateral inhibitory projections.
 Specifically, a neuron located at (i, j) will have the following weights in projection $p$:
@@
 
-In the model, as images are presented to the photoreceptors, cortical afferent connection weights $\omega_{i,j,A}$ from the ON/OFF sheets are adjusted once per iterations (after cortical settling is completed) using a simple Hebbian learning rule.
+In the model, as images are presented to the photoreceptors, the cortical afferent connection weights $\omega_{i,j,A}$ from the ON/OFF sheets are adjusted once per iteration (after cortical settling is completed) using a simple Hebbian learning rule.
 This rule results in connections that reflect correlations between the presynaptic ON/OFF unit activities and the postsynaptic cortical response.
@@
 \caption{The four considered intra-cortical connectivity variants.
-As in figure \ref{fig:model_architecture}, in all panels, the top line represent excitatory and bottom line inhibitory population of neurons.
+As in figure \ref{fig:model_architecture}, in all panels, the top line represents the excitatory and the bottom line the inhibitory population of neurons.
 Models (A-C) are using firing rate models of neural units with instantaneous
 translation of inputs into membrane-potential (see section \ref{sec:gcal}). 
-(A) Variant 1, assuming only local equaly wide excitatory and inhibitory connectivity, and ignoring direct inhibitory to inhibitory interactions.
+(A) Variant 1, assuming only local equally wide excitatory and inhibitory connectivity, and ignoring direct inhibitory to inhibitory interactions.
 (B) Variant 2, as in A but with added direct inhibitory to inhibitory connections of equal extent as the other excitatory and inhibitory projections.
@@
 The architecture of this model (model variant 4; figure \ref{fig:model_architecture} and \ref{fig:model_variants}D) is identical to the GCAL models 
-(model variants 1 through 3; figure \ref{fig:model_variants}A-C) with the exception of the equations \ref{eq:vm_model} and \ref{eq:neural_model}, which were replaced with equation taking into account the membrane time constant:
+(model variants 1 through 3; figure \ref{fig:model_variants}A-C) with the exception of the equations \ref{eq:vm_model} and \ref{eq:neural_model}, which were replaced with an equation taking into account the membrane time constant:
 
@@
 ($\tau_{E}$ = \insval{membrane_time_constant_E} ms, $\tau_{I}$ = \insval{membrane_time_constant_I} ms), and $\nu_{p}$ is the transmission delay 
-on the projection $p$ ($\nu_{EE}$ = \insval{td_EE}ms, $\nu_{EI}$ = \insval{td_EI} ms,$\nu_{IE}$ was varied (see section \ref{sec:SM4})).
-As a consequence the evolution of the network dynamics have to be simulated at higher resolution.
-We chose the update step to be \insval{rm_update_step}ms, and we let the activity in the model to settle for \insval{rm_settle_time}ms, thus resulting in \insval{rm_settle_steps} settling steps as opposed to the \insval{settle_steps} of the GCAL model.
+for projection $p$ ($\nu_{EE}$ = \insval{td_EE}ms, $\nu_{EI}$ = \insval{td_EI} ms,$\nu_{IE}$ was varied (see section \ref{sec:SM4})).
+As a consequence the evolution of the network dynamics has to be simulated at higher resolution.
+We chose the update step to be \insval{rm_update_step}ms, and we let the activity in the model settle for \insval{rm_settle_time}ms, thus resulting in \insval{rm_settle_steps} settling steps as opposed to the \insval{settle_steps} of the GCAL model.
 The sub-cortical parametrization of the model is identical to the model variants 1-3 (figure \ref{fig:model_variants}A-C), but because the addition of membrane time constants substantially changes the cortical dynamics
@@
 value for the spatial frequency parameter across all subsequent analysis.
-Next, sine grating inputs that cover the full range of remaining parameter values (combinations of all orientations and phases) are presented, and for each orientation, the peak response of the neuron is recorded.
+Next, sinusoidal grating inputs that cover the full range of remaining parameter values (combinations of all orientations and phases) are presented, and for each orientation, the peak response of the neuron is recorded.
 The orientation preference is calculated by constructing a vector for each orientation $\theta$ (between 0 and 180$\degree$), with the peak response as
@@
 
-In order to asses whether the proposed model develops orientation maps that match the structure of those found in real 
+In order to assess whether the proposed model develops orientation maps that match the structure of those found in real 
 animals we need to utilize an automatic metric that tells how close the maps are to animal data. To this end we will 
@@
 in biological orientation maps scales linearly with hypercolumn size across many different species \citep{Kaschube2010}.
-Specifically the pinwheel density per hypercolumn area ($\Lambda^2$) converges to $\pi$, when averaged across sufficiently 
-large cortical surface. For a detailed description of the procedure for calculating this metric we refer reader to 
+Specifically the pinwheel density per hypercolumn area ($\Lambda^2$) converges to $\pi$, when averaged across a sufficiently 
+large cortical surface. For a detailed description of the procedure for calculating this metric we refer the reader to 
 our previous work \citep{Stevens2013}, but briefly, its calculation involves three steps. First, the locations of the pinwheels
@@
 hypercolumn size is determined as the peak in the isotropic ring-like Fourier transform of the orientation maps.
-Third, using these two number we can derive the pinwheel density, but to transform it to a useful metric between 
-unity (high-quality map) and zero (low quality map) we pass it through normalized Gama distribution. 
+Third, using these two numbers we can derive the pinwheel density, but to transform it to a useful metric between 
+unity (high-quality map) and zero (low quality map) we pass it through a normalized Gamma distribution. 
 We have shown that this metric reliably distinguishes low and high quality maps \citep{Stevens2013}
@@
 In this article we will proceed through multiple gradually more complex models of orientation development, addressing several 
-of the major issues with modeling of this phenomena, eventually demonstrating that the experimentally identified fast inhibitory loop is a satisfactory explanation for how short-range inhibition can support development of cortical functional organization.
+of the major issues with modeling this phenomenon, eventually demonstrating that the experimentally identified fast inhibitory loop is a satisfactory explanation for how short-range inhibition can support the development of cortical functional organization.
 We will use two different computational abstractions to explore the questions at hand.
-In the fist part of the study we will use computational model that is derived from the LISSOM family of models \citep{CMVC}.
+In the fist part of the study we will use a computational model that is derived from the LISSOM family of models \citep{CMVC}.
 This choice has three advantages.
 It allows for a very straightforward explanation of why
-fast inhibitory loop enables short-range inhibition to induce competition. It makes our explanation directly comparable to the 
-extensive set of published LISSOM family models, and thus demonstrates that the solution proposed here generalizes to range of other 
+a fast inhibitory loop enables short-range inhibition to induce competition. It makes our explanation directly comparable to the 
+extensive set of published LISSOM family models, and thus demonstrates that the solution proposed here generalizes to a range of other 
 functional properties. And finally the LISSOM abstraction enables very fast simulations, thus allowing us to perform 
-parameter search analysis that would be otherwise computationally prohibitive.
-However, as we will explain further in section \ref{sec:SM4}, some simplifications made by the LISSOM abstraction, specifically the instantaneous translation of the neuronal input into it's activity, will leave
+a parameter search analysis that would otherwise be computationally prohibitive.
+However, as we will explain further in section \ref{sec:SM4}, some simplifications made by the LISSOM abstraction, specifically the instantaneous translation of the neuronal input into its activity, will leave
 certain questions unanswered. These will be addressed in section \ref{sec:SM4} using more detailed rate model framework.
 
-\subsection{Fast excitatory to inhibitory to excitatory loop enables competition in networks with short-range inhibitory connections.} \label{sec:SM1}
+\subsection{A fast excitatory to inhibitory to excitatory loop enables competition in networks with short-range inhibitory connections.} \label{sec:SM1}
 
@@
 combined di-synaptic delay from excitatory to inhibitory and inhibitory to excitatory cells is approximately as long as the mono-synaptic delay
-of the excitatory to excitatory connections. Drawing from this, the core insight of this study is that under such pattern of delays
+of the excitatory to excitatory connections. Drawing from this, the core insight of this study is that under such a pattern of delays
 the effective inhibitory interactions are well approximated by the convolution of the excitatory and inhibitory connection kernels, as we will
@@
 \item There is no inhibitory to inhibitory interaction
-\item The synaptic inputs into the excitatory and inhibitory neurons are instantaneously translated into their rate response via positive rectified transfer function.
+\item The synaptic inputs into the excitatory and inhibitory neurons are instantaneously translated into their rate response via a positive rectified transfer function.
 \item The connection kernels of all neurons are Gaussian kernels with spatial constant $\sigma_{e}$ for excitatory neurons and $\sigma_{i}$ for inhibitory neurons.
-\item We will assume only local connectivity (disregarding long-range excitatory connections), and assume that both excitatory and inhibitory have the same extent, thus $\sigma_{e} = \sigma_{i}$ 
+\item We will assume only local connectivity (disregarding long-range excitatory connections), and assume that both excitatory and inhibitory neurons have the same extent, thus $\sigma_{e} = \sigma_{i}$
 \end{enumerate}
 
-These conditions lead to following set of equations governing the cortico-cortical interaction.
+These conditions lead to the following set of equations governing the cortico-cortical interaction.
 \begin{align}
@@
 
-\noindent where $R_{p}(x,t)$ is the response of neuron of type $p$ located at position $x$ at time $t$, $I_{aff}(x,t)$ is the afferent input to neuron at position $x$ at time $t$, $N_{\sigma}$
-is normal distribution of variance $\sigma$, corresponding to the lateral connection kernel, and $\theta_{ab}$ is delay on connections from
+\noindent where $R_{p}(x,t)$ is the response of a neuron of type $p$ located at position $x$ at time $t$, $I_{aff}(x,t)$ is the afferent input to a neuron at position $x$ at time $t$, $N_{\sigma}$
+is a normal distribution of variance $\sigma$, corresponding to the lateral connection kernel, and $\theta_{ab}$ is a delay on connections from
 neural type $a$ to neural type $b$. When we expand for $R_{i}$ we obtain:
@@
 
-\noindent and thus due to symmetry of normal distribution:
+\noindent and thus due to the symmetry of the normal distribution:
 
@@
 
-\noindent Because convolution of two Normal distributions of variance $\sigma_{i}$  and $\sigma_{e}$ is normal distribution with variance $\sqrt{\sigma_{i}^2 + \sigma_{e}^2}$ and 
+\noindent Because convolution of two Normal distributions of variance $\sigma_{i}$ and $\sigma_{e}$ is a normal distribution with variance $\sqrt{\sigma_{i}^2 + \sigma_{e}^2}$ and 
 because $\sigma_{i} = \sigma_{e}$ we can simplify to:
@@
 
-This shows that under these specific assumptions the effective inhibitory interactions between excitatory neurons are $\sqrt{2}$ longer than excitatory and thus follow the Mexican-hat like profile required for lateral cortical competition underlying functional map development.
+This shows that under these specific assumptions the effective inhibitory interactions between excitatory neurons are $\sqrt{2}$ longer than the excitatory ones and thus follow the Mexican-hat like profile required for the lateral cortical competition underlying functional map development.
 Even though this essential condition of cortical competition is fulfilled in this model configuration, important constraints on the extent of the effective lateral
@@
 Conveniently, the lateral interaction in the LISSOM family of models is governed by the same equations as \ref{eqn:first}.
-In the following we will use the GCAL model \citep{Stevens2013}, the latest and most robust addition to the LISSOM family of models to demonstrate that the above specific configuration of effective lateral excitation and inhibition permits development of high-quality orientation maps.
+In the following we will use the GCAL model \citep{Stevens2013}, the latest and most robust addition to the LISSOM family of models to demonstrate that the above specific configuration of effective lateral excitation and inhibition permits the development of high-quality orientation maps.
 GCAL is the only model of stimulus dependent functional development which achieves emergence of biologically realistic
 orientation maps in terms of pinwheel density, a signature that is a useful objective measure of orientation map quality. We will utilize this
-map quality measure to reliably detect model configurations which permit successful emergence of orientation maps.
-To facilitate comparison we will use exactly the same model configuration 
-as in Stevens \etal\,\citep{Stevens2013} (see Materials and Methods), except to three modifications necessitated by the analysis above (see figure \ref{fig:model_variants}A):
+map quality measure to reliably detect model configurations which permit the successful emergence of orientation maps.
+To facilitate a comparison we will use exactly the same model configuration 
+as in Stevens \etal\,\citep{Stevens2013} (see Materials and Methods), except for three modifications necessitated by the analysis above (see figure \ref{fig:model_variants}A):
 
@@
 
-\item The change in point 1 will results in change in the balance of overall excitation and inhibition in the model which is critical 
-for successful functional development in the model. We will thus modify the strength of the lateral inhibition to compensate for changes due to the modification 1.
+\item The change in point 1 will result in a change in the balance of overall excitation and inhibition in the model which is critical 
+for successful functional development in the model. We will thus modify the strength of the lateral inhibition to compensate for changes due to modification 1.
 
-\item Analogously to 2, the changes in 1 change the overall balance between feedforward and lateral contributions to model cortical neuron's activity, 
+\item Analogously to 2, the changes in 1 change the overall balance between feedforward and lateral contributions to model a cortical neuron's activity, 
 which we will compensate by changing the overall strength of the lateral interactions.
@@
 
-In order to find working combination of parameters in point's 2 and 3, and also to show that the model is robust to certain level of changes
-in these two parameters we have performed a parameter search across these two parameters and evaluated the quality of the orientation map (see Materials and Methods) for each 
+In order to find a working combination of parameters in points 2 and 3, and also to show that the model is robust to a certain level of changes
+in these two parameters, we have performed a parameter search across these two parameters and evaluated the quality of the orientation map (see Materials and Methods) for each 
 parameter combination. As figure \ref{fig:figure1} shows, under a range of values of both parameters the model develops high-quality orientation maps indistinguishable
-from their experimental counterparts, thus concluding our first step towards showing that fast excitatory-to-inhibitory-to-excitatory loop can explain 
-how short-range inhibition can induce cortical competition and consequently development of topological organization of functional properties.
+from their experimental counterparts, thus concluding our first step towards showing that a fast excitatory-to-inhibitory-to-excitatory loop can explain 
+how short-range inhibition can induce cortical competition and consequently the development of topological organization of functional properties.
 
@@
 \includegraphics[width=16cm]{./SVG/Figure1/figure1.png}
-\caption{Orientation map development with short range inhibition and fast excitatory-to-inhibitory-to-excitatory loop. (A) Map quality (see Materials and Methods) at a range of 
+\caption{Orientation map development with short range inhibition and a fast excitatory-to-inhibitory-to-excitatory loop. (A) Map quality (see Materials and Methods) at a range of 
 lateral excitatory vs inhibitory projection strength ratios and afferent vs lateral projection strength ratios. (B-D) Functional organization in 3 example parameter configurations 
@@
 (B-C) Two examples of sub-optimal orientation maps. (D) Model configuration with the highest quality map found in this parameter search. 
-(E) The same model configuration but in this case ran with explicit simulation of inhibitory neurons and corresponding connections. E and D are nearly identical confirming correctness of our analysis.
+(E) The same model configuration but in this case it was run with explicit simulation of inhibitory neurons and corresponding connections.
+E and D are nearly identical confirming the correctness of our analysis.
 }
@@
 interactions between them, thus corresponding to equation \ref{eqn:last}. Even though above we have shown that equation \ref{eqn:last} is equivalent to equation \ref{eqn:first} to verify 
-correctness of our analysis we have ran a single simulation of the GCAL model corresponding to the parameter combination with the highest map 
-quality found in figure \ref{fig:figure1}, but with explicitly simulated inhibitory population (figure \ref{fig:figure1}E). In this model we thus do not model direct inhibitory 
-interactions between excitatory neurons, but instead add excitatory to inhibitory and inhibitory to excitatory connections of the same extend as those of excitatory to excitatory pathway (see the assumption \#5).
+the correctness of our analysis we have run a single simulation of the GCAL model corresponding to the parameter combination with the highest map 
+quality found in figure \ref{fig:figure1}, but with an explicitly simulated inhibitory population (figure \ref{fig:figure1}E). In this model we thus do not model direct inhibitory 
+interactions between excitatory neurons, but instead add excitatory to inhibitory and inhibitory to excitatory connections of the same extent as those of the excitatory to excitatory pathway (see the assumption \#5).
 As we have shown above this model should be mathematically equivalent to the simulation shown in figure \ref{fig:figure1}D. Note, however, that the GCAL simulations represent a discrete approximation 
@@
 In the previous section we have shown that under a set of specific assumptions that facilitate analytical treatment, short-range inhibition can induce 
-effective Mexican-hat like interaction and thus support development of orientation maps. However not all assumptions we 
+effective Mexican-hat like interaction and thus support the development of orientation maps. However, not all assumptions we 
 made were in line with experimental evidence. In this section we will show, now only numerically, that one of these assumptions is not necessary, 
-specifically that addition of inhibitory to inhibitory connections does not prevent emergence of orientation maps.
+specifically that the addition of inhibitory to inhibitory connections does not prevent emergence of orientation maps.
 
-To this end we will use the exact GCAL model configuration that we have found in previous section to posses the highest
-quality orientation map (see figure \ref{fig:figure1}). We will use GCAL configuration in which we will explicitly model the inhibitory neurons
+To this end, we will use the exact GCAL model configuration that we have found in the previous section to possess the highest
+quality orientation map (see figure \ref{fig:figure1}). We will use the GCAL configuration in which we will explicitly model the inhibitory neurons
 (see figure \ref{fig:model_variants}B, and equation \ref{eqn:first}) and consequently also explicitly the excitatory to inhibitory and inhibitory to excitatory connections.
@@
 \includegraphics[width=16cm]{./SVG/Figure2/figure2.png}
-\caption{Orientation map development with short range inhibition, fast excitatory-to-inhibitory-to-excitatory loop and inhibitory to inhibitory connections.
+\caption{Orientation map development with short range inhibition, a fast excitatory-to-inhibitory-to-excitatory loop and inhibitory to inhibitory connections.
 (A) Map quality (see Materials and Methods) at a range of
@@
 By explicitly modeling the inhibitory neurons in this model we have replaced a single parameter governing the strength of inhibitory
-later interactions in the model from previous section with three new parameters that set: (1) strength of excitatory to inhibitory,
+lateral interactions in the model from the previous section with three new parameters that set: (1) the strength of excitatory to inhibitory,
 (2) inhibitory to excitatory and (3) inhibitory to inhibitory connections (all other parameters remained the same as the 
-in the best parametrization found in previous section). Note that in principle, there is redundancy in these parameters as
+in the best parameterization found in the previous section). Note that in principle, there is redundancy in these parameters as
 the overall strength of the projections from inhibitory neurons onto both excitatory and inhibitory populations is scaled by the 
-excitatory to inhibitory projections strength. Therefore in figure \ref{fig:figure2} we have systematically varied the strength of the added inhibitory to inhibitory
+excitatory to inhibitory projection strength. Therefore in figure \ref{fig:figure2} we have systematically varied the strength of the added inhibitory to inhibitory
 projection expressed relatively to the strength of the inhibitory to inhibitory connections (which was set to 1), while also varying the 
@@
 \includegraphics[width=16cm]{./SVG/Figure3/figure3.png}
-\caption{Orientation map development with short range inhibition, fast excitatory-to-inhibitory-to-excitatory loop and long-range excitation. (A) Map quality (see Materials and Methods) at a range of 
+\caption{Orientation map development with short range inhibition, a fast excitatory-to-inhibitory-to-excitatory loop and long-range excitation.
+(A) Map quality (see Materials and Methods) at a range of
 short vs.
-long range excitatory connection strength ratios (y-axis) and range of afferent vs.
+long range excitatory connection strength ratios (y-axis) and a range of afferent vs.
 lateral connection strength ratios (x-axis).
@@
 However experimental evidence shows that excitatory cells send longer connections compared to inhibitory cells \citep{Buzas2006,Budd2001}.
-In this section we will explore what happens if we add long-range excitatory connectivity into the model variant 1.
+In this section we will explore what happens if we add long-range excitatory connectivity into model variant 1.
 Buz\'as \etal\,\citep{Buzas2006} have
-shown that the lateral connectivity in layer 2/3 can be best described as superimposition of two gabor connectivity likelihoods, one short-range not orientation 
-specific and one long-range orientation specific. Here we will assume such dual structure, leading us to add second excitatory to excitatory and excitatory to 
-inhibitory projection into the model variant 1, but with 3 times larger space constant (see figure \ref{fig:model_variants}C), in line with Buz\'as \etal \, \citep{Buzas2006} quantitative findings.
+shown that the lateral connectivity in layer 2/3 can be best described as superimposition of two gabor connectivity likelihoods, one short-range but not orientation 
+specific and one long-range and orientation specific. Here we will assume such dual structure, leading us to add a second excitatory to excitatory and excitatory to 
+inhibitory projection into model variant 1, but with a space constant that is 3 times larger (see figure \ref{fig:model_variants}C), in line with Buz\'as \etal \, \citep{Buzas2006} quantitative findings.
 
-In figure \ref{fig:figure3} we will examine what strength of the long-range excitatory connections, relative to the short-range excitatory ones (y axis in figure \ref{fig:figure3}A), leads to development of high-quality orientation map.
+In figure \ref{fig:figure3} we will examine what strength of the long-range excitatory connections, relative to the short-range excitatory ones (y axis in figure \ref{fig:figure3}A), leads to development of a high-quality orientation map.
 Adding the long-range excitation changes the spatial configurations over which excitatory interactions win over inhibitory ones, as well as the 
@@
 long-range excitatory projections.
-As can be seen in figure \ref{fig:figure3}D and E, under appropriately strong lateral interactions, substantial long-range excitatory connection still permit development of high quality orientation maps, demonstrating that the proposed model is consistent with the experimentally identified long-range excitatory connectivity.
+As can be seen in figure \ref{fig:figure3}D and E, under appropriately strong lateral interactions, substantial long-range excitatory connections still permit the development of high quality orientation maps, demonstrating that the proposed model is consistent with the experimentally identified long-range excitatory connectivity.
 
-\subsection{Non-equal effective excitatory and inhibitory delay} \label{sec:SM4}
+\subsection{Non-equal effective excitatory and inhibitory delays} \label{sec:SM4}
 
@@
 to the sum of excitatory to inhibitory and inhibitory to excitatory delays. This assumption is approximately supported by the experimental evidence 
-\citep{Ohana2012}, but we cannot assume it holds exactly in real biological substrate.
-However, we hypothesize, that the small discrepancies between the delay of the mono-synaptic excitatory and cumulative delay of the bi-synaptic inhibitory interactions can be absorbed in the membrane time-constant of the neurons.
-In this section we will verify this hypothesis by extending the modeling framework used thus far with a finite membrane time-constant (see Materials and Methods) and proceed to determine the magnitude of discrepancy in the delays between the excitatory and inhibitory interactions that can be managed by the model without impairing the resulting map orientation map quality.
+\citep{Ohana2012}, but we cannot assume it holds exactly in a real biological substrate.
+However, we hypothesize, that the small discrepancies between the delay of the mono-synaptic excitatory connections and the cumulative delay of the bi-synaptic inhibitory interactions can be absorbed into the membrane time-constant of the neurons.
+In this section we will verify this hypothesis by extending the modeling framework used thus far with a finite membrane time-constant (see Materials and Methods) and proceed to determine the magnitude of the discrepancy in the delays between the excitatory and inhibitory interactions that can be managed by the model without impairing the resulting orientation map quality.
 
-We use model parametrization similar to those determined for the
+We use a model parameterization similar to those determined for 
 model variant 1 (section \ref{sec:SM1}; figure \ref{fig:model_variants}A).
@@
 The membrane time-constant of excitatory neurons was set to \insval{membrane_time_constant_E}ms while those of inhibitory neurons to \insval{membrane_time_constant_I}ms.
-This faster inhibitory dynamics are necessary to prevent oscillations in the system \citep{Kang2003}.
+These faster inhibitory dynamics are necessary to prevent oscillations in the system \citep{Kang2003}.
 We set the excitatory to excitatory delay to 1.4 ms and excitatory to inhibitory delay to 0.5 ms based on Ohana \etal\,\citep{Ohana2012}.
-In order to understand how closely does the
-cumulative bi-synaptic inhibition delay has to match that of the direct excitatory to excitatory delay, we will vary the delay on the inhibitory to excitatory projection (note that we could achieve the same by varying excitatory
-to inhibitory delay, and this choice was arbitrary).
+In order to understand how closely the
+cumulative bi-synaptic inhibition delay has to match that of the direct excitatory to excitatory delay, we will vary the delay on the inhibitory to excitatory projection (note that we could achieve the same by varying the excitatory
+to inhibitory delays, and this choice was arbitrary).
 
@@
 \includegraphics[width=16cm]{./SVG/Figure4/figure4.png}
-\caption{Orientation map development in a rate model with synaptic time constant. (A) Map quality (see Materials and Methods) at a range of 
-cumulative inhibitory delays expressed as the difference from the direct excitatory to excitatory delay. (B-F) Functional organization in 5 parameter configurations 
+\caption{Orientation map development in a rate model with a synaptic time constant. (A) Map quality (see Materials and Methods) at a range of 
+cumulative inhibitory delays expressed as the difference from the direct excitatory to excitatory delays. (B-F) Functional organization in 5 parameter configurations 
 indicated by the red marks. The orientation map (top), and afferent connection fields from the ON LGN model sheet for 81 example model V1 neurons (bottom). 
-(B-D) Three examples of configuration where good quality orientation maps develop. (E) If the cumulative inhibitory delay is longer by more than approximately 0.8ms in comparison
-to the direct excitatory delay the map quality starts to drop. (F) In the configuration corresponding to the case where delays on all connections are equal (i.e. the cumulative delay
+(B-D) Three examples of configurations where good quality orientation maps develop. (E) If the cumulative inhibitory delay is longer by more than approximately 0.8ms in comparison
+to the direct excitatory delay the map quality starts to drop. (F) In the configuration corresponding to the case where delays on all connections are equal (i.e. the cumulative delays
 of the inhibitory interactions is twice as long as on the direct excitatory ones) orientation maps fail to develop.}
@@
 
-Figure \ref{fig:figure4} shows the resulting orientation maps and associated map quality measures of models with range of differences between the delays of monosynaptic excitatory and bi-synaptic inhibitory interactions, that are in the figure expressed as the sum of the excitatory to inhibitory and inhibitory to excitatory delays minus the excitatory to excitatory delay (i.e.
-the figure shows how much slower the inhibitory disynaptic interactions were in comparison to the monosynaptic excitatory ones).
+Figure \ref{fig:figure4} shows the resulting orientation maps and associated map quality measures of models with a range of differences between the delays of monosynaptic excitatory and bi-synaptic inhibitory interactions, that are in the figure expressed as the sum of the excitatory to inhibitory and inhibitory to excitatory delays minus the excitatory to excitatory delay (i.e.
+the figure shows how much slower the inhibitory disynaptic interactions were in comparison with the monosynaptic excitatory ones).
 As can be seen when the differences between the excitatory delay (1.4ms) and cumulative inhibitory delay is small (\textless 0.8 ms) high quality orientation maps develop in the model, confirming that sufficiently
 large discrepancy between the direct excitatory and bi-synaptic inhibitory delays can be accommodated in the model.
-However, as expected, as the difference between the delays increases the ability of the model to learn topological organized representation of orientation preference diminishes.
-Crucially, if the delays across all the projections would be equal (figure \ref{fig:figure4}), as is typically assumed, the model fails to develop orientation maps in line with the analysis by Muir \etal \, \citep{Muir2014}, thus confirming that the specific delay pattern between neural types identified by Ohana \etal\,\citep{Ohana2012} is key to achieving competitive dynamics in topologically organized neural models.
+However, as expected, as the difference between the delays increases the ability of the model to learn a topologically organized representation of orientation preference diminishes.
+Crucially, if the delays across all the projections were equal (figure \ref{fig:figure4}), as is typically assumed, the model fails to develop orientation maps in line with the analysis by Muir \etal \, \citep{Muir2014}, thus confirming that the specific delay pattern between neural types identified by Ohana \etal\,\citep{Ohana2012} is key to achieving competitive dynamics in topologically organized neural models.
 
@@
 In this study we have shown how recent findings on dependence of neural transmission delays on the
-type of pre- and post-synaptic neuron \citep{Ohana2012} can resolve long standing question on how short-range inhibition
-can support cortical competition and consequently the development cortical functional topological organization.
+type of pre- and post-synaptic neurons \citep{Ohana2012} can resolve a long-standing question on how short-range inhibition
+can support cortical competition and consequently the development of functional cortical topological organization.
 Under simplifying assumptions, we have analytically shown how disynaptic inhibition that is as fast as mono-synaptic 
@@
 that the proposed mechanisms are robust to the variations of the exact delay ratio between the mono-synaptic excitation and di-synaptic inhibition. 
-Overall, this study represents an important advance in our understanding how orientation map development can be supported by cortical neural substrate. These
+Overall, this study represents an important advance in our understanding of how orientation map development can be supported by the cortical neural substrate. These
 results generalize to the development of other functional features in the cortex and other cortical competition based mechanisms in general.
@@
 The most related past explanation of how cortical competition can arise under short range inhibition is that of Kang \etal\,\citep{Kang2003}, who have shown 
-that under the assumption of faster inhibitory time constant (as opposed to excitatory), the effective excitatory and inhibitory interactions will follow the 
+that under the assumption of a faster inhibitory time constant (as opposed to excitatory), the effective excitatory and inhibitory interactions will follow the 
 Mexican hat profile and thus support competition along the cortical surface. Indeed, in our final model variant 4 that explicitly considers membrane time-constant presented in section \ref{sec:SM4}, 
 we assume that inhibitory neurons have faster membrane time constants than excitatory ones, as otherwise we observe oscillatory behavior in line with the analytical findings of Kang \etal\,\citep{Kang2003}.
-Crucially, Kang \etal\,\citep{Kang2003} however assumed instantaneous neural transmission, and when this biologically implausible assumption is rectified by addition of transmission delays that are uniform
+Crucially, Kang \etal\,\citep{Kang2003} assumed instantaneous neural transmission, and when this biologically implausible assumption is rectified by addition of transmission delays that are uniform 
 across the connections between the different pre- and post-synaptic neural types, we find that the competitive dynamics in the neural model break (figure \ref{fig:figure4}F) in line with the analytical and
@@
 The analytical results in this study were obtained only under simplifying assumptions, specifically instantaneous translation of inputs to membrane potential, equal extent of 
-excitatory and inhibitory connections and lack of inhibitory-to-inhibitory interactions. Even though we have shown computationally that these assumptions are not necessary 
-for achieving the cortical competition and the consequential orientation map development sought in these study, further analytical work than can circumvent these simplifications would 
-undoubtedly provide deeper understanding of the dynamics of the studied neural system and its dependence on the various parameters. This sentiment is underlaid by the parameter 
-explorations presented here, which show that even though the model is robust to changes of the considered parameters, the existence or not of dynamics supporting development of 
+excitatory and inhibitory connections and a lack of inhibitory-to-inhibitory interactions. Even though we have shown computationally that these assumptions are not necessary 
+for achieving the cortical competition and the consequential orientation map development sought in this study, further analytical work then can circumvent these simplifications and would 
+undoubtedly provide deeper understanding of the dynamics of the studied neural system and its dependence on the various parameters. This sentiment underlies the parameter 
+explorations presented here, which show that even though the model is robust to changes in the considered parameters, the existence or not of dynamics supporting development of 
 orientation maps forms a complex pattern within the explored parameter spaces. Furthermore, the relative computational complexity of the studied models and the extensive set of parameters
-involved preclude systematic search across the full parameter space, and we have only explored parameters that we empirically found to have biggest impact. Finally, one simplifying assumption 
+involved preclude systematic search across the full parameter space, and we have only explored parameters that we empirically found to have the biggest impact. Finally, one simplifying assumption 
 that we have not treated in this study is the lack of direct thalamic input onto inhibitory cells. Since inhibitory cells in cortical layer 4 do receive thalamic input \citep{Binzegger2004} the 
-inclusion of external input in the inhibitory population needs to be considered in future.
+inclusion of external input in the inhibitory population needs to be considered in the future.
 
@@
 The advantage of this approach is that it allows us not only to show that some form of competition is possible, but also that it is of the form that actually supports implementation of specific cortical computations.
-Given that we show that our model implements effective Mexican hat lateral interactions (section \ref{sec:SM1}) and these have been in the past shown to be sufficient to explain cortical organization of other functional features (i.e.
+Given that we show that our model implements effective Mexican hat lateral interactions (section \ref{sec:SM1}) and these have in the past been shown to be sufficient to explain cortical organization of other functional features (i.e.
 retinotopy, ocular dominance, spatial frequency and color) it is very likely that our results will generalize to these other dimensions of sensory input as well.
-Cortical competition of other forms have been proposed to underly broad variety of other cortical operations, including associative memory, noise suppression, decision making, saliency detection and other forms of attentional computations.
+Cortical competition of other forms has been proposed to underly a broad variety of other cortical operations, including associative memory, noise suppression, decision making, saliency detection and other forms of attentional computations.
 Even though additional work will be required to determine if the mechanisms proposed here can generalize to these other neural computations, this study offers
-promising framework for anatomically plausible mechanistic explanations of these important aspect of brain function.
+a promising framework for anatomically plausible mechanistic explanations of these important aspects of brain function.
 

### 2017-02-23 paper.tex

@@
 excitation) while long-range inhibition facilitates competition between the local populations and stabilizes the activity in the 
-network. Such so called Mexican-hat arrangement of recurrent interactions, however, is in stark contrast to the know anatomical arrangement
+network. Such so called Mexican-hat arrangement of recurrent interactions (figure \ref{fig:lat_int}A), however, is in stark contrast to the know anatomical arrangement
 of cortical circuitry: the excitatory neurons (especially in superficial layers) tend to form long-range arborizations spanning 
 multiple columns while the axons of a majority of inhibitory neurons are confined to an area only several hundreds of micrometers 
-in diameter \citep{Buzas2006,Budd2001}.
+in diameter \citep{Buzas2006,Budd2001} (figure \ref{fig:lat_int}B).
 
-As pointed out by Muir \etal\, in their recent study \citep{Muir2014}, to solve this apparent conflict, previous topologically organized models of cortical competition have either relied on the anatomically unsupported Mexican-hat profile of lateral interactions \citep{VonderMalsburg1973,CMVC}, or relied on other biologically unrealistic properties, such as selective targeting of inhibitory neurons by long range excitatory connections \citep{law:phd09,Rutishauser2012} or instantaneous synaptic transmission coupled with omission of recurrent inhibition \citep{Kang2003,Levy2011,Grabska-Barwinska2008}.
+As pointed out by Muir \etal\, in their recent study \citep{Muir2014}, to solve this apparent conflict, previous topologically organized models of cortical competition have either relied on the anatomically unsupported
+Mexican-hat profile of lateral interactions \citep{VonderMalsburg1973,CMVC}, or relied on other biologically unrealistic properties, such as selective targeting of inhibitory neurons by long range excitatory
+connections \citep{law:phd09,Rutishauser2012} or instantaneous synaptic transmission coupled with omission of recurrent inhibition \citep{Kang2003,Levy2011,Grabska-Barwinska2008}.
 
-A common hypothesis about how to reconcile Mexican-hat lateral interactions with anatomical reality is that the range of the effective inhibitory influence of an excitatory neuron onto other excitatory neurons (via the disynaptic pathway from excitatory to inhibitory to excitatory neurons - the disynaptic inhibition) will correspond to the cumulative reach of the direct excitatory and the direct inhibitory connections.
-The effective
+A common hypothesis about how to reconcile Mexican-hat lateral interactions with anatomical reality is that the range of the effective inhibitory influence of an excitatory neuron 
+onto other excitatory neurons (via the disynaptic pathway from excitatory to inhibitory to excitatory neurons - the disynaptic inhibition) will correspond to the cumulative reach 
+of the direct excitatory and the direct inhibitory connections. The effective
 inhibitory interactions between excitatory neurons will thus have longer range than the direct excitatory connections, supporting the use
@@
 the recurrent disynaptic inhibition will lag the direct recurrent excitation. As it turns out this is crucial, 
-as recent model analysis by Muir \etal\,\citep{Muir2014} shows that, under the assumption of uniform transmission delays, the presence of competition across the cortical surface is predicted well by the anatomy of direct excitatory and inhibitory coupling and that multi-synaptic network effects are negligible, effectively rejecting the disynaptic explanation behind Maxican-hat interaction.
+as recent model analysis by Muir \etal\,\citep{Muir2014} shows that, under the assumption of uniform transmission delays, 
+the presence of competition across the cortical surface is predicted well by the anatomy of direct excitatory and inhibitory coupling and that multi-synaptic 
+network effects are negligible, effectively rejecting the disynaptic explanation behind Maxican-hat interaction. 
 In conclusion, currently no satisfactory explanation of how topological functional organization develops in cortical networks that is consistent with the present anatomical findings exists.
+
+\begin{figure}[htpb!] 
+\centering
+\includegraphics[width=16cm]{./SVG/FigureLateralInteractions/drawing.png}
+\caption{Cortical anatomy and effective lateral interactions. (A) Short-range excitation (red) and long-range inhibition (blue) leads to effective Mexican-hat lateral interactions (black curve). 
+(B) Anatomical evidence indicates the oposite organization of lateral connectivity in cortex, whereby excitatory neurons send long-range connections to other neurons, while inhibitory neurons only 
+local ones. Furthermore, uniform delays accross the intra-cortical projections is tipically assumed. It is not clear how such anatomical configuration can support the Mexican-hat effective interaction
+across cortical surface.
+(C) Under the assumption of slow excitatory-to-excitatory and fast excitatory-to-inhibitory and inhibitory-to-excitatory pathway (and disregarding the inhibitory-to-inhibitory interactions),
+the effective lateral inhibitory interactions correspond to the convolution of the excitatory-to-inhibitory and inhibitory-to-excitatory connection kernels.
+Uner the assumption that these kernels are Gaussian (G($\sigma$) in
+the figure), the effective inhibitory interactions will be Gaussian with variance (space-constant) greater then the excitatory-to-excitatory one, thus forming Mexican-hat lateral interaction profile.
+}
+\label{fig:lat_int}
+\end{figure}
 
@@
 to effective cortical competition.
-We employ computational models to show that the fast excitatory-to-inhibitory-to-excitatory pathway allows for the disynaptic inhibition to generate the effective Mexican-hat like lateral interactions, thus for the first time
+We employ computational models to show that the fast excitatory-to-inhibitory-to-excitatory pathway allows for the disynaptic inhibition to generate the effective Mexican-hat like lateral interactions (figure \ref{fig:lat_int}C), thus for the first time
 explaining competition in topologically organized cortical networks with no biologically implausible assumptions. To demonstrate the

### 2017-02-27 paper.tex
changes

@@
 
-%\title[Running Title]{Fast inhibitory loop explains functional map development with short-range inhibition.}
-%\author{J\'an Antol\'ik}
-
-%\begin{document}
-%\maketitle
-
 \def\keyFont{\fontsize{8}{11}\helveticabold }
@@
 \def\Address{$^{1}$ Unit\'e de Neurosciences, Information et Complexit\'e,CNRS UPR 3293, Gif-sur-Yvette, France }
-% The Corresponding Author should be marked with an asterisk
-% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
 \def\corrAuthor{J\'an Antol\'ik}
@@
 indicated by the red marks.
-From left to right, the orientation map, fast-Fourier transform of the orientation map and afferent connection fields from the ON LGN model sheet for 25 example model V1 neurons.
+From left to right, the orientation map, the lateral interaction kernel (LIK; this is calculated based on eq.
+\ref{eqn:last}), fast-Fourier transform of the orientation map and afferent connection fields from the ON LGN model sheet for 25 example model V1 neurons.
 (B-C) Two examples of sub-optimal orientation maps. (D) Model configuration with the highest quality map found in this parameter search. 
@@
 indicated by the red marks.
-From left to right, the orientation map, fast-Fourier transform of the orientation map and afferent connection fields from the ON LGN model sheet for 25 example model V1 neurons.
+From left to right, the orientation map, the lateral interaction kernel (LIK; this is calculated based on eq.
+\ref{eqn:last}), fast-Fourier transform of the orientation map and afferent connection fields from the ON LGN model sheet for 25 example model V1 neurons.
 }
@@
 \caption{Orientation map development in a rate model with a synaptic time constant. (A) Map quality (see Materials and Methods) at a range of 
-cumulative inhibitory delays expressed as the difference from the direct excitatory to excitatory delays. (B-F) Functional organization in 5 parameter configurations 
-indicated by the red marks. The orientation map (top), and afferent connection fields from the ON LGN model sheet for 81 example model V1 neurons (bottom). 
-(B-D) Three examples of configurations where good quality orientation maps develop. (E) If the cumulative inhibitory delay is longer by more than approximately 0.8ms in comparison
-to the direct excitatory delay the map quality starts to drop. (F) In the configuration corresponding to the case where delays on all connections are equal (i.e. the cumulative delays
-of the inhibitory interactions is twice as long as on the direct excitatory ones) orientation maps fail to develop.}
+cumulative inhibitory delays expressed as the difference from the direct excitatory to excitatory delays.
+(B-F) Functional organization in 5 parameter configurations indicated by the red marks.
+The orientation map (top), and afferent connection fields from the ON LGN model sheet for 81 example model V1 neurons (bottom).
+(B-D) Three examples of configurations where good quality orientation maps develop.
+(E) If the cumulative inhibitory delay is longer by more than approximately 0.8ms in comparison to the direct excitatory delay the map quality starts to drop.
+(F) In the configuration corresponding to the case where delays on all connections are equal (i.e.
+the cumulative delays of the inhibitory interactions is twice as long as on the direct excitatory ones) orientation maps fail to develop.
+The lateral interaction kernel is indicated on the left, but note that it is valid only
+for the first parametrization, corresponding to equal delays on the mono-synaptic excitation and dy-synaptic inhibition pathway.}
 \label{fig:figure4}

### 2017-02-27 paper.tex
more changes

@@
 that the proposed mechanisms are robust to the variations of the exact delay ratio between the mono-synaptic excitation and di-synaptic inhibition. 
-Overall, this study represents an important advance in our understanding of how orientation map development can be supported by the cortical neural substrate. These
-results generalize to the development of other functional features in the cortex and other cortical competition based mechanisms in general.
+Overall, this study represents an important advance in our understanding of how orientation map development can be supported by the cortical neural substrate.
+
+Using neuron reconstruction data of the recorded neurons, Ohana \etal,\citep{Ohana2012} identified the 
+placement of the different pre-post synaptic combinations on the dendrites of the target cell as the likely origin of this 
+transmission delay heterogeneity. They found that the excitatory-to-inhibitory synapses were closer to the soma
+of the post-synaptic neuron than the excitatory-to-excitatory and inhibitory-to-excitatory synapses. An axonal origin 
+for the transmission delay heterogeneity is unlikely, as Ohana \etal,\citep{Ohana2012} found that all the pre-post combinations were, on
+average, equally distant from the pre-synaptic cell body. They further supported the dendritic origin of the observed
+delay inhomogeneities by showing in a computational model that the observed delay magnitudes can be explained by
+these anatomical findings. Overall this indicates a mechanism for the generation of the differences in 
+inter-neuron transmission delays that is likely to generalize beyond the primary visual cortex. We thus suggest that the
+results presented in this study generalize to the development of other functional features and other 
+cortical competition based mechanisms whose origin is outside of primary visual cortex.
 
@@
 excitatory and inhibitory connections and a lack of inhibitory-to-inhibitory interactions. Even though we have shown computationally that these assumptions are not necessary 
-for achieving the cortical competition and the consequential orientation map development sought in this study, further analytical work then can circumvent these simplifications and would 
+for achieving the cortical competition and the consequential orientation map development sought in this study, further analytical work then can circumvent these simplifications would 
 undoubtedly provide deeper understanding of the dynamics of the studied neural system and its dependence on the various parameters. This sentiment underlies the parameter 
 explorations presented here, which show that even though the model is robust to changes in the considered parameters, the existence or not of dynamics supporting development of 
-orientation maps forms a complex pattern within the explored parameter spaces. Furthermore, the relative computational complexity of the studied models and the extensive set of parameters
+orientation maps can form a complex pattern within the explored parameter spaces. This is particularly the case for the model variant with inhibitory-to-inhibitory interactions, in which 
+(unlike in the other variants) the inhibitory population gains it's own dynamics. Furthermore, the relative computational complexity of the studied models and the extensive set of parameters
 involved preclude systematic search across the full parameter space, and we have only explored parameters that we empirically found to have the biggest impact. Finally, one simplifying assumption 
@@
 In this paper we have decided to investigate cortical competitive mechanisms through the prism of orientation map development.
-The advantage of this approach is that it allows us not only to show that some form of competition is possible, but also that it is of the form that actually supports implementation of specific cortical computations.
-Given that we show that our model implements effective Mexican hat lateral interactions (section \ref{sec:SM1}) and these have in the past been shown to be sufficient to explain cortical organization of other functional features (i.e.
-retinotopy, ocular dominance, spatial frequency and color) it is very likely that our results will generalize to these other dimensions of sensory input as well.
-Cortical competition of other forms has been proposed to underly a broad variety of other cortical operations, including associative memory, noise suppression, decision making, saliency detection and other forms of attentional computations.
-Even though additional work will be required to determine if the mechanisms proposed here can generalize to these other neural computations, this study offers
-a promising framework for anatomically plausible mechanistic explanations of these important aspects of brain function.
+The advantage of this approach is that it allows us not only to show that some form
+of competition is possible, but also that it is of the form that actually supports implementation of specific cortical computations.
+Given that we show that our model implements effective Mexican hat lateral
+interactions (section \ref{sec:SM1}) and these have in the past been shown to be sufficient to explain cortical organization of other functional features (i.e.
+retinotopy, ocular dominance, spatial frequency and color)
+it is very likely that our results will generalize to these other dimensions of sensory input as well.
+Cortical competition of other forms has been proposed to underly a broad variety of other cortical operations,
+including associative memory, noise suppression, decision making, saliency detection and other forms of attentional computations.
+Even though additional work will be required to determine if the mechanisms proposed
+here can generalize to these other neural computations, this study offers a promising framework for anatomically plausible mechanistic explanations of these important aspects of brain function.
 

### 2017-02-27 CODE/generate_figure_material.py

@@ -70,7 +70,39 @@ def figure2():
         pylab.xticks([0,len(np.unique(X))-1],[data[2][0],data[2][-1]])
         pylab.savefig('figure2.png',dpi=600)
     
-    
+
+    if True:
+        X = []
+	Y = []
+        qual = []
+
+
+        for a in os.listdir(dirr):
+            b = os.path.join(dirr,a);
+
+            if not stat.S_ISDIR(os.stat(b).st_mode):
+                continue
+	    
+	    f = open(os.path.join(b,'results.pickle'))
+            d = pickle.load(f)
+            
+	    X.append(d['lat_strength_ratio'])
+            Y.append(d['exc_inh_ratio'])
+
+            # load result file 
+            load_snapshot(os.path.join(b,'snapshot.typ'))
+    	    qual.append(numpy.mean(topo.sim["V1"].sheet_views["OrientationSelectivity"].view()[0]))
+	    print qual[-1]
+
+	data = np.histogram2d(Y, X, bins=[len(np.unique(Y)),len(np.unique(X))], weights=qual)
+        pylab.figure(dpi=600,facecolor='w',figsize=(5,5))
+        im = pylab.imshow(data[0],interpolation='none',cmap='gray')#,vmin=0.3)
+        pylab.colorbar(im,fraction=0.046, pad=0.04)
+        pylab.yticks([0,len(np.unique(Y))-1],[data[1][0],data[1][-1]])
+        pylab.xticks([0,len(np.unique(X))-1],[data[2][0],data[2][-1]])
+        pylab.savefig('figure_sel.png',dpi=600)
+
+
     if False:
        f = open('/home/jan/Doc/Papers/fast_inh_paper/DATA/GCAL_SHORTRANGE/lat_strength_ratio=2.2_exc_inh_ratio=0.85/results.pickle') 
        d = pickle.load(f)
@@ -130,7 +162,7 @@ def figure2():
        load_snapshot('/home/jan/Doc/Papers/fast_inh_paper/DATA/GCAL_SHORTRANGE/lat_strength_ratio=2.8_exc_inh_ratio=0.65/results.pickle')
        save_plotgroup("Projection",projection=topo.sim["V1"].projections().values()[0],density=1.5,saver_params={'filename_prefix' : '1'})
 
-    if True:
+    if False:
[…]

### 2017-03-02 CODE/generate_figure_material.py
added files

@@ -28,6 +28,49 @@ if False:
     0/0
 
 
+def plot_ortcs(snapshot_path,name_prefix):
+    def disable_top_right_axis(ax):
+		for loc, spine in ax.spines.iteritems():
+		    if loc in ['right', 'top']:
+	    		spine.set_color('none')  # don't draw spine
+		for tick in ax.yaxis.get_major_ticks():
+		        tick.tick2On = False
+		for tick in ax.xaxis.get_major_ticks():
+	    		tick.tick2On = False
+		ax.xaxis.tick_bottom()
+		ax.yaxis.tick_left()
+
+    def disable_left_axis(ax):
+	        for loc, spine in ax.spines.iteritems():
+	            if loc in ['left']:
+        		spine.set_color('none')  # don't draw spine
+	        for tick in ax.yaxis.get_major_ticks():
+		        tick.tick1On = False
+
+    def clean_figure():
+        import pylab
+        pylab.title('')
+        pylab.xlabel('')
+        pylab.ylabel('')
+        pylab.yticks([], [])
+        disable_top_right_axis(pylab.gca())
+        disable_left_axis(pylab.gca())
+        pylab.gca().get_legend().set_visible(False)
+        #pylab.gca().legend_.remove()
+	    
+	    
+    import topo.command.pylabplot
+    load_snapshot(snapshot_path)
+    topo.command.pylabplot.measure_or_tuning_fullfield.instance(sheet=topo.sim["V1"])()
+    topo.command.pylabplot.cyclic_tuning_curve.instance(x_axis="orientation",sheet=topo.sim["V1"],coords=[(0,0)])(); clean_figure(); pylab.savefig(name_prefix+"ORTC[0,0].png")
+    topo.command.pylabplot.cyclic_tuning_curve.instance(x_axis="orientation",sheet=topo.sim["V1"],coords=[(0.1,0.1)])(); clean_figure(); pylab.savefig(name_prefix+"ORTC[0.1,0.1].png")
[…]

### 2017-03-02 build_figures.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
 
-for x in ./SVG/Figure1/figure1.svg  ./SVG/Figure2/figure2.svg ./SVG/Figure3/figure3.svg  ./SVG/Figure4/figure4.svg ./SVG/FigureModelArchitecture/drawing.svg ./SVG/FigureModelVariants/drawing.svg ./SVG/FigureLateralInteractions/drawing.svg 
+for x in ./SVG/Figure1/figure1.svg  ./SVG/Figure2/figure2.svg ./SVG/Figure3/figure3.svg  ./SVG/Figure4/figure4.svg ./SVG/Figure5/figure5.svg ./SVG/FigureModelArchitecture/drawing.svg ./SVG/FigureModelVariants/drawing.svg ./SVG/FigureLateralInteractions/drawing.svg 
   do
   dir=`dirname $x`
   base=`basename $x .svg`

### 2017-03-02 paper.tex
new version

@@
 \caption{Cortical anatomy and effective lateral interactions. (A) Short-range excitation (red) and long-range inhibition (blue) leads to effective Mexican-hat lateral interactions (black curve). 
-(B) Anatomical evidence indicates the oposite organization of lateral connectivity in cortex, whereby excitatory neurons send long-range connections to other neurons, while inhibitory neurons only 
-local ones. Furthermore, uniform delays accross the intra-cortical projections is tipically assumed. It is not clear how such anatomical configuration can support the Mexican-hat effective interaction
+(B) Anatomical evidence indicates the opposite organization of lateral connectivity in cortex, whereby excitatory neurons send long-range connections to other neurons, while inhibitory neurons only 
+local ones. Furthermore, uniform delays across the intra-cortical projections is typically assumed. It is not clear how such anatomical configuration can support the Mexican-hat effective interaction
 across cortical surface.
@@
 the effective lateral inhibitory interactions correspond to the convolution of the excitatory-to-inhibitory and inhibitory-to-excitatory connection kernels.
-Uner the assumption that these kernels are Gaussian (G($\sigma$) in
+Under the assumption that these kernels are Gaussian (G($\sigma$) in
 the figure), the effective inhibitory interactions will be Gaussian with variance (space-constant) greater then the excitatory-to-excitatory one, thus forming Mexican-hat lateral interaction profile.
@@
 We employ computational models to show that the fast excitatory-to-inhibitory-to-excitatory pathway allows for the disynaptic inhibition to generate the effective Mexican-hat like lateral interactions (figure \ref{fig:lat_int}C), thus for the first time
-explaining competition in topologically organized cortical networks with no biologically implausible assumptions. To demonstrate the
-proposed properties, in this study we focus on models of development of orientation maps, because this is the most well explored
-example of competition driven functional cortical organization development, but these results generalize to development of other
-cortical topological properties, as well as potentially other competition based computations.
+explaining competition in topologically organized cortical networks with no biologically implausible assumptions.
+
+To demonstrate the proposed implementation of competitive interactions on a specific feature of the V1 processing, in this study we focus on models of development of functional cortical organization.
+Functional properties of neurons in primary visual cortex, such orientation, color and frequency preference is not randomly distributed along the cortical surface but rather form smooth topological maps \citep{swindale:network96,goodhill:neuron07}, whereby nearby neurons prefer similar stimulus features.
+Such smooth topological mapping of function on cortical surface is ubiquitous throughout the cortex \citep{Huth2016}, found in multiple species, and implicated in a number of functional properties of V1, and of cortex in general.
+Here we restrict our attention to orientation preference, as it is the most well explored example of competition driven functional cortical organization development, but these results generalize to development of other cortical topological properties, as well as potentially other competition based computations.
 
@@
 In sections \ref{sec:SM1} through \ref{sec:SM3} (model 1, 2 and 3; see figure \ref{fig:model_variants}) we use the GCAL model \citep{Stevens2013}  which is the most advanced variant of the LISSOM (Laterally Interconnected Synergetically Self-Organizing Map) algorithm introduced by Miikkulainen \etal\,\citep{CMVC}, which itself is based on earlier Self-Organizing Map models \citep{Kohonen1982}.
+Several mechanistic explanations of how orientation
+maps can develop in primary visual cortex have been proposed, but most, including the LISSOM family of models, involve two key ingredients: (1) stimulus
+driven Hebbian learning on the thalamo-cortical synapses, that ensures the formation of afferent connectivity pattern inducing Gabor like RFs (and consequently orientation, frequency and phase preference) to V1 neural units; and (2) a Mexican-hat-like effective lateral interactions within the  cortical population of neural units, that induce co-activation among proximate units while competition between more distal units.
 
-For clarity and consistency the following model description closely follows the methodology sections in our 
+To understand how these two mechanisms lead to development of orientation maps, consider first the initial state of the model with isotropic afferent connectivity.
+When stimulus is presented to the model for the first time, the activity of the cortical population will after few simulation steps settle into a pattern that can be best described as random placement of 'blobs', whereby nearby neurons tend to be either co-active or silent.
+At this point, the placement of the local activity centers (the `blobs`) is essentially random, determined by the interplay of the competitive influence of the lateral interactions with whatever source of variability present in the model (e.g.
+random noise in the initial connections or content of the stimulus).
+The Hebbian learning will ensure, that all neurons within the active local populations will adjust they afferent weights slightly towards the stimulus pattern appearing within their RFs.
+They will thus in subsequent iterations be slightly more activated by similar patterns.
+
+This iterative process of stimulus presentation, activity pattern formation and Hebbian adjustment of the thalamo-cortical weights will keep repeating.
+However note, that in the subsequent iterations, neurons whose afferent weights will be more similar with the currently presented stimulus falling withing their RF will tend to be more active then those whose RF at that point differs from the presented stimulus.
+Importantly even small differences of the initial activations will be magnified by the lateral competitive interactions thus overall driving nearby neurons to over time develop sharp selectivity for similar stimulus features while more distal neurons will be driven to develop selectivity to other stimulus features.
+Which features will be mapped onto
+cortical surface will depend on the exact statistics of the stimuli shown during training. It turns out that if the stimuli are natural images, or 
+other artificial stimuli with strong oriented components, smooth representation of orientation preference across the cortical surface (the orientation
+maps) will develop. In this study we will focus on this specific feature (orientation), as it is the most well studied one, but it is important to
+emphasize that given the right stimulation conditions, multiplexed representation of other features, such as ocular dominance, disparity, spatial frequency and others can develop \citep{CMVC}.
+
+In the reminder of this section we will describe one specific implementation of model with overall dynamics following the 
+broad description outlined in the previous paragraphs. For clarity and consistency the following model description closely follows the methodology sections in our 
 previous work \citep{Stevens2013,Antolik2011}. The architecture of the 3 GCAL variants presented in this 
 paper is depicted in figure \ref{fig:model_architecture} and figure \ref{fig:model_variants}A-C. The models are implemented in the Topographica simulator, freely available at www.topographica.org. 
-Each of the GCAL models consists of a set of sheets of model neural units, corresponding to: (1) the photoreceptors, (2) the combined effects of the retinal ganglion cells (RGC) and ON and OFF LGN cells.
+Each of the GCAL models consists of a set of sheets of model neural units, corresponding to: (1) the photo-receptors, (2) the combined effects of the retinal ganglion cells (RGC) and ON and OFF LGN cells.
 Furthermore each model has one V1 sheet of units with direct excitatory and inhibitory interactions (model 1; see figure \ref{fig:model_variants}), or two
@@
 All cortical sheets have nominal dimensions \insval{cortex_size}$\times$\insval{cortex_size} in sheet coordinates. The sizes of the
-RGC/LGN (\insval{lgn_size}$\times$\insval{lgn_size}) and photoreceptor (\insval{visual_field_size}$\times$\insval{visual_field_size}) sheets
+RGC/LGN (\insval{lgn_size}$\times$\insval{lgn_size}) and photo-receptor (\insval{visual_field_size}$\times$\insval{visual_field_size}) sheets
 were chosen to ensure that each unit in the receiving sheet has a complete set of connections, thus minimizing edge effects in the
 RGC/LGN and V1 sheets. The density of units per $1.0\times$1.0 area is \insval{cortex_density}$\times$\insval{cortex_density} 
-for the photoreceptor and RGC/LGN ON and OFF sheets, and \insval{cortex_density}$\times$\insval{cortex_density} for both cortical sheets.
+for the photo-receptor and RGC/LGN ON and OFF sheets, and \insval{cortex_density}$\times$\insval{cortex_density} for both cortical sheets.
 
@@
 
-The activation value $\psi_{i}$ of unit i in the photoreceptor sheet (P) is given by the gray-scale value in the chosen image at that point.
+The activation value $\psi_{i}$ of unit i in the photo-receptor sheet (P) is given by the gray-scale value in the chosen image at that point.
 In this study we use two-dimensional elongated Gaussian patterns, whose center coordinates and orientation are sampled from a uniform 
-distribution that covers the area of the photoreceptor sheet and the full range of orientations.
+distribution that covers the area of the photo-receptor sheet and the full range of orientations.
 Every \insval{settle_steps} time steps \insval{num_gaussian_stim} Gaussian patterns are superimposed to form the visual input.
@@
 
-The ON/OFF units are called RGC/LGN units because they represent the entire pathway between the retinal photoreceptors
+The ON/OFF units are called RGC/LGN units because they represent the entire pathway between the retinal photo-receptors
 and V1, including the retinal ganglion cells, LGN cells, and the connection pathways. The activation level for a unit at position $j$
@@
 %%
-The activation function $f$ is a half-wave rectifying function that ensures positive activation values, constant $\gamma_{F}=\insval{retina_to_lgn_strength}$ defines the overall strength of the afferent connections from the retina, constant $\gamma_{L}=\insval{lateral_lgn_strength}$ defines the strength of the lateral connections within the RGC/LGN sheet, constant $c=\insval{lateral_lgn_gain}$ defines the slope of the gain, $\Psi_{i}$ is the activation of unit $i$ taken from the set of photoreceptors
+The activation function $f$ is a half-wave rectifying function that ensures positive activation values, constant $\gamma_{F}=\insval{retina_to_lgn_strength}$ defines the overall strength of the afferent connections from the retina, constant $\gamma_{L}=\insval{lateral_lgn_strength}$ defines the strength of the lateral connections within the RGC/LGN sheet, constant $c=\insval{lateral_lgn_gain}$ defines the slope of the gain, $\Psi_{i}$ is the activation of unit $i$ taken from the set of photo-receptors
 from which RGC/LGN unit $j$ receives input (its connection field $F_j$), $\omega_{ij}$ is the connection weight from unit $i$ in the
 retina to unit $j$ in the RGC/LGN, and $l_{kj}$ is the lateral connection weight from RGC/LGN unit $k$ to RGC/LGN unit $j$.
-Weights from the photoreceptors to units in the ON and OFF channels are set to fixed strengths with a difference-of-Gaussians
+Weights from the photo-receptors to units in the ON and OFF channels are set to fixed strengths with a difference-of-Gaussians
 kernel ($\sigma_{center}=\insval{lgn_center_sigma}$, $\sigma_{surround}=\insval{lgn_surround_sigma}$, in sheet dimensions), with ON connection fields having a positive center and a negative surround and vice versa for OFF.
 The lateral RGC/LGN weights are 2D Gaussians with kernel size $\sigma=\insval{lgn_lateral_sigma}$.
-The center of the afferent and lateral connection field of each ON/OFF unit is mapped to the location in the photoreceptor and LGN sheet corresponding to the location of that unit in sheet coordinates, making all these projections retinotopic.
+The center of the afferent and lateral connection field of each ON/OFF unit is mapped to the location in the photo-receptor and LGN sheet corresponding to the location of that unit in sheet coordinates, making all these projections retinotopic.
 
@@
 
-In the model, as images are presented to the photoreceptors, the cortical afferent connection weights $\omega_{i,j,A}$ from the ON/OFF sheets are adjusted once per iteration (after cortical settling is completed) using a simple Hebbian learning rule.
+In the model, as images are presented to the photo-receptors, the cortical afferent connection weights $\omega_{i,j,A}$ from the ON/OFF sheets are adjusted once per iteration (after cortical settling is completed) using a simple Hebbian learning rule.
 This rule results in connections that reflect correlations between the presynaptic ON/OFF unit activities and the postsynaptic cortical response.
@@
 
-\item Analogously to 2, the changes in 1 change the overall balance between feedforward and lateral contributions to model a cortical neuron's activity, 
+\item Analogously to 2, the changes in 1 change the overall balance between feed-forward and lateral contributions to model a cortical neuron's activity, 
 which we will compensate by changing the overall strength of the lateral interactions.
@@
 We have also shown that these findings are applicable to the problem of functional development in primary visual cortex. We have then proceeded 
-to show using computational methods that the proposed models are robust to the addition of other well established features of cortical anatomy, commonly 
-ignored by similar studies, including the long-range excitatory connections and mutual inhibition among inhibitory neurons. Finally, we have shown 
-that the proposed mechanisms are robust to the variations of the exact delay ratio between the mono-synaptic excitation and di-synaptic inhibition. 
+to show using computational methods that the proposed models are robust to the addition of other well established features of cortical anatomy, commonly ignored by similar studies, including the long-range excitatory connections and mutual inhibition among inhibitory neurons.
+Finally, we have shown that the proposed mechanisms are robust to the variations of the exact delay ratio between the mono-synaptic excitation and di-synaptic inhibition.
 Overall, this study represents an important advance in our understanding of how orientation map development can be supported by the cortical neural substrate.
@@
 cortical competition based mechanisms whose origin is outside of primary visual cortex.
+
+In this study we have focused on the development of the global organization of the thalamo-cortical and cortico-cortical connectivity and 
+as a consequence the global organization of V1's most salient functional property - orientation tuning. How do, however, the individual model
+cortical units correspond to individual biological V1 neurons, e.g. do they have matching tuning properties? The detailed systematic investigation 
+of this question is outside the scope of this study; however, to offer at least a basic view of how the explored models behave at single-cell
+level, we show in figure \ref{fig:figure5} representative orientation tuning curves for all model parametrizations for which we have shown a 
+detailed view of their properties (i.e.
+we have presented their orientation map etc.).
+Figure \ref{fig:figure5} shows that orientation tuning of individual neurons  varies between the models and parametrizations, but all the models with good quality orientation maps also have reasonably realistic orientation tuning of single units.
+Several of these model parametrizations achieve sharp realistic contrast invariance of the orientation tuning width.
+At the same time, figure \ref{fig:figure5} shows that the relationship between the orientation maps and afferent connectivity patterns and the orientation tuning is not straightforward.
+For example some parametrizations with inferior maps achieve sharper tuning (e.g.
+figure \ref{fig:figure5} A2) in comparison with model parametrizations exhibiting high-quality orientation maps (e.g.
+figure \ref{fig:figure5} A3).
+Ultimately, further, more rigorous, quantification and systematic investigation of the single cell tuning properties and their dependence on the model parametrization will be necessary to fully understand these relationships.
+
+\begin{figure}[htpb!] 
+\centering
+\includegraphics[width=11cm]{./SVG/Figure5/figure5.png}
+\caption{Orientation tuning curves of individual V1 model units in models detailed in figures 4-6.
+Each row of the figure shows orientation tuning curves measured at 4 different contrasts of 5 model units, one in the center of the modelled cortical area and 4 positioned in the middle between the center and one of the 4 corners of the modelled cortical area.
+Each row corresponds to one of the model parametrizations that were detailed (i.e.
+their orientation map etc.
+was shown) in figure 3-5: tuning curves marked with A correspond to models from figure 4, B from figure 5 and C from digure 6.
+The order of the rows corresponding to a given figure 4,5 or 6 (i.e. marked with A,B or C) is the same as the order in which the model parametrization
+in the given figure were presented (i.e.
+orientation tuning curves measured in the model parametrization presented in the 2nd row of figure 5 will be displayed in the row marked as B2 in this figure).}
+\label{fig:figure5}
+\end{figure}
 
@@
 \bibliographystyle{frontiersinSCNS_ENG_HUMS}
-\bibliography{./mendeley}
+\bibliography{./mendeley,./external}
 
